---
title: "NIPS-JS"
author: "De Nguyen"
output:
  word_document: 
    toc: true
    fig_width: 8
    fig_height: 6
    keep_md: true
  html_document:
    toc: yes
    number_sections: yes
    df_print: default
    fig_width: 10
    fig_height: 6
  pdf_document:
    toc: yes
editor_options:
  markdown:
    wrap: 120
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, fig.retina = 5)
library(psych)
library(factoextra)
library(FactoMineR)
library(ggplot2)
library(dplyr)
library(cowplot)
library(broom)
library(knitr)
library(MASS)
```

# Import data

```{r , echo = F}
nips <- read.csv("NIPS.csv", header = T, row.names = 1)
nips$SUS <- as.factor(nips$SUS)
nips$EtOH <- as.factor(nips$EtOH)
dat <- nips
nips <- within(nips, SUS <- relevel(SUS, ref = 1))
nips <- within(nips, EtOH <- relevel(EtOH, ref = 1))
# nips
```

```{r}
boxplot(nips[,-c(1,2)], main = "Box plot")
boxplot(scale(nips[,-c(1,2)], center = T), main = "Box plot")
```

# Corelation test

## Pearson

```{r}
pairs.panels(
  nips[,-c(1,2,3,4,5,6,8,13,15,17)],
  pch = 21,
  cex.cor = 1.2,
  gap = 0.5,
  scale = F,
  density = T,
  ellipses = F,
  lm = F,
  smooth = F,
  main = "Pearson_Correlation_Analysis",
  cex.main = 1.5,
  font.main = 2,
  hist.col = 4,
  stars = T
)
```

## Spearman

```{r}
pairs.panels(
  nips[,-c(1,2,3,4,5,6,8,9,13,14,15,16,17,18)],
  method = "spearman",
  pch = 21,
  cex.cor = 1,
  gap = 0.5,
  scale = F,
  density = T,
  ellipses = F,
  lm = F,
  smooth = F,
  main = "Spearman_Correlation",
  cex.main = 1.5,
  font.main = 2,
  hist.col = 8,
  stars = T
)
```

```{r}
correlation_matrix <- cor(nips[,-c(1,2,3,4,5,6,8,9,13,14,15,16,17,18)], method = "spearman")
cor.mtest <- function(mat, ...) {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = "spearman", ...)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  p.mat
}

p_matrix <- cor.mtest(nips[,-c(1,2,3,4,5,6,8,9,13,14,15,16,17,18)])
decimal_places <- 3
rounded_correlation_matrix <- round(correlation_matrix, decimal_places)
rounded_p_matrix <- round(p_matrix, decimal_places)
rounded_correlation_matrix
rounded_p_matrix
write.csv(rounded_correlation_matrix, "spearman_correlation_matrix.csv")
write.csv(rounded_p_matrix, "spearman_p_matrix.csv")
```

# PCA

```{r}
res.pca <- PCA(dat[, c(11,12,13,14,15,16,17,18)], graph = F, ncp = 11)
# res.pca$var$coord

eig <- as.data.frame(res.pca[["eig"]])
eig

coord <- as.data.frame(res.pca[["var"]][["coord"]])
coord

contrib <- as.data.frame(res.pca[["var"]][["contrib"]])
contrib

cos2 <- as.data.frame(res.pca[["var"]][["cos2"]])
cos2

contrib <- as.data.frame(res.pca[["var"]][["contrib"]])
contrib

coord_ind <- as.data.frame(res.pca[["ind"]][["coord"]])
coord_ind
```

## Scree plot

```{r , echo = F}
a1 <- fviz_eig(res.pca, addlabels = T, ylim = c(1,90),
         title = "Scree plot", ylab = "Explained variances (%)"
         ) +
    theme(
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 11, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"),
        panel.border = element_rect(fill = F),
        legend.text = element_text(size = 15, face = "bold"),
        aspect.ratio = 1.7,
        )
b1 <- fviz_pca_var(res.pca,
                 title = "Variables \ncorrelation plot",
                 col.var = "cos2",
                 gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
                 repel = T,
                 arr.width = 25,) +
    theme(
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 11, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"),
        # = element_rect(fill = F),
        legend.title = element_text(size = 15, face = "bold"),
        legend.text = element_text(size = 15, face = "bold"),
        aspect.ratio = 1,
        )
c1 <- fviz_contrib(res.pca, choice = "var", axes = 1, top = 11, title = "Contribution \nDim1") +
    theme(
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"),
        panel.border = element_rect(fill = F),
        legend.text = element_text(size = 15, face = "bold"),
        aspect.ratio = 1.8,
        )
d1 <- fviz_contrib(res.pca, choice = "var", axes = 2, top = 11, title = "Contribution \nDim2") +
    theme(
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"),
        panel.border = element_rect(fill = F),
        legend.text = element_text(size = 15, face = "bold"),
        aspect.ratio = 1.8,
        )
e1 <- fviz_contrib(res.pca, choice = "var", axes = 3, top = 11, title = "Contribution \nDim3") +
    theme(
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"),
        panel.border = element_rect(fill = F),
        legend.text = element_text(size = 15, face = "bold"),
        aspect.ratio = 1.8,
        )
```

```{r}
plot_grid(a1,b1, ncol = 2, nrow = 1, labels = c("A","B"), label_size = 15,rel_widths = c(1,1), rel_heights = c(1,1))
```

```{r}
plot_grid(c1,d1,e1, ncol = 3, nrow = 1, labels = c("A","B","C"), label_size = 15,rel_widths = c(1,1), rel_heights = c(1,1))
```

## Biplot trên PC1-PC2

```{r , echo = F}
a2 <- fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(1, 2),
  geom.ind = "point",
  pointshape = 21,
  pointsize = 4,# dat$SUS,
  fill.ind = dat$SUS,
#  palette = c("red", "blue", "#747474", "green"),
  col.ind = "black",
  addEllipses = T,
#  alpha.var = "contrib",
  col.var = "black",# "contrib",
  arr.width = 25,
#  gradient.cols = c("#BBC4C2", "#7E7C73", "#464033", "#140005"),
  legend.title = list(fill = "SUS\n (%)", color = "Variable\ncontribution (%)", alpha = "Variable\ncontribution (%)", size = "SUS\n(%)"),
  mean.point = F,
  repel = T,
  ) +
  guides(fill = guide_legend(order = 1),
         alpha = guide_legend(order = 3),
         size = guide_legend(order = 2),
         color = F,
         ) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = .6,
        legend.position = "bottom"
        )

b2 <- fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(1, 2),
  geom.ind = "point",
  pointshape = 21,
  pointsize = 4,# dat$SUS,
  fill.ind = dat$EtOH,
#  palette = c("red", "blue", "#747474", "green"),
  col.ind = "black",
  addEllipses = T,
#  alpha.var = "contrib",
  col.var = "black",# "contrib",
  arr.width = 25,
#  gradient.cols = c("#BBC4C2", "#7E7C73", "#464033", "#140005"),
  legend.title = list(fill = "EtOH", color = "Variable\ncontribution (%)", alpha = "Variable\ncontribution (%)", size = "SUS\n(%)"),
  mean.point = F,
  repel = T,
  ) +
  guides(fill = guide_legend(order = 1),
         alpha = guide_legend(order = 3),
         size = guide_legend(order = 2),
         color = F,
         ) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = .6,
        legend.position = "bottom"
        )
plot_grid(a2,b2, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

## Biplot trên PC1-PC3

```{r , echo = F}
a3 <- fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(1, 3),
  geom.ind = "point",
  pointshape = 21,
  pointsize = 4,# dat$SUS,
  fill.ind = dat$SUS,
#  palette = c("red", "blue", "#747474", "green"),
  col.ind = "black",
  addEllipses = T,
#  alpha.var = "contrib",
  col.var = "black",# "contrib",
  arr.width = 25,
#  gradient.cols = c("#BBC4C2", "#7E7C73", "#464033", "#140005"),
  legend.title = list(fill = "SUS\n (%)", color = "Variable\ncontribution (%)", alpha = "Variable\ncontribution (%)", size = "SUS\n(%)"),
  mean.point = F,
  repel = T,
  ) +
  guides(fill = guide_legend(order = 1),
         alpha = guide_legend(order = 3),
         size = guide_legend(order = 2),
         color = F,
         ) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = .6,
        legend.position = "bottom"
        )

b3 <- fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(1, 3),
  geom.ind = "point",
  pointshape = 21,
  pointsize = 4,# dat$SUS,
  fill.ind = dat$EtOH,
#  palette = c("red", "blue", "#747474", "green"),
  col.ind = "black",
  addEllipses = T,
#  alpha.var = "contrib",
  col.var = "black",# "contrib",
  arr.width = 25,
#  gradient.cols = c("#BBC4C2", "#7E7C73", "#464033", "#140005"),
  legend.title = list(fill = "EtOH\n (%)", color = "Variable\ncontribution (%)", alpha = "Variable\ncontribution (%)", size = "SUS\n(%)"),
  mean.point = F,
  repel = T,
  ) +
  guides(fill = guide_legend(order = 1),
         alpha = guide_legend(order = 3),
         size = guide_legend(order = 2),
         color = F,
         ) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = .6,
        legend.position = "bottom"
        )
plot_grid(a3,b3, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

## Biplot trên PC2-PC3

```{r , echo = F}
a4 <- fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(2, 3),
  geom.ind = "point",
  pointshape = 21,
  pointsize = 4,# dat$SUS,
  fill.ind = dat$SUS,
#  palette = c("red", "blue", "#747474", "green"),
  col.ind = "black",
  addEllipses = T,
#  alpha.var = "contrib",
  col.var = "black",# "contrib",
  arr.width = 25,
#  gradient.cols = c("#BBC4C2", "#7E7C73", "#464033", "#140005"),
  legend.title = list(fill = "SUS\n (%)", color = "Variable\ncontribution (%)", alpha = "Variable\ncontribution (%)", size = "SUS\n(%)"),
  mean.point = F,
  repel = T,
  ) +
  guides(fill = guide_legend(order = 1),
         alpha = guide_legend(order = 3),
         size = guide_legend(order = 2),
         color = F,
         ) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = .6,
        legend.position = "bottom"
        )

b4 <- fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(2, 3),
  geom.ind = "point",
  pointshape = 21,
  pointsize = 4,# dat$SUS,
  fill.ind = dat$EtOH,
#  palette = c("red", "blue", "#747474", "green"),
  col.ind = "black",
  addEllipses = T,
#  alpha.var = "contrib",
  col.var = "black",# "contrib",
  arr.width = 25,
#  gradient.cols = c("#BBC4C2", "#7E7C73", "#464033", "#140005"),
  legend.title = list(fill = "EtOH\n (%)", color = "Variable\ncontribution (%)", alpha = "Variable\ncontribution (%)", size = "SUS\n(%)"),
  mean.point = F,
  repel = T,
  ) +
  guides(fill = guide_legend(order = 1),
         alpha = guide_legend(order = 3),
         size = guide_legend(order = 2),
         color = F,
         ) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = .6,
        legend.position = "bottom"
        )
plot_grid(a4,b4, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

## Biplot trên PC1-PC2 with labels

```{r , echo = F}
fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(1, 2),
  geom.ind = c("point", "text"),
  pointshape = 21,
  pointsize = 4,# dat$SUS,
  fill.ind = dat$SUS,
#  palette = c("red", "blue", "#747474", "green"),
  col.ind = "black",
  addEllipses = F,
#  alpha.var = "contrib",
  col.var = "black",# "contrib",
  arr.width = 25,
#  gradient.cols = c("#BBC4C2", "#7E7C73", "#464033", "#140005"),
  legend.title = list(fill = "SUS (%)", color = "Variable\ncontribution (%)", alpha = "Variable\ncontribution (%)", size = "SUS (%)"),
  mean.point = F,
  repel = T,
  ) +
  guides(fill = guide_legend(order = 1),
         alpha = guide_legend(order = 3),
         size = guide_legend(order = 2),
         color = F,
         ) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 13, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = .65
        )
```

```{r , echo = F}
fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(1, 2),
  geom.ind = c("point", "text"),
  pointshape = 21,
  pointsize = 4,# dat$SUS,
  fill.ind = dat$EtOH,
#  palette = c("red", "blue", "#747474", "green"),
  col.ind = "black",
  addEllipses = F,
#  alpha.var = "contrib",
  col.var = "black",# "contrib",
  arr.width = 25,
#  gradient.cols = c("#BBC4C2", "#7E7C73", "#464033", "#140005"),
  legend.title = list(fill = "EtOH (%)", color = "Variable\ncontribution (%)", alpha = "Variable\ncontribution (%)", size = "EtOH (%)"),
  mean.point = F,
  repel = T,
  ) +
  guides(fill = guide_legend(order = 1),
         alpha = guide_legend(order = 3),
         size = guide_legend(order = 2),
         color = F,
         ) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 13, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = .65
        )
```

## Biplot PC7-PC8

```{r , echo = F}
fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(7, 8),
  geom.ind = c("point", "text"),
  pointshape = 21,
  pointsize = 4,# dat$SUS,
  fill.ind = dat$SUS,
#  palette = c("red", "blue", "#747474", "green"),
  col.ind = "black",
  addEllipses = F,
#  alpha.var = "contrib",
  col.var = "black",# "contrib",
  arr.width = 25,
#  gradient.cols = c("#BBC4C2", "#7E7C73", "#464033", "#140005"),
  legend.title = list(fill = "SUS (%)", color = "Variable\ncontribution (%)", alpha = "Variable\ncontribution (%)", size = "SUS (%)"),
  mean.point = F,
  repel = T,
  ) +
  guides(fill = guide_legend(order = 1),
         alpha = guide_legend(order = 3),
         size = guide_legend(order = 2),
         color = F,
         ) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 13, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = .65
        )
```

# Normality check

Các biến tuân theo luật phân bố chuẩn \## Shapiro test

```{r}
shapiro.test(nips$Weight)
shapiro.test(nips$Diameter)
shapiro.test(nips$Height)
shapiro.test(nips$Volume)
shapiro.test(nips$Density)
shapiro.test(nips$Porosity)
shapiro.test(nips$Shrinkage)
shapiro.test(nips$Recovery)
shapiro.test(nips$Modulus)
shapiro.test(nips$A0)
shapiro.test(nips$A1)
shapiro.test(nips$t1)
shapiro.test(nips$A2)
shapiro.test(nips$t2)
shapiro.test(nips$A3)
shapiro.test(nips$t3)
shapiro.test(nips$T)
```

## qqnorm & histogram

```{r}
par(mfrow = c(3,6))
qqnorm(nips$Weight, main = "Weight"); qqline(nips$Weight)
qqnorm(nips$Diameter, main = "Diameter"); qqline(nips$Diameter)
qqnorm(nips$Height, main = "Height"); qqline(nips$Height)
qqnorm(nips$Volume, main = "Volume"); qqline(nips$Volume)
qqnorm(nips$Density, main = "Density"); qqline(nips$Density)
qqnorm(nips$Porosity, main = "Porosity"); qqline(nips$Porosity)
qqnorm(nips$Shrinkage, main = "Shrinkage"); qqline(nips$Shrinkage)
qqnorm(nips$Recovery, main = "Recovery"); qqline(nips$Recovery)
qqnorm(nips$Modulus, main = "Modulus"); qqline(nips$Modulus)
qqnorm(nips$A0, main = "A0"); qqline(nips$A0)
qqnorm(nips$A1, main = "A1"); qqline(nips$A1)
qqnorm(nips$t1, main = "t1"); qqline(nips$t1)
qqnorm(nips$A2, main = "A2"); qqline(nips$A2)
qqnorm(nips$t2, main = "t2"); qqline(nips$t2)
qqnorm(nips$A3, main = "A3"); qqline(nips$A3)
qqnorm(nips$t3, main = "t3"); qqline(nips$t3)
qqnorm(nips$T, main = "T"); qqline(nips$T)
```

# Hồi quy & ANOVA:

## Hồi quy riêng rẽ & ANOVA

### Diameter

```{r}
mod1 <- lm(nips$Diameter ~ SUS + EtOH, data = nips)
mod1 %>% summary()
mod1 %>% anova()
mod1 %>% tidy() %>% kable(digits = 3)
mod1 %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod1 %>% plot()
```

### Height

```{r}
mod2 <- lm(nips$Height ~ SUS + EtOH, data = nips)
mod2 %>% summary()
mod2 %>% anova()
mod2 %>% tidy() %>% kable(digits = 3)
mod2 %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod2 %>% plot()
```

### Volume

```{r}
mod3 <- lm(nips$Volume ~ SUS + EtOH, data = nips)
mod3 %>% summary()
mod3 %>% anova()
mod3 %>% tidy() %>% kable(digits = 3)
mod3 %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod3 %>% plot()
```

### Density

```{r}
mod4 <- lm(nips$Density ~ SUS + EtOH, data = nips)
mod4 %>% summary()
mod4 %>% anova()
mod4 %>% tidy() %>% kable(digits = 3)
mod4 %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod4 %>% plot()
```

### Porosity

```{r}
mod5 <- lm(nips$Porosity ~ SUS + EtOH, data = nips)
mod5 %>% summary()
mod5 %>% anova()
mod5 %>% tidy() %>% kable(digits = 3)
mod5 %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod5 %>% plot()
```

### Shrinkage

```{r}
mod6 <- lm(nips$Shrinkage ~ SUS + EtOH, data = nips)
mod6 %>% summary()
mod6 %>% anova()
mod6 %>% tidy() %>% kable(digits = 3)
mod6 %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod6 %>% plot()
```

### Recovery

```{r}
mod7 <- lm(nips$Recovery ~ SUS + EtOH, data = nips)
mod7 %>% summary()
mod7 %>% anova()
mod7 %>% tidy() %>% kable(digits = 3)
mod7 %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod7 %>% plot()
```

#### Recovery with interaction

```{r}
mod7a <- lm(nips$Recovery ~ SUS + EtOH + SUS*EtOH, data = nips)
mod7a %>% summary()
mod7a %>% anova()
mod7a %>% tidy() %>% kable(digits = 3)
mod7a %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod7a %>% plot()
```

### Modulus

```{r}
mod8 <- lm(nips$Modulus ~ SUS + EtOH, data = nips)
mod8 %>% summary()
mod8 %>% anova()
mod8 %>% tidy() %>% kable(digits = 3)
mod8 %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod8 %>% plot()
```

#### Modulus with interaction

```{r}
mod8a <- lm(nips$Modulus ~ SUS + EtOH + SUS*EtOH, data = nips)
mod8a %>% summary()
mod8a %>% anova()
mod8a %>% tidy() %>% kable(digits = 3)
mod8a %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod8a %>% plot()
```

### A0

```{r}
mod9 <- lm(nips$A0 ~ SUS + EtOH, data = nips)
mod9 %>% summary()
mod9 %>% anova()
mod9 %>% tidy() %>% kable(digits = 3)
mod9 %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod9 %>% plot()
```

#### A0 with interaction

```{r}
mod9a <- lm(nips$A0 ~ SUS + EtOH + SUS*EtOH, data = nips)
mod9a %>% summary()
mod9a %>% anova()
mod9a %>% tidy() %>% kable(digits = 3)
mod9a %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod9a %>% plot()
```

### A1

```{r}
mod10 <- lm(nips$A1 ~ SUS + EtOH, data = nips)
mod10 %>% summary()
mod10 %>% anova()
mod10 %>% tidy() %>% kable(digits = 3)
mod10 %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod10 %>% plot()
```

### t1

```{r}
mod11 <- lm(nips$t1 ~ SUS + EtOH, data = nips)
mod11 %>% summary()
mod11 %>% anova()
mod11 %>% tidy() %>% kable(digits = 3)
mod11 %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod11 %>% plot()
```

#### t1 with interaction

```{r}
mod11a <- lm(nips$t1 ~ SUS + EtOH + SUS*EtOH, data = nips)
mod11a %>% summary()
mod11a %>% anova()
mod11a %>% tidy() %>% kable(digits = 3)
mod11a %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod11a %>% plot()
```

### A2

```{r}
mod12 <- lm(nips$A2 ~ SUS + EtOH, data = nips)
mod12 %>% summary()
mod12 %>% anova()
mod12 %>% tidy() %>% kable(digits = 3)
mod12 %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod12 %>% plot()
```

### t2

```{r}
mod13 <- lm(nips$t2 ~ SUS + EtOH, data = nips)
mod13 %>% summary()
mod13 %>% anova()
mod13 %>% tidy() %>% kable(digits = 3)
mod13 %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod13 %>% plot()
```

#### t2 with interaction

```{r}
mod13a <- lm(nips$t2 ~ SUS + EtOH + SUS*EtOH, data = nips)
mod13a %>% summary()
mod13a %>% anova()
mod13a %>% tidy() %>% kable(digits = 3)
mod13a %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod13a %>% plot()
```

### A3

```{r}
mod14 <- lm(nips$A3 ~ SUS + EtOH, data = nips)
mod14 %>% summary()
mod14 %>% anova()
mod14 %>% tidy() %>% kable(digits = 3)
mod14 %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod14 %>% plot()
```

### t3

```{r}
mod15 <- lm(nips$t3 ~ SUS + EtOH, data = nips)
mod15 %>% summary()
mod15 %>% anova()
mod15 %>% tidy() %>% kable(digits = 3)
mod15 %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod15 %>% plot()
```

#### t3 with interaction

```{r}
mod15a <- lm(nips$t3 ~ SUS + EtOH + SUS*EtOH, data = nips)
mod15a %>% summary()
mod15a %>% anova()
mod15a %>% tidy() %>% kable(digits = 3)
mod15a %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod15a %>% plot()
```

### T

```{r}
mod16 <- lm(nips$T ~ SUS + EtOH, data = nips)
mod16 %>% summary()
mod16 %>% anova()
mod16 %>% tidy() %>% kable(digits = 3)
mod16 %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod16 %>% plot()
```

#### T with interaction

```{r}
mod16a <- lm(nips$T ~ SUS + EtOH + SUS*EtOH, data = nips)
mod16a %>% summary()
mod16a %>% anova()
mod16a %>% tidy() %>% kable(digits = 3)
mod16a %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod16a %>% plot()
```

### Residual_stress

```{r}
mod17 <- lm(nips$RS ~ SUS + EtOH, data = nips)
mod17 %>% summary()
mod17 %>% anova()
mod17 %>% tidy() %>% kable(digits = 3)
mod17 %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod17 %>% plot()
```

#### Residual_stress with interaction

```{r}
mod17a <- lm(nips$RS ~ SUS + EtOH + SUS*EtOH, data = nips)
mod17a %>% summary()
mod17a %>% anova()
mod17a %>% tidy() %>% kable(digits = 3)
mod17a %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod17a %>% plot()
```

## Hồi quy nhiều chiều & ANOVA

####lm(cbind(Modulus,A0,A1,t1,A2,t2,A3,t3) \~ SUS + EtOH)

```{r}
mod2.1 <- lm(cbind(nips$Modulus,nips$A0,nips$T,nips$Recovery) ~ SUS + EtOH, data = nips)
mod2.1 %>% summary()
mod2.1 %>% anova()
mod2.1 %>% tidy() %>% kable(digits = 3)
mod2.1 %>% anova %>% kable(digits = 3)
par(mfrow = c(2,2))
```

###lm(cbind(Modulus,A0,T,Recovery) \~ SUS + EtOH + SUS\*EtOH)

```{r}
mod2.1a <- lm(cbind(nips$Modulus,nips$A0,nips$T,nips$Recovery) ~ SUS + EtOH + SUS*EtOH, data = nips)
mod2.1a %>% summary()
mod2.1a %>% anova()
mod2.1a %>% tidy() %>% kable(digits = 3)
mod2.1a %>% anova %>% kable(digits = 3)
par(mfrow = c(2,2))
```

## Hồi quy nhiều chiều & ANOVA trên thành phần chính

### lm(cbind(Dim.1,Dim.2) \~ SUS + EtOH + SUS\*EtOH)

```{r}
score <- data.frame(res.pca$ind$coord)
mod3.1 <- lm(cbind(score$Dim.1,score$Dim.2) ~ SUS + EtOH + SUS*EtOH, data = nips)
mod3.1 %>% summary()
mod3.1 %>%  anova()
mod3.1 %>% tidy() %>% kable(digits = 3)
mod3.1 %>% anova %>% kable(digits = 3)
```

### lm(Dim.1 \~ SUS + EtOH)

```{r}
mod3.2 <- lm(score$Dim.1 ~ SUS + EtOH, data = nips)
mod3.2 %>% summary()
mod3.2 %>%  anova()
mod3.2 %>% tidy() %>% kable(digits = 3)
mod3.2 %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod3.2 %>% plot()
```

### Interaction lm(Dim.1 \~ SUS + EtOH + SUS\*EtOH)

```{r}
mod3.3 <- lm(score$Dim.1 ~ SUS + EtOH + SUS*EtOH, data = nips)
mod3.3 %>% summary()
mod3.3 %>%  anova()
mod3.3 %>% tidy() %>% kable(digits = 3)
mod3.3 %>% anova %>% kable(digits = 3)
```

```{r}
par(mfrow = c(2,2))
mod3.3 %>% plot()
```

# MANOVA

## SUS

manova(cbind(Modulus,A0,A1,t1,A2,t2,A3,t3) \~ SUS) Summary model

```{r}
mod4.1 <- manova(cbind(nips$Modulus,nips$A0,nips$A1,nips$t1,nips$A2,nips$t2,nips$A3,nips$t3) ~ SUS, data = nips)
mod4.1 %>% summary.manova(intercept = T, "Pillai")
mod4.1 %>% summary.manova(intercept = T, "Wilks")
mod4.1 %>% summary.manova(intercept = T, "Hotelling-Lawley")
mod4.1 %>% summary.manova(intercept = T, "Roy")
```

Summary.aov model

```{r}
mod4.1 %>% summary.aov()
mod4.1 %>% tidy() %>% kable(digits = 3)
```

## EtOH

manova(cbind(Modulus,A0,A1,t1,A2,t2,A3,t3) \~ EtOH) Summary model

```{r}
mod5.1 <- manova(cbind(nips$Modulus,nips$A0,nips$A1,nips$t1,nips$A2,nips$t2,nips$A3,nips$t3) ~ EtOH, data = nips)
mod5.1 %>% summary.manova(intercept = T, "Pillai")
mod5.1 %>% summary.manova(intercept = T, "Wilks")
mod5.1 %>% summary.manova(intercept = T, "Hotelling-Lawley")
mod5.1 %>% summary.manova(intercept = T, "Roy")
```

Summary.aov model

```{r}
mod5.1 %>% summary.aov()
mod5.1 %>% tidy() %>% kable(digits = 3)
```

<!-- # ASCA -->

<!-- Factor 1: SUS -->

<!-- Factor 2: EtOH -->

<!-- ```{r} -->

<!-- set.seed(1234) -->

<!-- nips.F = nips[, c(1,2)] -->

<!-- map.1 = c("0.27" = 1, "0.33" = 2, "0.44" = 3) -->

<!-- nips.F$SUS = map.1[nips.F$SUS] -->

<!-- map.2 = c("0.2" = 1, "0.5" = 2, "1" = 3) -->

<!-- nips.F$EtOH = map.2[nips.F$EtOH] -->

<!-- names(nips.F)[1] = paste("1") -->

<!-- names(nips.F)[2] = paste("2") -->

<!-- nips.F = as.matrix(nips.F) -->

<!-- nips.X = nips[, 8:16] -->

<!-- nips.X = as.matrix(nips.X) -->

<!-- ASCA.nips = ASCA.Calculate(nips.X, nips.F, equation.elements = "1,2,12", scaling = T) -->

<!-- ASCA.DoPermutationTest(ASCA.nips, perm = 1000) -->

<!-- ``` -->

# Boxplot

## One-Factor boxplot

### Diameter

```{r}
p1.1 <- ggplot(dat, aes(x = SUS, y = Diameter)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p2.1 <- ggplot(dat, aes(x = EtOH, y = Diameter)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p1.1,p2.1, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Height

```{r}
p1.2 <- ggplot(dat, aes(x = SUS, y = Height)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p2.2 <- ggplot(dat, aes(x = EtOH, y = Height)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p1.2,p2.2, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Volume

```{r}
p1.3 <- ggplot(dat, aes(x = SUS, y = Volume)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p2.3 <- ggplot(dat, aes(x = EtOH, y = Volume)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p1.3,p2.3, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Density

```{r}
p1.4 <- ggplot(dat, aes(x = SUS, y = Density)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p2.4 <- ggplot(dat, aes(x = EtOH, y = Density)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p1.4,p2.4, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Porosity

```{r}
p1.5 <- ggplot(dat, aes(x = SUS, y = Porosity)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p2.5 <- ggplot(dat, aes(x = EtOH, y = Porosity)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p1.5,p2.5, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Shrinkage

```{r}
p1.6 <- ggplot(dat, aes(x = SUS, y = Shrinkage)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p2.6 <- ggplot(dat, aes(x = EtOH, y = Shrinkage)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p1.6,p2.6, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Recovery

```{r}
p1.7 <- ggplot(dat, aes(x = SUS, y = Recovery)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p2.7 <- ggplot(dat, aes(x = EtOH, y = Recovery)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p1.7,p2.7, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Modulus

```{r}
p1.8 <- ggplot(dat, aes(x = SUS, y = Modulus)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p2.8 <- ggplot(dat, aes(x = EtOH, y = Modulus)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p1.8,p2.8, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### A0

```{r}
p1.9 <- ggplot(dat, aes(x = SUS, y = A0)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p2.9 <- ggplot(dat, aes(x = EtOH, y = A0)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p1.9,p2.9, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### A1

```{r}
p1.10 <- ggplot(dat, aes(x = SUS, y = A1)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p2.10 <- ggplot(dat, aes(x = EtOH, y = A1)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p1.10,p2.10, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### t1

```{r}
p1.11 <- ggplot(dat, aes(x = SUS, y = t1)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p2.11 <- ggplot(dat, aes(x = EtOH, y = t1)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p1.11,p2.11, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### A2

```{r}
p1.12 <- ggplot(dat, aes(x = SUS, y = A2)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p2.12 <- ggplot(dat, aes(x = EtOH, y = A2)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p1.12,p2.12, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### t2

```{r}
p1.13 <- ggplot(dat, aes(x = SUS, y = t2)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p2.13 <- ggplot(dat, aes(x = EtOH, y = t2)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p1.13,p2.13, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### A3

```{r}
p1.14 <- ggplot(dat, aes(x = SUS, y = A3)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p2.14 <- ggplot(dat, aes(x = EtOH, y = A3)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p1.14,p2.14, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### t3

```{r}
p1.15 <- ggplot(dat, aes(x = SUS, y = t3)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p2.15 <- ggplot(dat, aes(x = EtOH, y = t3)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p1.15,p2.15, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### T

```{r}
p1.16 <- ggplot(dat, aes(x = SUS, y = T)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p2.16 <- ggplot(dat, aes(x = EtOH, y = T)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p1.16,p2.16, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

## Two-Factors boxplot

### Diameter

```{r}
p3.1 <- ggplot(dat, aes(x = EtOH, y = Diameter)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p4.1 <- ggplot(dat, aes(x = SUS, y = Diameter)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p3.1,p4.1, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Height

```{r}
p3.2 <- ggplot(dat, aes(x = EtOH, y = Height)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p4.2 <- ggplot(dat, aes(x = SUS, y = Height)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p3.2,p4.2, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Volume

```{r}
p3.3 <- ggplot(dat, aes(x = EtOH, y = Volume)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p4.3 <- ggplot(dat, aes(x = SUS, y = Volume)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p3.3,p4.3, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Density

```{r}
p3.4 <- ggplot(dat, aes(x = EtOH, y = Density)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p4.4 <- ggplot(dat, aes(x = SUS, y = Density)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p3.4,p4.4, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Porosity

```{r}
p3.5 <- ggplot(dat, aes(x = EtOH, y = Porosity)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p4.5 <- ggplot(dat, aes(x = SUS, y = Porosity)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p3.5,p4.5, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Shrinkage

```{r}
p3.6 <- ggplot(dat, aes(x = EtOH, y = Shrinkage)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p4.6 <- ggplot(dat, aes(x = SUS, y = Shrinkage)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p3.6,p4.6, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Retension

```{r}
p3.7 <- ggplot(dat, aes(x = EtOH, y = Recovery)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p4.7 <- ggplot(dat, aes(x = SUS, y = Recovery)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p3.7,p4.7, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Modulus

```{r}
p3.8 <- ggplot(dat, aes(x = EtOH, y = Modulus)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

p4.8 <- ggplot(dat, aes(x = SUS, y = Modulus)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme(legend.title = element_text(size = 14, face = "bold.italic"),
        legend.text = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 15, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 15, face = "bold"),
        panel.border = element_rect(fill = F),
        aspect.ratio = 1,
        legend.position = "bottom"
        )

plot_grid(p3.8,p4.8, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

# Box Cox transformation

## Recovery

```{r}
par(mfrow = c(3,3))
boxcox(nips$Diameter ~ 1); mtext("Diameter", side=3, line=1, at=0, col="blue", las=1)
boxcox(nips$Height ~ 1); mtext("Height", side=3, line=1, at=0, col="blue", las=1)
boxcox(nips$Volume ~ 1); mtext("Volume", side=3, line=1, at=0, col="blue", las=1)
boxcox(nips$Density ~ 1); mtext("Density", side=3, line=1, at=0, col="blue", las=1)
boxcox(nips$Porosity ~ 1); mtext("Porosity", side=3, line=1, at=0, col="blue", las=1)
boxcox(nips$Shrinkage ~ 1); mtext("Shrinkage", side=3, line=1, at=0, col="blue", las=1)
boxcox(nips$Recovery ~ 1); mtext("Recovery", side=3, line=1, at=0, col="blue", las=1)
boxcox(nips$Modulus ~ 1); mtext("Modulus", side=3, line=1, at=0, col="blue", las=1)
boxcox(nips$A0 ~ 1); mtext("A0", side=3, line=1, at=0, col="blue", las=1)
boxcox(nips$A1 ~ 1); mtext("A1", side=3, line=1, at=0, col="blue", las=1)
boxcox(nips$A2 ~ 1); mtext("A2", side=3, line=1, at=0, col="blue", las=1)
boxcox(nips$A3 ~ 1); mtext("A3", side=3, line=1, at=0, col="blue", las=1)
boxcox(nips$t1 ~ 1); mtext("t1", side=3, line=1, at=0, col="blue", las=1)
boxcox(nips$t2 ~ 1); mtext("t2", side=3, line=1, at=0, col="blue", las=1)
boxcox(nips$t3 ~ 1); mtext("t3", side=3, line=1, at=0, col="blue", las=1)
boxcox(nips$T ~ 1); mtext("T", side=3, line=1, at=0, col="blue", las=1)
```

```{r}
x <- nips$Diameter; b <- boxcox(lm(x ~ 1), plotit = F); b$x[which.max(b$y)]
x <- nips$Height; b <- boxcox(lm(x ~ 1), plotit = F); b$x[which.max(b$y)]
x <- nips$Volume; b <- boxcox(lm(x ~ 1), plotit = F); b$x[which.max(b$y)]
x <- nips$Density; b <- boxcox(lm(x ~ 1), plotit = F); b$x[which.max(b$y)]
x <- nips$Porosity; b <- boxcox(lm(x ~ 1), plotit = F); b$x[which.max(b$y)]
x <- nips$Shrinkage; b <- boxcox(lm(x ~ 1), plotit = F); b$x[which.max(b$y)]
x <- nips$Recovery; b <- boxcox(lm(x ~ 1), plotit = F); b$x[which.max(b$y)]
x <- nips$Modulus; b <- boxcox(lm(x ~ 1), plotit = F); b$x[which.max(b$y)]
x <- nips$A0; b <- boxcox(lm(x ~ 1), plotit = F); b$x[which.max(b$y)]
x <- nips$A1; b <- boxcox(lm(x ~ 1), plotit = F); b$x[which.max(b$y)]
x <- nips$A2; b <- boxcox(lm(x ~ 1), plotit = F); b$x[which.max(b$y)]
x <- nips$A3; b <- boxcox(lm(x ~ 1), plotit = F); b$x[which.max(b$y)]
x <- nips$t1; b <- boxcox(lm(x ~ 1), plotit = F); b$x[which.max(b$y)]
x <- nips$t2; b <- boxcox(lm(x ~ 1), plotit = F); b$x[which.max(b$y)]
x <- nips$t3; b <- boxcox(lm(x ~ 1), plotit = F); b$x[which.max(b$y)]
x <- nips$T; b <- boxcox(lm(x ~ 1), plotit = F); b$x[which.max(b$y)]
```
