---
title: "NIPS-JS Sponge — Data & Statistical Analysis"
author:
  - "De Nguyen ([ORCID 0000-0001-5479-5927](https://orcid.org/0000-0001-5479-5927)), Kyoto Institute of Technology"
  - "Kenji Kinashi ([ORCID 0000-0003-4342-3516](https://orcid.org/0000-0003-4342-3516)), Kyoto Institute of Technology"
lang: en
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: true
    toc_depth: 3
    number_sections: yes
    anchor_sections: true
    code_download: true
    includes:
      in_header: head-meta.html
    df_print: default
    fig_width: 10
    fig_height: 6
  word_document:
    toc: true
    fig_width: 10
    fig_height: 6
    keep_md: true
  pdf_document:
    toc: yes
editor_options:
  markdown:
    wrap: 120
---

<div style="text-align: right; margin: 0.2em 0 1.2em; font-size: 0.95em;">
<strong>English</strong> · <a href="vi.html" hreflang="vi" lang="vi">Tiếng Việt</a> · <a href="ja.html" hreflang="ja" lang="ja">日本語</a>
</div>

```{r setup, include=FALSE}
# Global chunk options: code is hidden in the rendered report (echo = FALSE), so the
# knitted document shows only figures and tables. Read this .Rmd file for the code.
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.retina = 2,
                      dev = "svglite")  # vector figures; use dev = "png" for Word/PDF output

library(psych)       # pairs.panels(): scatter-plot matrices with correlations
library(factoextra)  # fviz_*(): publication-ready PCA graphics
library(FactoMineR)  # PCA(): principal component analysis
library(ggplot2)     # grammar-of-graphics plotting
library(dplyr)       # %>% pipe used throughout
library(cowplot)     # plot_grid(): combining plots into labelled panels
library(broom)       # tidy(): turn model summaries into data frames
library(knitr)       # kable(): formatted tables in the rendered document
library(MASS)        # boxcox(): Box–Cox transformation profiles
library(car)         # Anova(type = 2): order-free tests for the summary table
library(rsm)         # response-surface models: FO()/SO(), canonical analysis, contour.lm()
# emmeans is called via emmeans::emmeans() without attaching the package.

## ---- Figure styling: one colour system for the whole document -------------------
## Each design factor gets a single-hue ramp whose light -> dark order follows the
## concentration order, so hue identifies the factor and darkness identifies the
## level in every figure: SUS = blues, EtOH = reds. Both ramps are colour-blind-safe
## (adjacent-step separation validated for deutan/protan/tritan vision). Box fills are
## drawn at alpha 0.65 so that medians/whiskers stay readable on the darkest steps; the
## overlaid dots carry the full-strength colours, and axis position encodes the level
## in every gallery panel, so no information rides on the lightened fills alone.
pal_sus  <- c("0.24" = "#5FB0E2", "0.3" = "#2171B5", "0.4" = "#08306B")
pal_etoh <- c("0.2"  = "#FB6A4A", "0.5" = "#CB181D", "1"   = "#67000D")
scale_sus  <- scale_fill_manual(values = pal_sus)
scale_etoh <- scale_fill_manual(values = pal_etoh)
scale_sus_line  <- scale_colour_manual(values = pal_sus)   # ellipse outlines in biplots
scale_etoh_line <- scale_colour_manual(values = pal_etoh)
col_accent  <- "#B2352C"                       # annotation accent: fit/reference lines
col_bar     <- "#2171B5"                       # single-series bars (scree, contributions)
pal_surface <- colorRampPalette(c("#F7FBFF", "#C9DDF0", "#8FBEE0", "#4A94CB"))(64)

## Publication base theme: white background, black axis text, no decorative bolding.
theme_pub <- theme_classic(base_size = 14) +
  theme(
    axis.text    = element_text(size = 13, face = "bold", colour = "black"),
    axis.title   = element_text(size = 15, face = "bold"),
    axis.line    = element_line(linewidth = 0.5, colour = "black"),
    axis.ticks   = element_line(linewidth = 0.5, colour = "black"),
    plot.title   = element_text(size = 16, face = "bold"),
    legend.title = element_text(size = 13, face = "bold.italic"),
    legend.text  = element_text(size = 12, face = "bold")
  )
```

```{css doc-style, echo=FALSE}
/* Typography and reading comfort (HTML output only) */
body { font-size: 16px; line-height: 1.65; color: #1a1a1a; }
h1, h2, h3, h4 { color: #08306B; }
h1 { border-bottom: 2px solid #C9DDF0; padding-bottom: .25em; }

/* Download strip: the one box a data-seeking visitor is looking for */
.download {
  background: #FDF6F5;
  border-left: 4px solid #B2352C;
  border-radius: 0 6px 6px 0;
  padding: .7em 1.1em;
  margin: 1.2em 0;
}

/* Interpretation call-outs: the discussion blocks are visually separated from
   methods text so readers can scan figure -> box -> next section */
.interpretation {
  background: #F2F7FC;
  border-left: 4px solid #2171B5;
  border-radius: 0 6px 6px 0;
  padding: .85em 1.15em;
  margin: 1.3em 0;
}
.interpretation > p:first-child { margin: 0 0 0.35em 0; }
.interpretation > p:first-child > strong:first-child { color: #2171B5; font-size: 1.04em; }

/* Tables: quiet zebra striping, strong header rule */
table { border-collapse: collapse; margin: 1em 0; }
th { border-bottom: 2px solid #08306B; padding: 5px 12px; text-align: left; }
td { padding: 4px 12px; }
tbody tr:nth-child(even) { background: #F7FAFD; }

/* Console output blocks: recessive, clearly non-prose */
pre { border-left: 3px solid #DCE7F2; background: #FAFBFD; }
```


# About this document

This notebook contains the complete dataset documentation and statistical analysis behind
the NIPS-JS sponge studies:

1. *Nonsolvent-Induced Phase Separation–Jet Spinning: An Innovative Technique for Producing
   Cellulosic Nanofilms, Suspensions, and Nanofilm-Based Sponges* (2025), ACS Omega,
   <https://doi.org/10.1021/acsomega.5c02353>
2. *Mechanically tunable nanofilm-based cellulose acetate sponges via crosslinker-free
   cryo-templating* (2025), Polymer Journal,
   <https://doi.org/10.1038/s41428-025-01083-z>

If this dataset or analysis is useful in your research, please cite the corresponding
publication(s) above — ready-to-use BibTeX entries are provided in the repository's
`README.md`.

Forty sponges were fabricated in a 3 × 3 full factorial design over two experimental
factors — the polymer suspension concentration (`SUS`: 0.24, 0.3, 0.4 wt%) and the ethanol
cosolvent concentration (`EtOH`: 0.2, 0.5, 1 wt%), with 4–7 replicates per combination
(counted from the shipped dataset; the Polymer Journal Experimental section quotes 3–5) —
and characterized by a set of physical and mechanical response variables.

The document is organized as a pipeline from raw data to models:

1. **Study data** — import and overview of all responses.
2. **Distributional checks** — Shapiro–Wilk tests, normal Q–Q plots and Box–Cox profiles,
   run before any modeling.
3. **Factor effects I: distributions by factor level** — one-factor boxplots for the
   responses and two-factor boxplots for selected responses.
4. **Factor effects II: regression & ANOVA** — a linear model for each response that enters the models (all but `Weight`) with `SUS` and
   `EtOH` as categorical predictors, plus interaction models for selected responses.
5. **Correlations between responses** — Pearson and Spearman analyses.
6. **Multivariate analysis** — PCA with biplots, multi-response regression, regression on
   component scores, MANOVA, and an ASCA partition of the multivariate variation.
7. **Response-surface analysis** — `SUS` and `EtOH` as continuous predictors: first- and
   second-order surfaces for Shrinkage, canonical analysis, a validation test against
   the EtOH = 3 wt% sponges, and
   cross-checks with ordinary regression.
8. **Structure–property scaling** — Gibson–Ashby power-law fit of relative modulus vs
   relative density.
9. **Concluding remarks** — the overall picture in brief.

Two representations of the dataset are used: the factor version `nips` (the design levels
as discrete groups; used everywhere except the response-surface part) and the numeric
version `dat` (the response-surface models need continuous predictors) — see the import
section.

::: {.download}
**Get the data and code.** Everything on this page runs from two small CSV files that
accompany it: **[`NIPS.csv`](NIPS.csv)** (the 40-sponge dataset) and
**[`NIPS_validation.csv`](NIPS_validation.csv)** (the three RSM-validation sponges) —
click to download, ready to run. The analysis source file is behind the **Code ▾**
button at the top right of this page, and the complete repository (data + code + license)
lives at <https://github.com/D-E-Nguyen/NIPS-JS-Sponge>. The papers linked above are the
place to read the science; the SI tables there contain the raw data, but these CSVs are
the ready-to-use version (see below for how the derived columns were computed). Data are
CC BY 4.0 — please cite the publications when reusing them.
:::

**Data.** The full dataset ships with this repository as `NIPS.csv` (40 sponges), licensed
CC BY 4.0 — see `DATA_LICENSE.md`. The geometry variables (Weight … Shrinkage) correspond
to Table S1 of the ACS Omega Supporting Information and the mechanical variables
(Recovery … RS) to Table S1 of the Polymer Journal Supplementary Information; both SI
tables list the same 40 sponges in the same row order. The derived geometry responses
(Volume, Density, Porosity, Shrinkage) are stored at full precision, recomputed from
Weight/Diameter/Height with the papers' equations — the SI tables print them rounded, and
the published statistics were computed on the unrounded values. Three additional sponges
used to validate the RSM prediction (EtOH = 3 wt%) are provided in `NIPS_validation.csv`.

**Column dictionary** (measurement protocols are described in the papers):

| Column | Role | Description |
|---|---|---------|
| `ID` | identifier | Sample number; becomes the row name on import. |
| `SUS` | design factor | Polymer suspension concentration (wt%; levels 0.24 / 0.3 / 0.4). |
| `EtOH` | design factor | Ethanol cosolvent concentration (wt%; levels 0.2 / 0.5 / 1). |
| `Weight` | response | Sponge weight (g); primary measurement (electronic balance). |
| `Diameter`, `Height` | response | Sponge dimensions (mm); primary measurements (caliper). |
| `Volume` | response | Bulk volume (cm³), derived as a cylinder: V = π/4 · h · d². |
| `Density` | response | Bulk density (kg·m⁻³), derived: ρ = m/V. |
| `Porosity` | response | Porosity (%), derived: 100 · (1 − ρ/ρ~CA~) with ρ~CA~ = 1311 kg·m⁻³. |
| `Shrinkage` | response | Volumetric shrinkage (%), derived: 100 · (1 − V/V~suspension~) with 15 mL of suspension. |
| `Recovery` | response | Percentage of the initial height retained after the stress-relaxation test (%). |
| `Modulus` | response | Apparent elastic modulus E\* from compression (kPa). |
| `A0`; `A1`,`t1`; `A2`,`t2`; `A3`,`t3` | response | Three-branch generalized Maxwell fit of the stress-relaxation curves: equilibrium stress σ~e~ (`A0`, kPa), branch stresses σ~1–3~ (`A1`–`A3`, kPa) and relaxation times τ~1–3~ (`t1`–`t3`, s). |
| `T` | response | Characteristic relaxation time τ (s): the sum of the constituent relaxation times, t1 + t2 + t3. Caution: `T` is also R's shorthand for `TRUE`; the column is therefore always accessed through the data frame (e.g. `nips$T`), never as a bare `T`. |
| `RS` | response | Residual stress σ~r~ (dimensionless): the normalized stress still carried by the sponge when the relaxation test ends. |

**Correspondence to the published results.** The additive models and ANOVA for Height,
Diameter, Shrinkage and Density reproduce ACS Omega Table 1 and Tables S2–S5; the one-factor
boxplots of those variables correspond to ACS Omega Figure 5. The interaction models for
Modulus, RS, T and Recovery reproduce Polymer Journal Table S2, the Spearman analysis
corresponds to Polymer Journal Figure S6 (which uses the subset Modulus, RS, T, Recovery),
and the Gibson–Ashby fit reproduces the scaling constants of Polymer Journal Figure 5b
(*C* = 2.04, *n* = 2.48). The first-order response-surface model reproduces ACS Omega Table 2,
Table S6 and Figure 6j. One caveat concerns `T` (the characteristic relaxation time τ):
the statistics for that variable in Polymer Journal Table S2 and Figure S6 came from an
earlier version of the Maxwell-fit parameters, and re-running the analysis on this final
dataset changes them noticeably. Most visibly, the published table reports a strongly
significant suspension-concentration effect on τ, whereas on this dataset that effect
weakens to the margin of significance: *P* = 0.051 in the additive model and 0.038 with
the interaction included using the sequential (Type I) tests that match the published
analysis, or 0.080 and 0.062 respectively under the order-adjusted Type II tests used in
the cross-response summary later in this document — so it clears the 0.05 threshold in
only one of the four specifications. The broader conclusions — ethanol lengthens relaxation, and
the interaction is not significant — hold on both versions, and every other variable
reproduces the published values. Analyses beyond these — the
Pearson panel, PCA, MANOVA, ASCA, multi-response models, Box–Cox profiles and the
second-order response surface — are exploratory work that supported, but did not appear
in, the publications.

# Study data

## Import

```{r import-data}
# Read the data; the first column (ID) becomes row names, so samples keep their labels
# in plots such as the labelled PCA biplots.
nips <- read.csv("NIPS.csv", header = TRUE, row.names = 1)

# Two representations of the same dataset are used in this document:
#   dat  — SUS and EtOH stay numeric: the response-surface models need continuous
#          predictors;
#   nips — SUS and EtOH become factors: everywhere else the design levels are treated
#          as discrete groups.
dat <- nips

nips$SUS  <- as.factor(nips$SUS)
nips$EtOH <- as.factor(nips$EtOH)

# Make the baseline explicit: the first (lowest) level of each factor is the reference
# group that regression coefficients are contrasted against. (Level 1 is already the
# default reference, so this is documentation more than transformation.)
nips <- within(nips, SUS  <- relevel(SUS,  ref = 1))
nips <- within(nips, EtOH <- relevel(EtOH, ref = 1))

# head(nips)  # uncomment to preview the imported data
```

## Overview boxplots

```{r overview-boxplots}
# Quick overview of every response variable (SUS and EtOH excluded).
# Left: raw values — scales differ wildly between variables.
# Right: the same data standardised (z-scores), which makes spreads comparable
# and helps to spot outliers on a common scale.
resp_vars <- setdiff(names(nips), c("SUS", "EtOH"))
par(cex.main = 1.4, font.lab = 2, cex.lab = 1.2)
boxplot(nips[, resp_vars], main = "Box plot: raw values", ylab = "Raw value", col = "#C6DBEF", las = 2, cex.axis = 0.8)
boxplot(scale(nips[, resp_vars], center = TRUE), main = "Box plot: standardized z-scores", ylab = "z-score", col = "#C6DBEF", las = 2, cex.axis = 0.8)
```

::: {.interpretation}
**Interpretation.**

The raw-scale panel mostly demonstrates how different the measurement
scales are — `t3` and `T` span hundreds of seconds while Density and the stress parameters
sit near zero — which is why the standardized panel is the more informative one. Note what
standardizing does and does not do: it forces every variable to the same standard
deviation, so box *heights* no longer say which response varies most. What they do reveal
is distribution shape, since a box is tall relative to a now-fixed standard deviation when
the data are spread towards the quartiles rather than concentrated in the middle — often
the signature of several treatment groups sitting at different levels, as with Density and
Modulus here. It is also the right mental model for the PCA later on, which likewise
standardizes the variables before extracting components.
:::

# Distributional checks

The linear models below assume approximately normal residuals, so each response is first
screened for normality.

## Shapiro–Wilk test

H₀: the sample comes from a normal distribution. A p-value **above** 0.05 means the data
show no significant departure from normality (consistent with the normal assumption); a
p-value below 0.05 flags a departure.

```{r shapiro-tests}
shapiro.test(nips$Weight)
shapiro.test(nips$Diameter)
shapiro.test(nips$Height)
shapiro.test(nips$Volume)
shapiro.test(nips$Density)
shapiro.test(nips$Porosity)
shapiro.test(nips$Shrinkage)
shapiro.test(nips$Recovery)
shapiro.test(nips$Modulus)
shapiro.test(nips$A0)
shapiro.test(nips$A1)
shapiro.test(nips$t1)
shapiro.test(nips$A2)
shapiro.test(nips$t2)
shapiro.test(nips$A3)
shapiro.test(nips$t3)
shapiro.test(nips$T)
shapiro.test(nips$RS)
```

::: {.interpretation}
**Interpretation.**

At the 0.05 level, Weight, Density, Porosity, Modulus and most of the
Maxwell parameters (`A0`–`A3`, `t1`, `t2`) depart significantly from normality, while
Diameter, Height, Volume, Shrinkage, Recovery, `t3`, `T` and `RS` do not. Two details are worth
noting. First, Density and Porosity return exactly the same statistic (*W* ≈ 0.908):
Porosity is an affine transformation of Density, and the test is invariant to such
transformations — the same happens for Volume and Shrinkage. Second, these are tests of
the *marginal* distributions, which pool all nine treatment groups: a variable that
responds strongly to SUS (such as Modulus) is a mixture of three well-separated groups,
so its pooled distribution looks multimodal and "fails" normality even if the
measurements within each condition are perfectly well behaved. The assumption that
actually matters for the models below is normality of *residuals*, which is checked per
model with the diagnostic plots.
:::

## Normal Q–Q plots

Visual counterpart of the Shapiro–Wilk test: points falling along the reference line are
consistent with normality; systematic curvature indicates skew, and points peeling off at
the ends indicate heavy tails or outliers.

```{r qq-plots}
par(mfrow = c(3, 6), cex.main = 1.6, font.axis = 2)   # 18 plots fill the 3 x 6 grid
qqnorm(nips$Weight, main = "Weight"); qqline(nips$Weight, col = col_accent, lwd = 1.2)
qqnorm(nips$Diameter, main = "Diameter"); qqline(nips$Diameter, col = col_accent, lwd = 1.2)
qqnorm(nips$Height, main = "Height"); qqline(nips$Height, col = col_accent, lwd = 1.2)
qqnorm(nips$Volume, main = "Volume"); qqline(nips$Volume, col = col_accent, lwd = 1.2)
qqnorm(nips$Density, main = "Density"); qqline(nips$Density, col = col_accent, lwd = 1.2)
qqnorm(nips$Porosity, main = "Porosity"); qqline(nips$Porosity, col = col_accent, lwd = 1.2)
qqnorm(nips$Shrinkage, main = "Shrinkage"); qqline(nips$Shrinkage, col = col_accent, lwd = 1.2)
qqnorm(nips$Recovery, main = "Recovery"); qqline(nips$Recovery, col = col_accent, lwd = 1.2)
qqnorm(nips$Modulus, main = "Modulus"); qqline(nips$Modulus, col = col_accent, lwd = 1.2)
qqnorm(nips$A0, main = "A0"); qqline(nips$A0, col = col_accent, lwd = 1.2)
qqnorm(nips$A1, main = "A1"); qqline(nips$A1, col = col_accent, lwd = 1.2)
qqnorm(nips$t1, main = "t1"); qqline(nips$t1, col = col_accent, lwd = 1.2)
qqnorm(nips$A2, main = "A2"); qqline(nips$A2, col = col_accent, lwd = 1.2)
qqnorm(nips$t2, main = "t2"); qqline(nips$t2, col = col_accent, lwd = 1.2)
qqnorm(nips$A3, main = "A3"); qqline(nips$A3, col = col_accent, lwd = 1.2)
qqnorm(nips$t3, main = "t3"); qqline(nips$t3, col = col_accent, lwd = 1.2)
qqnorm(nips$T, main = "T"); qqline(nips$T, col = col_accent, lwd = 1.2)
qqnorm(nips$RS, main = "RS"); qqline(nips$RS, col = col_accent, lwd = 1.2)
```

::: {.interpretation}
**Interpretation.**

The Q–Q panels tell the same story graphically: the variables flagged
by the Shapiro–Wilk test (e.g. Modulus, `A0`) bend away from the reference line,
consistent with the group-mixture explanation above, while variables such as Height,
Recovery and Shrinkage track the line closely.
:::

## Box–Cox transformation

A complementary check on the distribution of each modeled response (`Weight`, which
enters the analysis only through Density = m/V, is omitted). `boxcox()` profiles the
log-likelihood of the power-transformation parameter λ for the model `y^(λ) ~ 1`
(λ = 1: no transformation needed; λ = 0: log transform; λ = −1: reciprocal). If the
95% confidence interval around the peak includes 1, the untransformed variable is
adequate. Note that Box–Cox requires strictly positive data.

### Log-likelihood profiles

```{r boxcox-profiles}
par(mfrow = c(3, 3))   # 17 profiles: rendered as two 3 x 3 pages
boxcox(nips$Diameter ~ 1, lambda = seq(-8, 8, 0.1)); mtext("Diameter", side = 3, line = 1, at = 0, col = "#2171B5", las = 1, cex = 1.1, font = 2)
boxcox(nips$Height ~ 1, lambda = seq(-8, 8, 0.1)); mtext("Height", side = 3, line = 1, at = 0, col = "#2171B5", las = 1, cex = 1.1, font = 2)
boxcox(nips$Volume ~ 1, lambda = seq(-8, 8, 0.1)); mtext("Volume", side = 3, line = 1, at = 0, col = "#2171B5", las = 1, cex = 1.1, font = 2)
boxcox(nips$Density ~ 1, lambda = seq(-8, 8, 0.1)); mtext("Density", side = 3, line = 1, at = 0, col = "#2171B5", las = 1, cex = 1.1, font = 2)
boxcox(nips$Porosity ~ 1, lambda = seq(-8, 8, 0.1)); mtext("Porosity", side = 3, line = 1, at = 0, col = "#2171B5", las = 1, cex = 1.1, font = 2)
boxcox(nips$Shrinkage ~ 1, lambda = seq(-8, 8, 0.1)); mtext("Shrinkage", side = 3, line = 1, at = 0, col = "#2171B5", las = 1, cex = 1.1, font = 2)
boxcox(nips$Recovery ~ 1, lambda = seq(-8, 8, 0.1)); mtext("Recovery", side = 3, line = 1, at = 0, col = "#2171B5", las = 1, cex = 1.1, font = 2)
boxcox(nips$Modulus ~ 1, lambda = seq(-8, 8, 0.1)); mtext("Modulus", side = 3, line = 1, at = 0, col = "#2171B5", las = 1, cex = 1.1, font = 2)
boxcox(nips$A0 ~ 1, lambda = seq(-8, 8, 0.1)); mtext("A0", side = 3, line = 1, at = 0, col = "#2171B5", las = 1, cex = 1.1, font = 2)
boxcox(nips$A1 ~ 1, lambda = seq(-8, 8, 0.1)); mtext("A1", side = 3, line = 1, at = 0, col = "#2171B5", las = 1, cex = 1.1, font = 2)
boxcox(nips$A2 ~ 1, lambda = seq(-8, 8, 0.1)); mtext("A2", side = 3, line = 1, at = 0, col = "#2171B5", las = 1, cex = 1.1, font = 2)
boxcox(nips$A3 ~ 1, lambda = seq(-8, 8, 0.1)); mtext("A3", side = 3, line = 1, at = 0, col = "#2171B5", las = 1, cex = 1.1, font = 2)
boxcox(nips$t1 ~ 1, lambda = seq(-8, 8, 0.1)); mtext("t1", side = 3, line = 1, at = 0, col = "#2171B5", las = 1, cex = 1.1, font = 2)
boxcox(nips$t2 ~ 1, lambda = seq(-8, 8, 0.1)); mtext("t2", side = 3, line = 1, at = 0, col = "#2171B5", las = 1, cex = 1.1, font = 2)
boxcox(nips$t3 ~ 1, lambda = seq(-8, 8, 0.1)); mtext("t3", side = 3, line = 1, at = 0, col = "#2171B5", las = 1, cex = 1.1, font = 2)
boxcox(nips$T ~ 1, lambda = seq(-8, 8, 0.1)); mtext("T", side = 3, line = 1, at = 0, col = "#2171B5", las = 1, cex = 1.1, font = 2)
boxcox(nips$RS ~ 1, lambda = seq(-8, 8, 0.1)); mtext("RS", side = 3, line = 1, at = 0, col = "#2171B5", las = 1, cex = 1.1, font = 2)
```

### Optimal λ per response

The λ value that maximizes each profile above, collected into a single table.

```{r boxcox-lambda}
boxcox_vars <- c("Diameter", "Height", "Volume", "Density", "Porosity", "Shrinkage",
                 "Recovery", "Modulus", "A0", "A1", "A2", "A3", "t1", "t2", "t3", "T", "RS")

lambda_opt <- sapply(boxcox_vars, function(v) {
  # boxcox() refits the model internally, so the response must be reachable from the
  # formula's own environment — extract it into a local variable and use the formula
  # interface (boxcox(y ~ 1)); boxcox(lm(nips[[v]] ~ 1)) breaks under knitting because
  # the refit cannot see `v`.
  y <- nips[[v]]
  # The default search grid is lambda in [-2, 2]; several variables here peak outside it
  # and would be silently reported at the boundary, so the grid is widened. Values still
  # sitting at +/-8 below are bounds, not estimates.
  prof <- boxcox(y ~ 1, lambda = seq(-8, 8, 0.1), plotit = FALSE)
  prof$x[which.max(prof$y)]     # λ at the log-likelihood maximum
})

kable(data.frame(Variable = names(lambda_opt), `Optimal λ` = round(lambda_opt, 2),
                 check.names = FALSE),
      row.names = FALSE)
```

::: {.interpretation}
**Interpretation.**

The optimal λ varies enormously across responses, and for the stress
quantities (Modulus, `A0`–`A3`) it sits near 0, i.e. a log transformation — unsurprising,
since these variables grow multiplicatively with suspension concentration. A practical
warning is embedded in this table: `boxcox()` searches λ ∈ [−2, 2] by default, and several
of these variables peak outside that window, which the default would have reported as a
neat "−2" or "+2" rather than as the boundary of the search. The grid is widened here, and
any value still resting at the new limits should be read as a bound rather than an
estimate. These profiles also share the caveat of the Shapiro–Wilk tests:
they are computed on the pooled data (`y ~ 1`), so much of the apparent non-normality is
the factorial design itself rather than skewness within conditions. The analyses in this
document therefore stay on the original scales, where coefficients keep their physical
units, and rely on per-model residual diagnostics; re-fitting the amplitude variables on
a log scale is a sensible exercise for the interested reader.
:::

# Factor effects I: distributions by factor level

Distribution of the responses by factor level. Boxes show the median and interquartile
range; overlaid dots are the individual samples, so group sizes and spread are visible at
a glance.

```{r boxplot-theme}
# Shared theme for all boxplot panels in this section (identical styling throughout).
theme_box <- theme_pub + theme(aspect.ratio = 1, legend.position = "bottom")
```

## One-factor boxplots

Each response against one factor at a time: panel A groups by `SUS`, panel B by `EtOH`.

### Diameter

```{r box1f-diameter}
p1.1 <- ggplot(nips, aes(x = SUS, y = Diameter)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus + guides(fill = "none") +
  labs(x = "SUS (wt%)", y = "Diameter (mm)")

p2.1 <- ggplot(nips, aes(x = EtOH, y = Diameter)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh + guides(fill = "none") +
  labs(x = "EtOH (wt%)", y = "Diameter (mm)")

plot_grid(p1.1, p2.1, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Height

```{r box1f-height}
p1.2 <- ggplot(nips, aes(x = SUS, y = Height)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus + guides(fill = "none") +
  labs(x = "SUS (wt%)", y = "Height (mm)")

p2.2 <- ggplot(nips, aes(x = EtOH, y = Height)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh + guides(fill = "none") +
  labs(x = "EtOH (wt%)", y = "Height (mm)")

plot_grid(p1.2, p2.2, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Volume

```{r box1f-volume}
p1.3 <- ggplot(nips, aes(x = SUS, y = Volume)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus + guides(fill = "none") +
  labs(x = "SUS (wt%)", y = expression(bold("Volume (cm")^bold("3")*bold(")")))

p2.3 <- ggplot(nips, aes(x = EtOH, y = Volume)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh + guides(fill = "none") +
  labs(x = "EtOH (wt%)", y = expression(bold("Volume (cm")^bold("3")*bold(")")))

plot_grid(p1.3, p2.3, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Density

```{r box1f-density}
p1.4 <- ggplot(nips, aes(x = SUS, y = Density)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus + guides(fill = "none") +
  labs(x = "SUS (wt%)", y = expression(bold("Density (kg m")^bold("-3")*bold(")")))

p2.4 <- ggplot(nips, aes(x = EtOH, y = Density)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh + guides(fill = "none") +
  labs(x = "EtOH (wt%)", y = expression(bold("Density (kg m")^bold("-3")*bold(")")))

plot_grid(p1.4, p2.4, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Porosity

```{r box1f-porosity}
p1.5 <- ggplot(nips, aes(x = SUS, y = Porosity)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus + guides(fill = "none") +
  labs(x = "SUS (wt%)", y = "Porosity (%)")

p2.5 <- ggplot(nips, aes(x = EtOH, y = Porosity)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh + guides(fill = "none") +
  labs(x = "EtOH (wt%)", y = "Porosity (%)")

plot_grid(p1.5, p2.5, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Shrinkage

```{r box1f-shrinkage}
p1.6 <- ggplot(nips, aes(x = SUS, y = Shrinkage)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus + guides(fill = "none") +
  labs(x = "SUS (wt%)", y = "Shrinkage (%)")

p2.6 <- ggplot(nips, aes(x = EtOH, y = Shrinkage)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh + guides(fill = "none") +
  labs(x = "EtOH (wt%)", y = "Shrinkage (%)")

plot_grid(p1.6, p2.6, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Recovery

```{r box1f-recovery}
p1.7 <- ggplot(nips, aes(x = SUS, y = Recovery)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus + guides(fill = "none") +
  labs(x = "SUS (wt%)", y = "Recovery (%)")

p2.7 <- ggplot(nips, aes(x = EtOH, y = Recovery)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh + guides(fill = "none") +
  labs(x = "EtOH (wt%)", y = "Recovery (%)")

plot_grid(p1.7, p2.7, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Modulus

```{r box1f-modulus}
p1.8 <- ggplot(nips, aes(x = SUS, y = Modulus)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus + guides(fill = "none") +
  labs(x = "SUS (wt%)", y = "Apparent modulus E* (kPa)")

p2.8 <- ggplot(nips, aes(x = EtOH, y = Modulus)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh + guides(fill = "none") +
  labs(x = "EtOH (wt%)", y = "Apparent modulus E* (kPa)")

plot_grid(p1.8, p2.8, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### A0

```{r box1f-a0}
p1.9 <- ggplot(nips, aes(x = SUS, y = A0)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus + guides(fill = "none") +
  labs(x = "SUS (wt%)", y = expression(bold("A0, ")*bold(sigma)[bold(e)]*bold(" (kPa)")))

p2.9 <- ggplot(nips, aes(x = EtOH, y = A0)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh + guides(fill = "none") +
  labs(x = "EtOH (wt%)", y = expression(bold("A0, ")*bold(sigma)[bold(e)]*bold(" (kPa)")))

plot_grid(p1.9, p2.9, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### A1

```{r box1f-a1}
p1.10 <- ggplot(nips, aes(x = SUS, y = A1)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus + guides(fill = "none") +
  labs(x = "SUS (wt%)", y = expression(bold("A1, ")*bold(sigma)[bold(1)]*bold(" (kPa)")))

p2.10 <- ggplot(nips, aes(x = EtOH, y = A1)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh + guides(fill = "none") +
  labs(x = "EtOH (wt%)", y = expression(bold("A1, ")*bold(sigma)[bold(1)]*bold(" (kPa)")))

plot_grid(p1.10, p2.10, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### t1

```{r box1f-t1}
p1.11 <- ggplot(nips, aes(x = SUS, y = t1)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus + guides(fill = "none") +
  labs(x = "SUS (wt%)", y = expression(bold("t1, ")*bold(tau)[bold(1)]*bold(" (s)")))

p2.11 <- ggplot(nips, aes(x = EtOH, y = t1)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh + guides(fill = "none") +
  labs(x = "EtOH (wt%)", y = expression(bold("t1, ")*bold(tau)[bold(1)]*bold(" (s)")))

plot_grid(p1.11, p2.11, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### A2

```{r box1f-a2}
p1.12 <- ggplot(nips, aes(x = SUS, y = A2)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus + guides(fill = "none") +
  labs(x = "SUS (wt%)", y = expression(bold("A2, ")*bold(sigma)[bold(2)]*bold(" (kPa)")))

p2.12 <- ggplot(nips, aes(x = EtOH, y = A2)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh + guides(fill = "none") +
  labs(x = "EtOH (wt%)", y = expression(bold("A2, ")*bold(sigma)[bold(2)]*bold(" (kPa)")))

plot_grid(p1.12, p2.12, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### t2

```{r box1f-t2}
p1.13 <- ggplot(nips, aes(x = SUS, y = t2)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus + guides(fill = "none") +
  labs(x = "SUS (wt%)", y = expression(bold("t2, ")*bold(tau)[bold(2)]*bold(" (s)")))

p2.13 <- ggplot(nips, aes(x = EtOH, y = t2)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh + guides(fill = "none") +
  labs(x = "EtOH (wt%)", y = expression(bold("t2, ")*bold(tau)[bold(2)]*bold(" (s)")))

plot_grid(p1.13, p2.13, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### A3

```{r box1f-a3}
p1.14 <- ggplot(nips, aes(x = SUS, y = A3)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus + guides(fill = "none") +
  labs(x = "SUS (wt%)", y = expression(bold("A3, ")*bold(sigma)[bold(3)]*bold(" (kPa)")))

p2.14 <- ggplot(nips, aes(x = EtOH, y = A3)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh + guides(fill = "none") +
  labs(x = "EtOH (wt%)", y = expression(bold("A3, ")*bold(sigma)[bold(3)]*bold(" (kPa)")))

plot_grid(p1.14, p2.14, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### t3

```{r box1f-t3}
p1.15 <- ggplot(nips, aes(x = SUS, y = t3)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus + guides(fill = "none") +
  labs(x = "SUS (wt%)", y = expression(bold("t3, ")*bold(tau)[bold(3)]*bold(" (s)")))

p2.15 <- ggplot(nips, aes(x = EtOH, y = t3)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh + guides(fill = "none") +
  labs(x = "EtOH (wt%)", y = expression(bold("t3, ")*bold(tau)[bold(3)]*bold(" (s)")))

plot_grid(p1.15, p2.15, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### T

```{r box1f-bigT}
# Note: inside aes(), `T` refers to the data column T (data masking), not to TRUE.
p1.16 <- ggplot(nips, aes(x = SUS, y = T)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus + guides(fill = "none") +
  labs(x = "SUS (wt%)", y = expression(bold("T, ")*bold(tau)*bold(" (s)")))

p2.16 <- ggplot(nips, aes(x = EtOH, y = T)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh + guides(fill = "none") +
  labs(x = "EtOH (wt%)", y = expression(bold("T, ")*bold(tau)*bold(" (s)")))

plot_grid(p1.16, p2.16, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

## Two-factor boxplots

Both factors shown at once: one factor on the x-axis, the other as the fill color (and
vice versa in the second panel). Parallel patterns across groups suggest additive effects;
crossing patterns hint at an interaction.

### Diameter

```{r box2f-diameter}
p3.1 <- ggplot(nips, aes(x = EtOH, y = Diameter)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus +
  labs(x = "EtOH (wt%)", fill = "SUS (wt%)", y = "Diameter (mm)")

p4.1 <- ggplot(nips, aes(x = SUS, y = Diameter)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh +
  labs(x = "SUS (wt%)", fill = "EtOH (wt%)", y = "Diameter (mm)")

plot_grid(p3.1, p4.1, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Height

```{r box2f-height}
p3.2 <- ggplot(nips, aes(x = EtOH, y = Height)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus +
  labs(x = "EtOH (wt%)", fill = "SUS (wt%)", y = "Height (mm)")

p4.2 <- ggplot(nips, aes(x = SUS, y = Height)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh +
  labs(x = "SUS (wt%)", fill = "EtOH (wt%)", y = "Height (mm)")

plot_grid(p3.2, p4.2, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Volume

```{r box2f-volume}
p3.3 <- ggplot(nips, aes(x = EtOH, y = Volume)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus +
  labs(x = "EtOH (wt%)", fill = "SUS (wt%)", y = expression(bold("Volume (cm")^bold("3")*bold(")")))

p4.3 <- ggplot(nips, aes(x = SUS, y = Volume)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh +
  labs(x = "SUS (wt%)", fill = "EtOH (wt%)", y = expression(bold("Volume (cm")^bold("3")*bold(")")))

plot_grid(p3.3, p4.3, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Density

```{r box2f-density}
p3.4 <- ggplot(nips, aes(x = EtOH, y = Density)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus +
  labs(x = "EtOH (wt%)", fill = "SUS (wt%)", y = expression(bold("Density (kg m")^bold("-3")*bold(")")))

p4.4 <- ggplot(nips, aes(x = SUS, y = Density)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh +
  labs(x = "SUS (wt%)", fill = "EtOH (wt%)", y = expression(bold("Density (kg m")^bold("-3")*bold(")")))

plot_grid(p3.4, p4.4, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Porosity

```{r box2f-porosity}
p3.5 <- ggplot(nips, aes(x = EtOH, y = Porosity)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus +
  labs(x = "EtOH (wt%)", fill = "SUS (wt%)", y = "Porosity (%)")

p4.5 <- ggplot(nips, aes(x = SUS, y = Porosity)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh +
  labs(x = "SUS (wt%)", fill = "EtOH (wt%)", y = "Porosity (%)")

plot_grid(p3.5, p4.5, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Shrinkage

```{r box2f-shrinkage}
p3.6 <- ggplot(nips, aes(x = EtOH, y = Shrinkage)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus +
  labs(x = "EtOH (wt%)", fill = "SUS (wt%)", y = "Shrinkage (%)")

p4.6 <- ggplot(nips, aes(x = SUS, y = Shrinkage)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh +
  labs(x = "SUS (wt%)", fill = "EtOH (wt%)", y = "Shrinkage (%)")

plot_grid(p3.6, p4.6, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Recovery

```{r box2f-recovery}
p3.7 <- ggplot(nips, aes(x = EtOH, y = Recovery)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus +
  labs(x = "EtOH (wt%)", fill = "SUS (wt%)", y = "Recovery (%)")

p4.7 <- ggplot(nips, aes(x = SUS, y = Recovery)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh +
  labs(x = "SUS (wt%)", fill = "EtOH (wt%)", y = "Recovery (%)")

plot_grid(p3.7, p4.7, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Modulus

```{r box2f-modulus}
p3.8 <- ggplot(nips, aes(x = EtOH, y = Modulus)) + geom_boxplot(aes(fill = SUS), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_sus +
  labs(x = "EtOH (wt%)", fill = "SUS (wt%)", y = "Apparent modulus E* (kPa)")

p4.8 <- ggplot(nips, aes(x = SUS, y = Modulus)) + geom_boxplot(aes(fill = EtOH), alpha = 0.65, position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box + scale_etoh +
  labs(x = "SUS (wt%)", fill = "EtOH (wt%)", y = "Apparent modulus E* (kPa)")

plot_grid(p3.8, p4.8, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

::: {.interpretation}
**What the galleries show.**

Reading the one-factor panels across the responses, the two
factors divide the labor. Suspension concentration mainly sets *how much material* ends
up in the sponge: Density and Modulus climb steeply with SUS, and the stress amplitudes
`A0`–`A3` follow. Ethanol concentration mainly influences *how the structure forms during
freezing*: Diameter increases and Shrinkage decreases with EtOH, and the relaxation-related
quantities shown here (`T`, Recovery) shift upwards. (`RS` behaves the same way, but it has
no panel in this gallery — the regression section quantifies it.) Height rises with both factors, although
its EtOH trend falls short of significance in the model below (*P* = 0.066). In the
two-factor panels, roughly parallel box patterns suggest additive behavior for most
responses, whereas for Modulus the EtOH effect looks visibly different at each SUS level —
anticipating the significant SUS × EtOH interaction found in the next part. The formal
tests attach *P*-values to each of these impressions (the geometry responses are discussed
in the ACS Omega paper, the mechanical ones in the Polymer Journal paper).
:::

# Factor effects II: regression & ANOVA

The graphical impressions from the boxplot gallery are quantified here with linear
models on the factor version of the data.

## Per-response linear models & ANOVA

Each response is modeled as `response ~ SUS + EtOH`, i.e. an additive two-factor model
with both factors categorical. For every response, four outputs are shown:

- `summary()` — coefficient estimates (each level contrasted against the reference level),
  their significance, and the model's *R*².
- `anova()` — the ANOVA table testing the overall effect of each factor.
- The same two tables re-rendered with `tidy()`/`kable()` as compact, publication-ready
  tables.
- Standard diagnostic plots (residuals vs fitted, normal Q–Q, scale–location, residuals vs
  leverage) to check the model assumptions.

Two reading aids for the tables that follow. *Degrees of freedom*: each factor has three
levels and therefore 2 df; the additive model leaves 40 − 1 − 2 − 2 = 35 residual df, and
the interaction models add a 4-df `SUS:EtOH` term, leaving 31. *Coefficients*: with R's
default treatment coding, a coefficient such as `SUS0.3` is the estimated difference
between the 0.3 wt% group and the 0.24 wt% baseline at a fixed ethanol level, and the
intercept is the predicted response at the two baselines together (SUS 0.24, EtOH
0.2 wt%) — not the overall mean.

For selected responses, a second model adds the `SUS:EtOH` interaction to test whether the
effect of one factor depends on the level of the other. (In R's formula notation,
`SUS + EtOH + SUS*EtOH` expands to main effects plus the interaction term — it is
equivalent to `SUS*EtOH`.) The four responses analyzed with interaction in the Polymer
Journal paper (Table S2) are Modulus, RS, T and Recovery; the additive models for Height,
Diameter, Shrinkage and Density are the ones published in ACS Omega (Table 1, Tables S2–S5).

### Diameter

```{r lm-diameter}
mod1 <- lm(nips$Diameter ~ SUS + EtOH, data = nips)
mod1 %>% summary()
mod1 %>% anova()
mod1 %>% tidy() %>% kable(digits = 3)
mod1 %>% anova %>% kable(digits = 3)
```

```{r lm-diameter-diag}
par(mfrow = c(2, 2))
mod1 %>% plot()
```

### Height

```{r lm-height}
mod2 <- lm(nips$Height ~ SUS + EtOH, data = nips)
mod2 %>% summary()
mod2 %>% anova()
mod2 %>% tidy() %>% kable(digits = 3)
mod2 %>% anova %>% kable(digits = 3)
```

```{r lm-height-diag}
par(mfrow = c(2, 2))
mod2 %>% plot()
```

### Volume

```{r lm-volume}
mod3 <- lm(nips$Volume ~ SUS + EtOH, data = nips)
mod3 %>% summary()
mod3 %>% anova()
mod3 %>% tidy() %>% kable(digits = 3)
mod3 %>% anova %>% kable(digits = 3)
```

```{r lm-volume-diag}
par(mfrow = c(2, 2))
mod3 %>% plot()
```

### Density

```{r lm-density}
mod4 <- lm(nips$Density ~ SUS + EtOH, data = nips)
mod4 %>% summary()
mod4 %>% anova()
mod4 %>% tidy() %>% kable(digits = 3)
mod4 %>% anova %>% kable(digits = 3)
```

```{r lm-density-diag}
par(mfrow = c(2, 2))
mod4 %>% plot()
```

### Porosity

```{r lm-porosity}
mod5 <- lm(nips$Porosity ~ SUS + EtOH, data = nips)
mod5 %>% summary()
mod5 %>% anova()
mod5 %>% tidy() %>% kable(digits = 3)
mod5 %>% anova %>% kable(digits = 3)
```

```{r lm-porosity-diag}
par(mfrow = c(2, 2))
mod5 %>% plot()
```

### Shrinkage

```{r lm-shrinkage}
mod6 <- lm(nips$Shrinkage ~ SUS + EtOH, data = nips)
mod6 %>% summary()
mod6 %>% anova()
mod6 %>% tidy() %>% kable(digits = 3)
mod6 %>% anova %>% kable(digits = 3)
```

```{r lm-shrinkage-diag}
par(mfrow = c(2, 2))
mod6 %>% plot()
```

### Recovery

```{r lm-recovery}
mod7 <- lm(nips$Recovery ~ SUS + EtOH, data = nips)
mod7 %>% summary()
mod7 %>% anova()
mod7 %>% tidy() %>% kable(digits = 3)
mod7 %>% anova %>% kable(digits = 3)
```

```{r lm-recovery-diag}
par(mfrow = c(2, 2))
mod7 %>% plot()
```

#### Recovery with interaction

```{r lm-recovery-int}
mod7a <- lm(nips$Recovery ~ SUS + EtOH + SUS*EtOH, data = nips)
mod7a %>% summary()
mod7a %>% anova()
mod7a %>% tidy() %>% kable(digits = 3)
mod7a %>% anova %>% kable(digits = 3)
```

```{r lm-recovery-int-diag}
par(mfrow = c(2, 2))
mod7a %>% plot()
```

### Modulus

```{r lm-modulus}
mod8 <- lm(nips$Modulus ~ SUS + EtOH, data = nips)
mod8 %>% summary()
mod8 %>% anova()
mod8 %>% tidy() %>% kable(digits = 3)
mod8 %>% anova %>% kable(digits = 3)
```

```{r lm-modulus-diag}
par(mfrow = c(2, 2))
mod8 %>% plot()
```

#### Modulus with interaction

```{r lm-modulus-int}
mod8a <- lm(nips$Modulus ~ SUS + EtOH + SUS*EtOH, data = nips)
mod8a %>% summary()
mod8a %>% anova()
mod8a %>% tidy() %>% kable(digits = 3)
mod8a %>% anova %>% kable(digits = 3)
```

```{r lm-modulus-int-diag}
par(mfrow = c(2, 2))
mod8a %>% plot()
```

### A0

```{r lm-a0}
mod9 <- lm(nips$A0 ~ SUS + EtOH, data = nips)
mod9 %>% summary()
mod9 %>% anova()
mod9 %>% tidy() %>% kable(digits = 3)
mod9 %>% anova %>% kable(digits = 3)
```

```{r lm-a0-diag}
par(mfrow = c(2, 2))
mod9 %>% plot()
```

#### A0 with interaction

```{r lm-a0-int}
mod9a <- lm(nips$A0 ~ SUS + EtOH + SUS*EtOH, data = nips)
mod9a %>% summary()
mod9a %>% anova()
mod9a %>% tidy() %>% kable(digits = 3)
mod9a %>% anova %>% kable(digits = 3)
```

```{r lm-a0-int-diag}
par(mfrow = c(2, 2))
mod9a %>% plot()
```

### A1

```{r lm-a1}
mod10 <- lm(nips$A1 ~ SUS + EtOH, data = nips)
mod10 %>% summary()
mod10 %>% anova()
mod10 %>% tidy() %>% kable(digits = 3)
mod10 %>% anova %>% kable(digits = 3)
```

```{r lm-a1-diag}
par(mfrow = c(2, 2))
mod10 %>% plot()
```

### t1

```{r lm-t1}
mod11 <- lm(nips$t1 ~ SUS + EtOH, data = nips)
mod11 %>% summary()
mod11 %>% anova()
mod11 %>% tidy() %>% kable(digits = 3)
mod11 %>% anova %>% kable(digits = 3)
```

```{r lm-t1-diag}
par(mfrow = c(2, 2))
mod11 %>% plot()
```

#### t1 with interaction

```{r lm-t1-int}
mod11a <- lm(nips$t1 ~ SUS + EtOH + SUS*EtOH, data = nips)
mod11a %>% summary()
mod11a %>% anova()
mod11a %>% tidy() %>% kable(digits = 3)
mod11a %>% anova %>% kable(digits = 3)
```

```{r lm-t1-int-diag}
par(mfrow = c(2, 2))
mod11a %>% plot()
```

### A2

```{r lm-a2}
mod12 <- lm(nips$A2 ~ SUS + EtOH, data = nips)
mod12 %>% summary()
mod12 %>% anova()
mod12 %>% tidy() %>% kable(digits = 3)
mod12 %>% anova %>% kable(digits = 3)
```

```{r lm-a2-diag}
par(mfrow = c(2, 2))
mod12 %>% plot()
```

### t2

```{r lm-t2}
mod13 <- lm(nips$t2 ~ SUS + EtOH, data = nips)
mod13 %>% summary()
mod13 %>% anova()
mod13 %>% tidy() %>% kable(digits = 3)
mod13 %>% anova %>% kable(digits = 3)
```

```{r lm-t2-diag}
par(mfrow = c(2, 2))
mod13 %>% plot()
```

#### t2 with interaction

```{r lm-t2-int}
mod13a <- lm(nips$t2 ~ SUS + EtOH + SUS*EtOH, data = nips)
mod13a %>% summary()
mod13a %>% anova()
mod13a %>% tidy() %>% kable(digits = 3)
mod13a %>% anova %>% kable(digits = 3)
```

```{r lm-t2-int-diag}
par(mfrow = c(2, 2))
mod13a %>% plot()
```

### A3

```{r lm-a3}
mod14 <- lm(nips$A3 ~ SUS + EtOH, data = nips)
mod14 %>% summary()
mod14 %>% anova()
mod14 %>% tidy() %>% kable(digits = 3)
mod14 %>% anova %>% kable(digits = 3)
```

```{r lm-a3-diag}
par(mfrow = c(2, 2))
mod14 %>% plot()
```

### t3

```{r lm-t3}
mod15 <- lm(nips$t3 ~ SUS + EtOH, data = nips)
mod15 %>% summary()
mod15 %>% anova()
mod15 %>% tidy() %>% kable(digits = 3)
mod15 %>% anova %>% kable(digits = 3)
```

```{r lm-t3-diag}
par(mfrow = c(2, 2))
mod15 %>% plot()
```

#### t3 with interaction

```{r lm-t3-int}
mod15a <- lm(nips$t3 ~ SUS + EtOH + SUS*EtOH, data = nips)
mod15a %>% summary()
mod15a %>% anova()
mod15a %>% tidy() %>% kable(digits = 3)
mod15a %>% anova %>% kable(digits = 3)
```

```{r lm-t3-int-diag}
par(mfrow = c(2, 2))
mod15a %>% plot()
```

### T

```{r lm-bigT}
mod16 <- lm(nips$T ~ SUS + EtOH, data = nips)
mod16 %>% summary()
mod16 %>% anova()
mod16 %>% tidy() %>% kable(digits = 3)
mod16 %>% anova %>% kable(digits = 3)
```

```{r lm-bigT-diag}
par(mfrow = c(2, 2))
mod16 %>% plot()
```

#### T with interaction

```{r lm-bigT-int}
mod16a <- lm(nips$T ~ SUS + EtOH + SUS*EtOH, data = nips)
mod16a %>% summary()
mod16a %>% anova()
mod16a %>% tidy() %>% kable(digits = 3)
mod16a %>% anova %>% kable(digits = 3)
```

```{r lm-bigT-int-diag}
par(mfrow = c(2, 2))
mod16a %>% plot()
```

### Residual stress (RS)

```{r lm-rs}
mod17 <- lm(nips$RS ~ SUS + EtOH, data = nips)
mod17 %>% summary()
mod17 %>% anova()
mod17 %>% tidy() %>% kable(digits = 3)
mod17 %>% anova %>% kable(digits = 3)
```

```{r lm-rs-diag}
par(mfrow = c(2, 2))
mod17 %>% plot()
```

#### Residual stress (RS) with interaction

```{r lm-rs-int}
mod17a <- lm(nips$RS ~ SUS + EtOH + SUS*EtOH, data = nips)
mod17a %>% summary()
mod17a %>% anova()
mod17a %>% tidy() %>% kable(digits = 3)
mod17a %>% anova %>% kable(digits = 3)
```

```{r lm-rs-int-diag}
par(mfrow = c(2, 2))
mod17a %>% plot()
```

### Summary of the factor effects

The individual ANOVA tables above are condensed here into one map: the additive-model
*P*-values for both factors, and the model fit, across every response.

```{r factor-effect-summary}
resp_models <- list(
  Diameter = mod1, Height = mod2, Volume = mod3, Density = mod4, Porosity = mod5,
  Shrinkage = mod6, Recovery = mod7, Modulus = mod8, A0 = mod9, A1 = mod10, t1 = mod11,
  A2 = mod12, t2 = mod13, A3 = mod14, t3 = mod15, T = mod16, RS = mod17)

# Type II sums of squares (car::Anova) rather than the sequential Type I of anova():
# the replicate counts are unequal (4-7 per cell), so the two factors are not quite
# orthogonal and sequential tests would depend on the order of the terms — the factor
# entered first would be credited with the shared variation. Type II tests each factor
# adjusted for the other, which is what this side-by-side comparison needs.
# (The individual tables above keep anova() so that they match the published analyses.)
pmap <- t(sapply(resp_models, function(m) {
  a <- car::Anova(m, type = 2)
  c(a$`Pr(>F)`[1], a$`Pr(>F)`[2], summary(m)$r.squared)
}))
fmt_p <- function(p) ifelse(p < 1e-4, "< 0.0001", sprintf("%.4f", p))
kable(data.frame(`SUS (*P*)` = fmt_p(pmap[, 1]), `EtOH (*P*)` = fmt_p(pmap[, 2]),
                 `*R*²` = sprintf("%.3f", pmap[, 3]),
                 row.names = rownames(pmap), check.names = FALSE),
      align = "rrr")
```

::: {.interpretation}
**Interpretation.**

The map shows a broad division of roles. SUS governs the
material-quantity side: Density, Porosity, Modulus and all four stress amplitudes
`A0`–`A3` respond to it at *P* < 0.001 — more polymer per volume gives thicker cell walls
and a stiffer network. The two-factor models explain these responses well (*R*² from 0.79
for `A0` up to 0.96 for Modulus), with SUS carrying most of that on its own (for Modulus,
*R*² = 0.92 from suspension concentration alone). EtOH governs the structure-formation
side: Diameter, Volume/Shrinkage, Recovery and `RS` all move with ethanol at *P* < 0.05
while showing no detectable SUS effect. Among the shape-related responses (Diameter,
Height, Volume/Shrinkage), Height is the one driven primarily by SUS (*P* = 0.001). The
relaxation timescales (`t1`–`t3`, `T`) are among the least determined responses in the
table (*R*² between 0.17 and 0.29): for `t3` and `T` ethanol reaches significance
(*P* ≈ 0.03) while SUS sits just outside it (*P* ≈ 0.08), and for `t1` and `t2` neither
factor reaches it, so no factor controls the relaxation timescales strongly — note that the Polymer Journal analysis, run on the earlier Maxwell-fit
values of τ, found a clearly significant SUS effect there, one of the places where that
earlier fit and this dataset differ.

Two cautions on reading this table. First, the SUS column here uses Type II tests and therefore
differs from the sequential (Type I) values printed in the individual tables above, which
are kept in that form to match the published analyses; `T` is the clearest example (0.080
here against 0.051 above). The EtOH column, by contrast, is identical in both — EtOH
enters the sequential fits last, so its Type I sum of squares is already adjusted for
SUS. That coincidence is itself a useful illustration of what "sequential" means. Second, these are 2-degree-of-freedom omnibus tests of a
whole factor, whereas the published tables report individual level contrasts, which can
reach significance when the omnibus test does not — the ethanol effect on Height is
significant as the EtOH = 1 wt% contrast in ACS Omega Table S2 (*P* = 0.021) while the
factor-level test here gives *P* = 0.066, because the omnibus test pools that contrast with
the flat EtOH = 0.5 wt% one. The same applies to Shrinkage and Volume, where the whole-factor
SUS test above reads *P* = 0.11 even though the SUS = 0.3 wt% contrast reaches *P* = 0.049 in
ACS Omega Table S4 — so "no detectable SUS effect" in this table means no *whole-factor*
effect, not the absence of any individual contrast. Neither analysis is wrong; they answer
different questions, and the published tables report the contrasts.

Two built-in identities are worth recognizing: Porosity is an affine transformation of
Density, and Shrinkage of Volume, so each pair necessarily shares identical test results.

Finally, the interaction models above add a refinement: Modulus (*P* = 0.0002) and the
equilibrium stress `A0` (*P* = 0.001) show significant SUS × EtOH interactions, whereas
Recovery, `RS`, `T` and the three relaxation times do not. For Modulus the interaction is
easiest to read from the group means: raising ethanol from 0.2 to 0.5 wt% roughly halves the modulus
at SUS 0.24 wt% (3.7 → 1.7 kPa) but leaves it essentially unchanged at SUS 0.4 wt%
(10.6 → 10.5 kPa). In other words the ethanol penalty is not a fixed offset — thicker
cell walls largely absorb it — which is precisely what an additive model cannot express.
:::

# Correlations between responses

Pairwise associations between selected responses. Each panel figure shows scatter plots
(below the diagonal), variable histograms with density curves (diagonal), and correlation
coefficients with significance stars (above the diagonal): \*\*\* p < 0.001, \*\* p < 0.01,
\* p < 0.05.

## Pearson

Pearson's *r* measures **linear** association and is used here for the wider variable set.

```{r pearson-panel}
# Variables entering the Pearson panel (selection used in the papers).
vars_pearson <- c("Density", "Shrinkage", "Recovery", "Modulus", "A0", "t1", "t2", "t3", "T")

pairs.panels(
  nips[, vars_pearson],
  pch = 21,
  cex.cor = 1.2,     # size of the correlation labels
  gap = 0.5,
  scale = FALSE,     # do not scale label size by correlation strength
  density = TRUE,    # add density curves to the diagonal histograms
  ellipses = FALSE,
  lm = FALSE,
  smooth = FALSE,
  main = "Pearson_Correlation_Analysis",
  cex.main = 1.5,
  font.main = 2,
  hist.col = "#9ECAE1",
  cex.labels = 1.6,  # diagonal variable names
  font.labels = 2,
  cex.axis = 1.15,
  stars = TRUE       # significance stars next to the coefficients
)
```

::: {.interpretation}
**Interpretation.**

Three blocks stand out. Density, Modulus and the stress quantities
form a tightly linked group (Density–Modulus r = 0.96, Modulus–`A0` r = 0.87): the denser
the sponge, the stiffer it is and the more stress every Maxwell branch carries. The three
relaxation times form a second block (`t1`–`t2` and `t2`–`t3` both r = 0.83) that is only
weakly and negatively related to the first (r between about −0.2 and −0.5): dense, stiff
sponges tend to relax somewhat faster. And `T` is essentially collinear with `t3`
(r = 1.00) — by construction, since T = t1 + t2 + t3 and t3 is two orders of magnitude
larger than t1. Shrinkage and Recovery are only weakly correlated with the rest of the
panel (|r| ≤ 0.33) — a first hint that resilience is not a by-product of stiffness.
:::

## Spearman

Spearman's ρ is rank-based, so it captures **monotonic** (not necessarily linear)
relationships and is robust to outliers. A reduced variable set is examined here. (Note:
the published Spearman figure — Polymer Journal Figure S6 — uses the subset `Modulus`,
`RS`, `T`, `Recovery`; the panel below is the repository's variant of that analysis.)

```{r spearman-panel}
# Variables entering the Spearman panel and the exported matrices below.
# The published figure (Polymer Journal Fig. S6) used c("Modulus", "RS", "T", "Recovery").
vars_spearman <- c("Density", "Recovery", "Modulus", "A0", "T")

pairs.panels(
  nips[, vars_spearman],
  method = "spearman",
  pch = 21,
  cex.cor = 1,
  gap = 0.5,
  scale = FALSE,
  density = TRUE,
  ellipses = FALSE,
  lm = FALSE,
  smooth = FALSE,
  main = "Spearman_Correlation",
  cex.main = 1.5,
  font.main = 2,
  hist.col = "#9ECAE1",
  cex.labels = 1.6,  # diagonal variable names
  font.labels = 2,
  cex.axis = 1.15,
  stars = TRUE
)
```

```{r spearman-export}
# Spearman correlation matrix for the same variable set...
correlation_matrix <- cor(nips[, vars_spearman], method = "spearman")

# ...and the matching matrix of p-values. cor() alone does not return p-values,
# so this helper runs cor.test() on every variable pair and collects the results
# into a symmetric matrix (diagonal = 0 by convention).
cor.mtest <- function(mat, ...) {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = "spearman", ...)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  p.mat
}

p_matrix <- cor.mtest(nips[, vars_spearman])

decimal_places <- 3
rounded_correlation_matrix <- round(correlation_matrix, decimal_places)
rounded_p_matrix <- round(p_matrix, decimal_places)
rounded_correlation_matrix
rounded_p_matrix

# Side effect: export both matrices as CSV files next to this document
# (used for the papers' supplementary tables).
write.csv(rounded_correlation_matrix, "spearman_correlation_matrix.csv")
write.csv(rounded_p_matrix, "spearman_p_matrix.csv")
```

::: {.interpretation}
**Interpretation.**

On ranks, the same structure appears — Density–Modulus ρ = 0.93,
Modulus–`A0` ρ = 0.84 — plus one relation the Pearson view understates: Recovery
correlates positively with `T` (ρ = 0.48) even though it is unrelated to Modulus
(ρ = −0.09). Slower-relaxing sponges recover more of their height after prolonged
compression, while stiffness buys no resilience at all; the Polymer Journal paper builds
on exactly this contrast (its Figure S6 uses the subset Modulus, `RS`, `T`, Recovery,
within which the strongest link is Recovery–`RS`, ρ = 0.82 — `RS` is not part of the
panel printed above). The exported CSV files contain this matrix together with the
matching *P*-values.
:::

# Multivariate analysis

The responses are interrelated, so this part analyzes them jointly: principal component
analysis condenses the eight mechanical variables into a few uncorrelated components,
multi-response regression and MANOVA test the factor effects on several responses at
once, and the component scores themselves are regressed on the factors.

## Principal component analysis (PCA)

Principal component analysis of the **mechanical/stress-relaxation variables** (modulus plus
the exponential-fit parameters). PCA re-expresses the correlated variables as a small number
of uncorrelated components, revealing which properties vary together and how the samples
separate by processing condition. Variables are standardized internally by `PCA()`.

```{r pca-fit}
# Variables entering the PCA (8 variables → at most 8 components).
vars_pca <- c("Modulus", "A0", "A1", "t1", "A2", "t2", "A3", "t3")
res.pca <- PCA(nips[, vars_pca], graph = FALSE, ncp = 8)

# Eigenvalues: variance explained by each component.
eig <- as.data.frame(res.pca[["eig"]])
eig

# Variable coordinates (loadings) on the components.
coord <- as.data.frame(res.pca[["var"]][["coord"]])
coord

# Contribution (%) of each variable to each component.
contrib <- as.data.frame(res.pca[["var"]][["contrib"]])
contrib

# Squared cosines: quality of representation of each variable on each component.
cos2 <- as.data.frame(res.pca[["var"]][["cos2"]])
cos2

# Individual (sample) coordinates, i.e. the PC scores — reused later for
# regression on principal components.
coord_ind <- as.data.frame(res.pca[["ind"]][["coord"]])
coord_ind
```

::: {.interpretation}
**Interpretation.**

The mechanical variables are highly redundant: the first component
alone explains 65.8% of their joint variance and the first two explain 89.2%, so eight
measured quantities effectively carry about two independent pieces of information. The
near-zero last components (below 0.1%) are the flip side of the same redundancy, and part
of it is built in rather than discovered — see the PC7/PC8 diagnostic below. The
variable coordinates show what the two big components mean: all five stress quantities
point the same way on Dim.1 (Modulus 0.94, `A1` 0.96, `A2` 0.93, `A3` 0.92, `A0` 0.75),
while the relaxation times split between a negative share of Dim.1 and the bulk of Dim.2
(0.61–0.75) — in short, a "how strong" axis and a "how slow" axis.
:::

### Scree plot and variable map

The scree plot shows how much variance each component explains — components before the
"elbow" summarize most of the structure. The variable correlation plot (loading map) places
each variable by its correlation with PC1 and PC2; variables pointing the same way are
positively correlated, and color encodes representation quality (cos²). The contribution
bar charts identify which variables define each of the first three components (the dashed
line marks the contribution expected if all variables contributed equally).

```{r pca-plot-objects}
# Build the PCA summary plots; they are arranged into panels in the next two chunks.
a1 <- fviz_eig(res.pca, addlabels = TRUE, barfill = col_bar, barcolor = col_bar, ylim = c(1, 90),
               title = "Scree plot", ylab = "Explained variances (%)") +
  theme_pub + theme(aspect.ratio = 1.7)

b1 <- fviz_pca_var(res.pca,
                   title = "Variables \ncorrelation plot",
                   col.var = "cos2",
                   gradient.cols = c("#C6DBEF", "#2171B5", "#08306B"),
                   repel = TRUE,
                   arr.width = 25,) +
  theme_pub + theme(aspect.ratio = 1)

# Shared theme for the three per-dimension contribution charts.
theme_contrib <- theme_pub + theme(aspect.ratio = 1.8)
c1 <- fviz_contrib(res.pca, choice = "var", fill = col_bar, color = col_bar, axes = 1, title = "Contribution \nDim1") + theme_contrib
d1 <- fviz_contrib(res.pca, choice = "var", fill = col_bar, color = col_bar, axes = 2, title = "Contribution \nDim2") + theme_contrib
e1 <- fviz_contrib(res.pca, choice = "var", fill = col_bar, color = col_bar, axes = 3, title = "Contribution \nDim3") + theme_contrib
```

```{r pca-scree-loadings}
plot_grid(a1, b1, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15,
          rel_widths = c(1, 1), rel_heights = c(1, 1))
```

```{r pca-contributions}
plot_grid(c1, d1, e1, ncol = 3, nrow = 1, labels = c("A", "B", "C"), label_size = 15,
          rel_widths = c(1, 1), rel_heights = c(1, 1))
```

::: {.interpretation}
**Interpretation.**

The scree plot's elbow after the second component confirms that two
dimensions suffice. The contribution bars make the axis meanings precise: Dim.1 is built
almost evenly from the five stress quantities (about 11–18% each), Dim.2 is dominated by
the three relaxation times (`t2` 30%, `t1` 26%, `t3` 20%), and Dim.3 — already minor at
6.8% — is essentially the equilibrium stress `A0` (49%), the one stress parameter that
carries some information independent of the others.
:::

### Biplots

Each biplot overlays the samples (points, scored on the components) with the variable
loadings (arrows). Points close together are samples with similar mechanical profiles;
a sample lying in the direction of an arrow has a high value of that variable. The same
plot is drawn twice — colored by `SUS` (panel A) and by `EtOH` (panel B) — with 95%
concentration ellipses per group, to show which factor drives the separation. (These are
*data* ellipses — regions expected to contain about 95% of each group's samples under a
normal model — not confidence regions for the group means, which would be far tighter;
overlapping data ellipses are therefore compatible with clearly separated means.)

```{r pca-theme-defs}
# Shared themes for the biplot figures below (values identical across panels).
theme_biplot     <- theme_pub + theme(aspect.ratio = .6, legend.position = "bottom")
theme_biplot_lab <- theme_pub + theme(aspect.ratio = .65)
```

#### Biplot: PC1 vs PC2

```{r pca-biplot-12}
a2 <- fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(1, 2),
  geom.ind = "point",
  pointshape = 21,
  pointsize = 4,
  fill.ind = nips$SUS,          # colour samples by suspension concentration
  col.ind = "black",
  addEllipses = TRUE,          # 95% concentration ellipse per group
  col.var = "black",
  arr.width = 25,
  legend.title = list(fill = "SUS\n (wt%)"),
  mean.point = FALSE,
  repel = TRUE,
  ) +
  guides(fill = guide_legend(order = 1), color = "none") +
  theme_biplot + scale_sus + scale_sus_line

b2 <- fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(1, 2),
  geom.ind = "point",
  pointshape = 21,
  pointsize = 4,
  fill.ind = nips$EtOH,         # same plot, coloured by ethanol concentration
  col.ind = "black",
  addEllipses = TRUE,
  col.var = "black",
  arr.width = 25,
  legend.title = list(fill = "EtOH\n (wt%)"),
  mean.point = FALSE,
  repel = TRUE,
  ) +
  guides(fill = guide_legend(order = 1), color = "none") +
  theme_biplot + scale_etoh + scale_etoh_line

plot_grid(a2, b2, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

::: {.interpretation}
**Interpretation.**

The samples order themselves along Dim.1 by suspension concentration
(panel A), with the 0.24 wt% group on the low-stress side and the 0.4 wt% group on the
high-stress side, whereas the EtOH groups (panel B) overlap substantially. The overlap is
a matter of effect size rather than direction: both factors shift samples along the same
Dim.1 axis, but the SUS group means span about 4.8 score units against roughly 1.8 for
EtOH, so the ethanol ellipses sit inside one another instead of separating. Neither factor
moves samples much along Dim.2. The score regressions below put numbers on this.
:::

#### Biplot: PC1 vs PC3

```{r pca-biplot-13}
a3 <- fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(1, 3),
  geom.ind = "point",
  pointshape = 21,
  pointsize = 4,
  fill.ind = nips$SUS,
  col.ind = "black",
  addEllipses = TRUE,
  col.var = "black",
  arr.width = 25,
  legend.title = list(fill = "SUS\n (wt%)"),
  mean.point = FALSE,
  repel = TRUE,
  ) +
  guides(fill = guide_legend(order = 1), color = "none") +
  theme_biplot + scale_sus + scale_sus_line

b3 <- fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(1, 3),
  geom.ind = "point",
  pointshape = 21,
  pointsize = 4,
  fill.ind = nips$EtOH,
  col.ind = "black",
  addEllipses = TRUE,
  col.var = "black",
  arr.width = 25,
  legend.title = list(fill = "EtOH\n (wt%)"),
  mean.point = FALSE,
  repel = TRUE,
  ) +
  guides(fill = guide_legend(order = 1), color = "none") +
  theme_biplot + scale_etoh + scale_etoh_line

plot_grid(a3, b3, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

#### Biplot: PC2 vs PC3

```{r pca-biplot-23}
a4 <- fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(2, 3),
  geom.ind = "point",
  pointshape = 21,
  pointsize = 4,
  fill.ind = nips$SUS,
  col.ind = "black",
  addEllipses = TRUE,
  col.var = "black",
  arr.width = 25,
  legend.title = list(fill = "SUS\n (wt%)"),
  mean.point = FALSE,
  repel = TRUE,
  ) +
  guides(fill = guide_legend(order = 1), color = "none") +
  theme_biplot + scale_sus + scale_sus_line

b4 <- fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(2, 3),
  geom.ind = "point",
  pointshape = 21,
  pointsize = 4,
  fill.ind = nips$EtOH,
  col.ind = "black",
  addEllipses = TRUE,
  col.var = "black",
  arr.width = 25,
  legend.title = list(fill = "EtOH\n (wt%)"),
  mean.point = FALSE,
  repel = TRUE,
  ) +
  guides(fill = guide_legend(order = 1), color = "none") +
  theme_biplot + scale_etoh + scale_etoh_line

plot_grid(a4, b4, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

#### Biplot: PC1 vs PC2 with sample labels

Same projection as above, but each point is labeled with its sample ID (the `ID` column,
kept as row names on import) — useful for identifying individual samples and outliers.

```{r pca-biplot-12-labels-sus}
fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(1, 2),
  geom.ind = c("point", "text"),   # points plus sample-ID labels
  pointshape = 21,
  pointsize = 4,
  fill.ind = nips$SUS,
  col.ind = "black",
  addEllipses = FALSE,
  col.var = "black",
  arr.width = 25,
  legend.title = list(fill = "SUS (wt%)"),
  mean.point = FALSE,
  repel = TRUE,
  ) +
  guides(fill = guide_legend(order = 1), color = "none") +
  theme_biplot_lab + scale_sus
```

```{r pca-biplot-12-labels-etoh}
fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(1, 2),
  geom.ind = c("point", "text"),
  pointshape = 21,
  pointsize = 4,
  fill.ind = nips$EtOH,
  col.ind = "black",
  addEllipses = FALSE,
  col.var = "black",
  arr.width = 25,
  legend.title = list(fill = "EtOH (wt%)"),
  mean.point = FALSE,
  repel = TRUE,
  ) +
  guides(fill = guide_legend(order = 1), color = "none") +
  theme_biplot_lab + scale_etoh
```

#### Biplot: PC7 vs PC8

Diagnostic view of the two **last** components. These carry almost no variance; a
near-zero component corresponds to an almost exact linear dependency among the variables,
so this plot mainly serves as a check on redundancy/collinearity in the variable set. Here
it finds something concrete: the final component is dominated by Modulus and the stress
amplitudes, and indeed Modulus is very nearly a fixed multiple of `A0` + `A1` + `A2` + `A3`
(r = 0.996). That is expected from how the quantities are defined — the equilibrium and branch
stresses together make up the total stress carried at the start of relaxation,
σ(0) = σ~e~ + σ~1~ + σ~2~ + σ~3~ — but it means part of the
redundancy reported by the PCA is structural rather than an empirical finding about the
sponges.

```{r pca-biplot-78}
fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(7, 8),
  geom.ind = c("point", "text"),
  pointshape = 21,
  pointsize = 4,
  fill.ind = nips$SUS,
  col.ind = "black",
  addEllipses = FALSE,
  col.var = "black",
  arr.width = 25,
  legend.title = list(fill = "SUS (wt%)"),
  mean.point = FALSE,
  repel = TRUE,
  ) +
  guides(fill = guide_legend(order = 1), color = "none") +
  theme_biplot_lab + scale_sus
```

## Multi-response (multivariate) regression & ANOVA

Here several responses are fitted **jointly** against the same predictors with
`lm(cbind(...) ~ ...)`. The coefficients equal those of the separate per-response models,
but `anova()` on a multi-response fit performs multivariate (MANOVA-type) tests, taking the
correlations between the responses into account.

### Four responses jointly: additive model

Model: `lm(cbind(Modulus, A0, T, Recovery) ~ SUS + EtOH)`

```{r mlm-additive}
mod2.1 <- lm(cbind(nips$Modulus, nips$A0, nips$T, nips$Recovery) ~ SUS + EtOH, data = nips)
mod2.1 %>% summary()
mod2.1 %>% anova()
# tidy() labels each row with the response as written in the cbind() call
# ("nips$Modulus"); the "$" pair would be read as math delimiters in the knitted
# table, so the data-frame prefix is stripped for display only.
mod2.1 %>% tidy() %>% mutate(response = sub("^[^$]*\\$", "", response)) %>% kable(digits = 3)
mod2.1 %>% anova %>% kable(digits = 3)
```

### Four responses jointly: with interaction

Model: `lm(cbind(Modulus, A0, T, Recovery) ~ SUS * EtOH)`

```{r mlm-interaction}
mod2.1a <- lm(cbind(nips$Modulus, nips$A0, nips$T, nips$Recovery) ~ SUS + EtOH + SUS*EtOH, data = nips)
mod2.1a %>% summary()
mod2.1a %>% anova()
mod2.1a %>% tidy() %>% mutate(response = sub("^[^$]*\\$", "", response)) %>% kable(digits = 3)
mod2.1a %>% anova %>% kable(digits = 3)
```

::: {.interpretation}
**Interpretation.**

The coefficient estimates are identical to those of the separate
per-response fits — fitting responses jointly does not change least squares — but the
`anova()` table now tests each factor's effect on all four responses *at once*, taking
their correlations into account. Both factors are clearly significant in the joint test
(additive model: SUS Pillai ≈ 1.00, *P* ≈ 10⁻⁷; EtOH Pillai ≈ 0.79, *P* ≈ 3 × 10⁻⁵). The
value of the multivariate formulation is protection against cherry-picking: one overall
test per factor instead of four separate ones.
:::

## Regression & ANOVA on principal-component scores

Instead of modeling each variable separately, the first PCA scores are used as *composite*
responses: Dim.1 and Dim.2 summarize the joint variation of all eight mechanical variables,
so factor effects on the scores indicate effects on the overall mechanical profile.

### Both component scores jointly: with interaction

Model: `lm(cbind(Dim.1, Dim.2) ~ SUS * EtOH)`

```{r pca-scores-mlm}
# Sample scores on the principal components, from the PCA fitted above.
score <- data.frame(res.pca$ind$coord)

mod3.1 <- lm(cbind(score$Dim.1, score$Dim.2) ~ SUS + EtOH + SUS*EtOH, data = nips)
mod3.1 %>% summary()
mod3.1 %>% anova()
mod3.1 %>% tidy() %>% mutate(response = sub("^[^$]*\\$", "", response)) %>% kable(digits = 3)
mod3.1 %>% anova %>% kable(digits = 3)
```

### The stiffness axis (Dim.1): additive model

Model: `lm(Dim.1 ~ SUS + EtOH)`

```{r pca-scores-dim1}
mod3.2 <- lm(score$Dim.1 ~ SUS + EtOH, data = nips)
mod3.2 %>% summary()
mod3.2 %>% anova()
mod3.2 %>% tidy() %>% kable(digits = 3)
mod3.2 %>% anova %>% kable(digits = 3)
```

```{r pca-scores-dim1-diag}
par(mfrow = c(2, 2))
mod3.2 %>% plot()
```

### The timescale axis (Dim.2): additive model

Model: `lm(Dim.2 ~ SUS + EtOH)`

The same additive model applied to the second component, for comparison with Dim.1 above.

```{r pca-scores-dim2}
mod3.2b <- lm(score$Dim.2 ~ SUS + EtOH, data = nips)
mod3.2b %>% summary()
mod3.2b %>% anova()
mod3.2b %>% tidy() %>% kable(digits = 3)
mod3.2b %>% anova %>% kable(digits = 3)
```

### The stiffness axis (Dim.1): with interaction

Model: `lm(Dim.1 ~ SUS * EtOH)`

```{r pca-scores-dim1-int}
mod3.3 <- lm(score$Dim.1 ~ SUS + EtOH + SUS*EtOH, data = nips)
mod3.3 %>% summary()
mod3.3 %>% anova()
mod3.3 %>% tidy() %>% kable(digits = 3)
mod3.3 %>% anova %>% kable(digits = 3)
```

```{r pca-scores-dim1-int-diag}
par(mfrow = c(2, 2))
mod3.3 %>% plot()
```

::: {.interpretation}
**Interpretation.**

Treating the component scores as responses compresses eight ANOVAs
into two. Dim.1 — the overall stress level — is strongly determined by the fabrication
factors (*R*² = 0.86): suspension concentration is the dominant term (*P* < 0.001) and ethanol
a clear secondary one (*P* < 0.001), acting on the same axis with a smaller effect. Dim.2 —
the relaxation-time axis — is much harder to steer: a weak SUS effect (*P* = 0.036) and no
detectable ethanol effect (*P* = 0.407). So the factors move sponges a long way along the
stiffness axis and only slightly along the timescale axis, and the levers are not
cleanly separated: both act mainly on the same first component.
:::

## MANOVA

Multivariate analysis of variance tests whether the mean vector of the eight mechanical
variables (`Modulus`, `A0`, `A1`, `t1`, `A2`, `t2`, `A3`, `t3`) differs between factor
levels — one joint test instead of eight separate ANOVAs. Four standard test statistics
are reported: **Pillai's trace** (most robust to assumption violations), **Wilks' lambda**
(the classical likelihood-ratio criterion), **Hotelling–Lawley trace**, and **Roy's largest
root** (most powerful when group differences lie along a single dimension, but most
liberal). Agreement across all four strengthens the conclusion. The `summary.aov()` output
then breaks the joint result down into the univariate follow-up ANOVA for each response.

### SUS

`manova(cbind(Modulus, A0, A1, t1, A2, t2, A3, t3) ~ SUS)` — joint test of the SUS effect:

```{r manova-sus}
mod4.1 <- manova(cbind(nips$Modulus, nips$A0, nips$A1, nips$t1, nips$A2, nips$t2, nips$A3, nips$t3) ~ SUS, data = nips)
mod4.1 %>% summary.manova(intercept = TRUE, "Pillai")
mod4.1 %>% summary.manova(intercept = TRUE, "Wilks")
mod4.1 %>% summary.manova(intercept = TRUE, "Hotelling-Lawley")
mod4.1 %>% summary.manova(intercept = TRUE, "Roy")
```

Univariate follow-up ANOVAs (one per response):

```{r manova-sus-aov}
mod4.1 %>% summary.aov()
mod4.1 %>% tidy() %>% kable(digits = 3)
```

### EtOH

`manova(cbind(Modulus, A0, A1, t1, A2, t2, A3, t3) ~ EtOH)` — joint test of the EtOH effect:

```{r manova-etoh}
mod5.1 <- manova(cbind(nips$Modulus, nips$A0, nips$A1, nips$t1, nips$A2, nips$t2, nips$A3, nips$t3) ~ EtOH, data = nips)
mod5.1 %>% summary.manova(intercept = TRUE, "Pillai")
mod5.1 %>% summary.manova(intercept = TRUE, "Wilks")
mod5.1 %>% summary.manova(intercept = TRUE, "Hotelling-Lawley")
mod5.1 %>% summary.manova(intercept = TRUE, "Roy")
```

Univariate follow-up ANOVAs (one per response):

```{r manova-etoh-aov}
mod5.1 %>% summary.aov()
mod5.1 %>% tidy() %>% kable(digits = 3)
```

::: {.interpretation}
**Interpretation.**

Both factors significantly shift the joint mean vector of the eight
mechanical variables: SUS with Pillai = 1.14 (*P* ≈ 10⁻⁶) and EtOH with Pillai = 0.68
(*P* = 0.027). All four test statistics lead to the same conclusion for each factor, which
is precisely the robustness check they are printed for. The contrast in strength mirrors
everything above — suspension concentration is the primary mechanical lever, ethanol the
secondary one — and the univariate follow-ups identify which individual variables carry
each joint effect. One caveat on how these two tests are built: each is a separate
one-way MANOVA, so neither factor is adjusted for the other. The large SUS effect
therefore stays inside the residual variation of the EtOH model, which makes the ethanol
test conservative. The two-factor multi-response model above improves on this only
partly — `anova()` on a multivariate fit is also sequential, so there EtOH is adjusted for
SUS but SUS is not adjusted for EtOH. Both routes reach the same conclusion here, and for
a fully order-free multivariate test `car::Anova(mod2.1, type = 2)` is the tool to reach
for.
:::

## ASCA: partitioning the multivariate variation

ANOVA–simultaneous component analysis (ASCA) ties the two threads of this part together.
Like the MANOVA, it asks how much of the multivariate variation each design term is
responsible for; like the PCA, it then looks *inside* each term's share with a component
decomposition. The autoscaled data matrix is split into additive effect matrices built
from the level means — one for SUS, one for EtOH, one for their interaction — plus a
residual matrix; each effect matrix gets a sum of squares (its share of the total
variation) and its own PCA. Significance is assessed by permutation: the condition labels
are reshuffled across sponges (jointly, so the cell sizes stay intact) and the observed
sums of squares are compared with the reshuffled distribution. The implementation is
written out in full below rather than imported from a package, so every step can be
followed and modified.

```{r asca-analysis}
set.seed(1234)
X <- scale(as.matrix(nips[, vars_pca]))   # autoscaled mechanical variables (40 x 8)

# Effect matrices from level means (the grand mean is 0 after autoscaling).
asca_decomp <- function(a, b) {
  Ma <- apply(X, 2, ave, a)
  Mb <- apply(X, 2, ave, b)
  Mi <- apply(X, 2, ave, interaction(a, b)) - Ma - Mb
  list(Ma = Ma, Mb = Mb, Mi = Mi,
       ssq = c(`SUS` = sum(Ma^2), `EtOH` = sum(Mb^2), `SUS x EtOH` = sum(Mi^2)))
}
obs <- asca_decomp(nips$SUS, nips$EtOH)
resid_ssq <- sum((X - obs$Ma - obs$Mb - obs$Mi)^2)

# Permutation test: reshuffle which sponge received which condition (the SUS/EtOH
# label pairs are permuted jointly, so the design's cell sizes are preserved).
n_perm <- 2000
perm <- replicate(n_perm, {
  i <- sample(nrow(X))
  asca_decomp(nips$SUS[i], nips$EtOH[i])$ssq
})
p_perm <- (rowSums(perm >= obs$ssq) + 1) / (n_perm + 1)

kable(data.frame(
  Effect = c(names(obs$ssq), "Residual"),
  `Sum of squares` = round(c(obs$ssq, resid_ssq), 1),
  `Share of total (%)` = round(100 * c(obs$ssq, resid_ssq) / sum(X^2), 1),
  `Permutation *P*` = c(ifelse(p_perm < 0.001, "< 0.001", sprintf("%.3f", p_perm)), ""),
  check.names = FALSE), row.names = FALSE)
```

```{r asca-effect-pca}
# PCA of each main-effect matrix: with three levels an effect matrix has rank two,
# and its first component tells us what kind of mechanical change the factor produces.
effect_pc1 <- function(M) {
  sv <- svd(M)
  list(share = sv$d[1]^2 / sum(sv$d^2), loading = setNames(sv$v[, 1], colnames(X)))
}
pc_sus <- effect_pc1(obs$Ma); pc_etoh <- effect_pc1(obs$Mb)
cat(sprintf("SUS-effect PC1 carries %.1f%% of the SUS effect; loadings:\n", 100 * pc_sus$share))
round(pc_sus$loading, 2)
cat(sprintf("EtOH-effect PC1 carries %.1f%% of the EtOH effect; loadings:\n", 100 * pc_etoh$share))
round(pc_etoh$loading, 2)
```

::: {.interpretation}
**Interpretation.**

The decomposition puts one number on what this whole part has been
showing: suspension concentration owns a little over half of the total variation in the
eight mechanical variables (≈ 56%, permutation *P* < 0.001), while ethanol's share (≈ 10%)
is of the same order as the interaction (≈ 8%) and both are dwarfed by the residual
(≈ 29%). Notice that ethanol does *not* clear this permutation test (*P* ≈ 0.09) even
though the MANOVA declared it significant. That is not a contradiction but a difference
in the question asked: the ASCA test asks whether ethanol's share of raw variation
exceeds chance, whereas the MANOVA weighs the effect against the residual covariance —
a sharper instrument when a large competing effect (here SUS) inflates the total. A
bookkeeping note: with unequal replicate counts the effect matrices are not exactly
orthogonal, so the shares add to slightly more than 100% (by about 2%) — the same
unbalance that motivated the Type II tests earlier.

The effect-level PCAs then compress each factor's action to essentially one direction —
PC1 carries 97–98% of either effect matrix, meaning the three level means lie almost
on a line in the eight-dimensional response space. The SUS direction loads all five
stress quantities together, with weak opposite loadings on the relaxation times: raising
the suspension concentration raises every stress parameter and mildly shortens the
timescales — the Dim.1 story of the global PCA restated as a design effect. The ethanol
direction is the more distinctive one: it pairs the stress amplitudes and the relaxation
times with *opposite* signs, so moving from 0.2 to 1 wt% ethanol slides sponges along a
single latent axis toward lower stress amplitudes *and* longer relaxation times. That is
ethanol's softer-but-slower signature — which the univariate models could only describe
one response at a time — captured as one direction in response space.
:::

# Response-surface analysis (continuous factors)

Everywhere above, `SUS` and `EtOH` are treated as categorical design levels. This part
instead treats them as **continuous** variables and uses response surface methodology
(RSM) to model `Shrinkage` (and, at the end, `Density`) as a smooth function of both
factors, working on the numeric data copy `dat`. This view supports interpolation between
the tested levels and indicates in which direction the response decreases fastest.

The first-order Shrinkage model below is the analysis published in the ACS Omega paper:
its coefficients appear there as Table 2, its ANOVA (including the lack-of-fit test) as
Table S6, and its contour plot as Figure 6j. The RSM prediction was validated with three
additional sponges fabricated at EtOH = 3 wt% (one per SUS level; ACS Omega SI Table S1,
rows 43–45; provided here as `NIPS_validation.csv`):
the three sponges measured 49–56% shrinkage against 46–48% predicted by the plane, closest
at SUS 0.4 wt% (49.4% against 46.4%). The "Which model extrapolates?" section below examines
that comparison properly.
The second-order model and canonical analysis are exploratory extensions beyond the
published analysis, and the extrapolation test below shows why the plane, not the
curved surface, is the model to trust outside the design region.

## First-order (FO) model

`FO(SUS, EtOH)` fits a first-order response surface — a plane over the two factors. The
`summary()` includes a lack-of-fit test, which indicates whether this simple plane is
adequate or whether curvature (a second-order model) is needed. The test exploits the
replicates: sponges fabricated at identical conditions scatter around their own condition
mean, giving a model-free noise estimate (*pure error*), while the nine condition means
scatter around the fitted plane (*lack of fit*); the plane is called into question only
when the second kind of scatter clearly exceeds the first. The predictors enter in
their original units (uncoded), so coefficients are interpretable directly on the
concentration scales. The coefficient table equals ACS Omega Table 2, and the ANOVA block
of the summary equals Table S6 (model *P* = 0.011; lack of fit insignificant, *P* = 0.260).

```{r rsm-fo-fit}
rsm1 <- rsm(Shrinkage ~ FO(SUS, EtOH), data = dat)
summary(rsm1)
rsm1 %>% tidy() %>% kable(digits = 3)
```

::: {.interpretation}
**Interpretation.**

On the fitted plane, shrinkage falls by about 3.8 percentage points
per 1 wt% of ethanol (*P* = 0.005), while the SUS slope (−8.9 per wt%, i.e. only about
−1.4 points across the narrow 0.24–0.4 wt% range) is not statistically distinguishable
from zero (*P* = 0.176). The ANOVA block is as informative as the coefficients: the model
is significant (*P* = 0.011) while the lack-of-fit test is not (*P* = 0.260) — the replicate
data inside the design region give no evidence *against* a plane. Read that as a failure
to reject rather than proof of adequacy: with 6 and 31 degrees of freedom the test has
limited power, and the model accounts for only about 22% of the variance in shrinkage
(*R*² = 0.216), so a good deal of sample-to-sample variation remains unexplained. The test
also says nothing about behavior outside the tested range; the EtOH = 3 wt% sponges are
genuine extrapolation, and how they compare with this plane is assessed on its own terms
later in this part — one of the three lands close, the other two are 7.5–8.5 points off.
:::

### Contour plot of the fitted plane

Filled contour map of predicted Shrinkage. The red rectangle marks the experimental
(design) region actually covered by the study — predictions outside it are extrapolation
and should be read with caution.

```{r rsm-fo-contour}
par(cex.axis = 1.3, cex.lab = 1.45, font.lab = 2, font.axis = 2)
contour(rsm1, ~ SUS + EtOH, image = TRUE, img.col = pal_surface, labcex = 1, xlabs = c("EtOH (wt%)", "SUS (wt%)"), bounds = list(EtOH = c(0.1, 4), SUS = c(0, 0.5)),
        zlim = c(40, 62), decode = FALSE)
lines(c(0.4, 0.24, 0.24, 0.4, 0.4), c(0.2, 0.2, 1, 1, 0.2), col = col_accent, lwd = 2.5)  # design region
text(0.32, 1.18, "design region", col = col_accent, font = 2, cex = 1.05)
```

## Second-order (SO) model & canonical analysis

`SO(SUS, EtOH)` adds quadratic and interaction terms, allowing a curved surface. The
canonical analysis then locates the stationary point of that surface and reports the
eigenvalues of the quadratic form: all negative → maximum, all positive → minimum, mixed
signs → saddle. `threshold = 0` keeps the eigenvalues exactly as estimated (no
thresholding of near-zero values).

```{r rsm-so-fit}
rsm2 <- rsm(Shrinkage ~ SO(SUS, EtOH), data = dat)
canonical(rsm2, threshold = 0)
```

Eigenvalues are scale-dependent — they are second derivatives of the surface, in units
of the response per squared predictor unit — so on the raw wt% scales the two are not
comparable: SUS is varied over a range five times narrower than EtOH, which inflates its
apparent curvature by a factor of about 25. Refitting on **coded** factors — each scaled
so that the design range runs from −1 to +1 — removes that artifact and is the standard
way to read a canonical analysis.

```{r rsm-so-coded}
# Code both factors to [-1, +1] over their design ranges: SUS 0.24-0.4, EtOH 0.2-1 wt%.
dat_coded <- coded.data(dat, x1 ~ (SUS - 0.32)/0.08, x2 ~ (EtOH - 0.6)/0.4)
rsm2c <- rsm(Shrinkage ~ SO(x1, x2), data = dat_coded)
summary(rsm2c)
canonical(rsm2c, threshold = 0)
```

::: {.interpretation}
**Interpretation.**

Both eigenvalues are positive, so the fitted quadratic has a minimum
at the stationary point, located at roughly SUS 0.35 wt%, EtOH 0.90 wt% — that is, inside
the design region rather than at an optimum beyond it. The uncoded eigenvalues (≈ 247 and
≈ 8) look wildly unequal, but that ratio is almost entirely the unit mismatch just
described: on coded factors they become ≈ 1.7 and ≈ 1.2. This is a useful general lesson —
never compare the magnitudes of canonical eigenvalues computed from uncoded predictors,
because the factor varied over the narrower range will always appear to dominate.

The bigger caveat is that none of this is statistically resolved. The coded summary shows
that no quadratic or interaction term is significant (`x1^2` *P* = 0.12, `x2^2` *P* = 0.21,
`x1:x2` *P* = 0.49), so the shape being described — the sign of the eigenvalues, hence
whether this is a bowl or a saddle, and where the stationary point sits — rests on
coefficients that the data do not distinguish from zero. With only three levels per factor
a full quadratic sits at the edge of what the design can support, and neither the
first-order lack-of-fit test nor the quadratic terms themselves give any evidence that the
extra curvature was needed. Read the
canonical analysis here as an illustration of the method rather than a finding about these
sponges; the comparison below shows what happens when it is taken literally.
:::

### Contour plot with the canonical path

The blue circles trace the canonical path returned by `canonical.path()`, walking outwards
from the stationary point (the distinct symbol at distance 0) along the axis of *least*
curvature. The function is called with `descent = TRUE`, but on this surface nothing
actually descends: because both eigenvalues are positive the stationary point is a
minimum, so predicted shrinkage rises in every direction away from it — gently along this
path and steeply elsewhere. The filled blue square marks the validation condition
SUS = 0.4 wt%, EtOH = 3 wt%, whose measured shrinkage was 49.4%; note that it sits far
off this fitted surface, which is the point made below. The red rectangle is again the
design region.

```{r rsm-so-contour-path}
par(cex.axis = 1.3, cex.lab = 1.45, font.lab = 2, font.axis = 2)
contour(rsm2, EtOH ~ SUS, image = TRUE, img.col = pal_surface, labcex = 1, xlabs = c("EtOH (wt%)", "SUS (wt%)"), bounds = list(EtOH = c(0.1, 4), SUS = c(0, 0.5)),
        zlim = c(30, 80), decode = FALSE)
lines(c(0.4, 0.24, 0.24, 0.4, 0.4), c(0.2, 0.2, 1, 1, 0.2), col = col_accent, lwd = 2.5)  # design region
text(0.32, 1.18, "design region", col = col_accent, font = 2, cex = 1.05)
points(EtOH ~ SUS, data = canonical.path(rsm2, descent = TRUE, threshold = 0),
       col = "#08306B", pch = 1 + 6*(dist == 0))   # pch 7 marks the stationary point (dist == 0)
points(0.4, 3, col = "black", pch = 22, cex = 2, bg = "#2171B5")  # validation condition (SUS 0.4, EtOH 3 wt%; measured shrinkage 49.4%)
text(0.4, 3.45, "validation sponge", col = "black", font = 2, cex = 1.0)
```

### Contour plot centered at the stationary point

The same second-order surface, re-drawn with the slice taken through the stationary point
(`at = summary(rsm2)$canonical$xs`) and a wider EtOH range.

```{r rsm-so-contour-stationary}
par(cex.axis = 1.3, cex.lab = 1.45, font.lab = 2, font.axis = 2)
contour(rsm2, EtOH ~ SUS, image = TRUE, img.col = pal_surface, labcex = 1, xlabs = c("EtOH (wt%)", "SUS (wt%)"), at = summary(rsm2)$canonical$xs,
        bounds = list(EtOH = c(0.1, 5), SUS = c(0, 0.5)), zlim = c(40, 100), decode = FALSE)
```

### Which model extrapolates? A test against the validation sponges

The three EtOH = 3 wt% sponges were fabricated well beyond the design region, which makes
them an honest out-of-sample test of the two competing surfaces.

```{r rsm-validation-check}
val <- read.csv("NIPS_validation.csv", header = TRUE, row.names = 1)

comparison <- data.frame(
  SUS = val$SUS, EtOH = val$EtOH,
  `Measured (%)` = round(val$Shrinkage, 1),
  `First order (%)` = round(predict(rsm1, val[, c("SUS", "EtOH")]), 1),
  `Second order (%)` = round(predict(rsm2, val[, c("SUS", "EtOH")]), 1),
  check.names = FALSE)
kable(comparison, row.names = FALSE)
```

::: {.interpretation}
**Interpretation.**

Read this as a comparison between the two models, not as a triumph for
either. The plane predicts 46–48% shrinkage against 49–56% measured, so it errs by 3.0,
7.5 and 8.5 percentage points — up to about three times the model's own residual standard
error (2.74), and larger than the entire 4.5-point swing it predicts across the design
grid. It also gets the trend only partly right: from EtOH 1 wt% to EtOH 3 wt% the measured
shrinkage fell at SUS 0.4 wt% (54.0 → 49.4%) and at SUS 0.24 wt% (56.9 → 56.3%), but rose
at SUS 0.3 wt% (53.1 → 54.8%), whereas the plane predicts a uniform fall of 7.6 points
(twice its EtOH slope of −3.78 per wt%). The honest summary is that the plane stays in the
right range and captures the direction at the highest suspension concentration, which is
where the paper's optimization argument was aimed.

The quadratic, by contrast, is not merely imprecise but qualitatively wrong: it predicts
about 88–95% shrinkage where 49–56% was measured. The reason is structural rather than
accidental. A fitted quadratic with a minimum inside the design region must curve upwards
in every direction outside it, so extrapolating it towards higher ethanol — the very
direction the experiment wanted to explore — is guaranteed to overshoot. That is the
general lesson worth taking from this table: polynomial surfaces describe the region they
were fitted to, and added flexibility inside that region buys nothing outside it. The
published analysis used the first-order model, and this comparison shows why.
:::

## Predicted means over the design grid

Estimated marginal means of Shrinkage from the first-order model, evaluated at the nine
design combinations of the study (SUS 0.24/0.3/0.4 wt% × EtOH 0.2/0.5/1 wt%;
`mode = "decoded"` reports them on the original concentration scales). On an additive
model these estimated marginal means are simply the fitted plane evaluated at the grid
points; they differ slightly from the raw cell means because the model pools information
across all 40 sponges, and unlike raw means they come with model-based standard errors.

```{r rsm-emmeans-grid}
# Note: emmeans reconstructs the data from the model call, so `dat` must exist in the
# workspace — which is the case when this document is knitted normally.
emmeans::emmeans(rsm1, ~ SUS * EtOH, mode = "decoded",
                 at = list(SUS = c(0.24, 0.3, 0.4), EtOH = c(0.2, 0.5, 1)))
```

::: {.interpretation}
**Interpretation.**

The estimated marginal means put numbers on the design grid:
predicted shrinkage spans from 58.4% at the gentlest condition (SUS 0.24, EtOH 0.2) down
to 53.9% at (SUS 0.4, EtOH 1) — a swing of about 4.5 percentage points, with standard
errors below one point. The monotone decrease along both axes is what motivated the
validation runs beyond the design region at EtOH = 3 wt% (`NIPS_validation.csv`), assessed above in
the "Which model extrapolates?" section.
:::

## Cross-checks with ordinary lm()

The same first-order surface can be fitted with plain `lm()`; these fits reproduce the
RSM results and add the standard regression diagnostics.

### First-degree polynomial fit

`poly(SUS, EtOH, degree = 1)` spans the same model space as `FO(SUS, EtOH)`, so this
contour should match the FO contour above. (The point SUS = 0.4, EtOH = 3 is the same
validation condition as before.)

```{r rsm-fo-poly-check}
fo_poly <- lm(Shrinkage ~ poly(SUS, EtOH, degree = 1), data = dat)
par(cex.axis = 1.3, cex.lab = 1.45, font.lab = 2, font.axis = 2)
contour(fo_poly, ~ SUS + EtOH, image = TRUE, img.col = pal_surface, labcex = 1, xlabs = c("EtOH (wt%)", "SUS (wt%)"), bounds = list(EtOH = c(0.1, 4), SUS = c(0, 0.5)),
        zlim = c(40, 62), decode = FALSE)
lines(c(0.4, 0.24, 0.24, 0.4, 0.4), c(0.2, 0.2, 1, 1, 0.2), col = col_accent, lwd = 2.5)  # design region
text(0.32, 1.18, "design region", col = col_accent, font = 2, cex = 1.05)
points(0.4, 3, col = "black", pch = 22, cex = 2, bg = "#2171B5")  # validation condition (SUS 0.4, EtOH 3 wt%; measured shrinkage 49.4%)
text(0.4, 3.45, "validation sponge", col = "black", font = 2, cex = 1.0)
```

### Additive linear model for Shrinkage, with diagnostics

Plain additive regression `Shrinkage ~ SUS + EtOH` on the numeric factors, its ANOVA
table, a contour zoomed to the design region, and the standard diagnostic plots
(residuals vs fitted, normal Q–Q, scale–location, residuals vs leverage).

```{r rsm-lm-shrinkage}
lm2 <- lm(Shrinkage ~ SUS + EtOH, data = dat)
summary(lm2)
anova(lm2)
par(cex.axis = 1.3, cex.lab = 1.45, font.lab = 2, font.axis = 2)
contour(lm2, ~ SUS + EtOH, image = TRUE, img.col = pal_surface, labcex = 1, xlabs = c("EtOH (wt%)", "SUS (wt%)"), bounds = list(EtOH = c(0, 1), SUS = c(0, 0.5)),
        zlim = c(50, 65), decode = FALSE)
lines(c(0.4, 0.24, 0.24, 0.4, 0.4), c(0.2, 0.2, 1, 1, 0.2), col = col_accent, lwd = 2.5)  # design region
text(0.32, 0.55, "design region", col = col_accent, font = 2, cex = 1.05)
par(mfrow = c(2, 2))
lm2 %>% plot()
```

### Additive linear model for Density

The same approach applied to Density as the response.

```{r rsm-lm-density}
lm3 <- lm(Density ~ SUS + EtOH, data = dat)
summary(lm3)
par(cex.axis = 1.3, cex.lab = 1.45, font.lab = 2, font.axis = 2)
contour(lm3, ~ SUS + EtOH, image = TRUE, img.col = pal_surface, labcex = 1, xlabs = c("EtOH (wt%)", "SUS (wt%)"), bounds = list(EtOH = c(0, 1), SUS = c(0, 0.5)), decode = FALSE)
lines(c(0.4, 0.24, 0.24, 0.4, 0.4), c(0.2, 0.2, 1, 1, 0.2), col = col_accent, lwd = 2.5)  # design region
text(0.32, 0.55, "design region", col = col_accent, font = 2, cex = 1.05)
```

# Structure–property scaling (Gibson–Ashby)

The apparent elastic modulus scales with density following the Gibson–Ashby model for
cellular solids, E\*/*E*~s~ = C · (ρ/ρ~s~)^n^, with the solid cellulose acetate reference
values *E*~s~ = 1178 MPa and ρ~s~ = 1311 kg·m⁻³. The model is fitted by nonlinear least
squares on the original scale, which reproduces the constants reported in Polymer Journal
Figure 5b (*C* = 2.04, *n* = 2.48). The exponent n is the informative parameter statistically:
values near 2 are the classical signature of bending-dominated deformation in open-cell
solids. The prefactor C measures how stiff the network is at a given relative density
compared with that idealized reference; the Polymer Journal paper discusses what its value
implies about the cell architecture.

```{r gibson-ashby}
Erel   <- (nips$Modulus / 1000) / 1178   # kPa -> MPa, relative to solid CA (Es = 1178 MPa)
rhorel <- nips$Density / 1311            # relative to solid CA (rho_s = 1311 kg m^-3)

ga_fit <- nls(Erel ~ C * rhorel^n, start = list(C = 1, n = 2))
summary(ga_fit)
cov2cor(vcov(ga_fit))   # C and n are almost perfectly correlated (see the discussion)

par(cex.axis = 1.25, cex.lab = 1.4, font.lab = 2, font.axis = 2, cex.main = 1.5)
plot(rhorel, Erel, log = "xy",
     pch = 21, cex = 1.4, bg = pal_sus[as.integer(nips$SUS)],
     xlab = "Relative density", ylab = "Relative modulus",
     main = "Gibson-Ashby scaling")
curve(coef(ga_fit)[["C"]] * x^coef(ga_fit)[["n"]], add = TRUE, col = col_accent, lwd = 2)
legend("topleft", legend = levels(nips$SUS), pt.bg = pal_sus,
       pch = 21, title = "SUS (wt%)", bty = "n")
```

::: {.interpretation}
**Interpretation.**

The two constants are estimated with very different precision: the
exponent comes out at *n* = 2.48 ± 0.15 and exceeds the ideal open-cell value of 2, while
the prefactor carries a wide interval (*C* = 2.04 ± 1.56). It is worth understanding *why*
C is hard to pin down here, because the reason is geometric rather than a defect of the
data. The sponges span relative densities of only about 0.004–0.007, while C is defined
at ρ~rel~ = 1 — an extrapolation of more than two decades. Over such a lever arm the
intercept and exponent are almost perfectly entangled (the correlation matrix printed
above gives r(*C*, *n*) ≈ 0.9995): tilting the fitted line slightly within the data window
swings the extrapolated prefactor enormously — each 0.1 change in n moves C by a factor
of about 1.7. And the window itself cannot easily be widened: fabricating sponges over a
much broader density range *while preserving the same internal network configuration* is
exactly what a fixed-chemistry cryo-templating process makes difficult, so the narrow
density span is a physical constraint of the material system, not a shortcoming of the
experiment.

The same entanglement explains the difference between fitting methods. Nonlinear least
squares on the original scale gives (*C*, *n*) = (2.04, 2.48); ordinary linear regression of
log(*E*~rel~) on log(ρ~rel~) gives (7.5, 2.73). These are not competing physical claims:
within the measured window the two curves agree to within roughly 10% (up to \~12% at
the window's lower edge), and they diverge only when extrapolated toward ρ~rel~ = 1, where the log-scale fit's slightly steeper
slope compounds into an almost fourfold larger intercept (a factor of roughly
(1/0.005)^Δ*n*^ ≈ 3.7 for Δ*n* = 0.25, where 0.005 is the mid-window relative density). The two (*C*, *n*) pairs are two points on the same
narrow ridge of parameter combinations that describe these sponges equally well — the
log-scale value *C* = 7.5 therefore has no separate structural meaning to interpret, and
attempting to read it as one would double-count what is simply a reweighting of the same
information.

Which value should be reported and interpreted? In the porous-materials literature the
reading of C is empirical: values are compared against the accumulated body of measured
foams and lattices and against idealized unit-cell models, a tradition anchored in the
standard Gibson–Ashby form fitted directly — which is the convention the published value
*C* = 2.04 follows, and the appropriate one for comparison with prior reports. Read in that
tradition, and alongside the independently measured cell architecture (the X-ray CT
analysis in the Polymer Journal paper), the point estimate is consistent with the
stiffer-than-ideal, more highly connected network described there; the wide statistical
interval records that the accessible density window alone cannot pin the prefactor down,
not that the structural reading is wrong. Two practical habits follow for anyone reusing
this fit: always quote C together with its paired n and the fitting convention (along the
ridge, C is only meaningful at its own exponent), and when a method-independent stiffness
summary is needed, quote the fitted relative modulus at a density *inside* the measured
window — at ρ~rel~ = 0.006 both fits give *E*~rel~ ≈ 6.4 × 10⁻⁶ — which is estimated
precisely and does not depend on the extrapolation at all.
:::

# Concluding remarks

Taken together, the analyses tell one coherent story with two levers. The suspension
concentration sets how much material the cryo-templated network contains: density,
porosity, stiffness and the Maxwell stress amplitudes all follow it almost
deterministically (*R*² up to 0.96), and the Gibson–Ashby fit condenses this into a single
power law whose exponent (*n* = 2.48) sits above the value of 2 expected for ideal
open-cell bending. The ethanol concentration acts earlier in the process, on
ice-crystal formation, and therefore shows up in the shape- and time-related quantities:
shrinkage decreases while diameter, recovery, residual stress and the characteristic
relaxation time all increase. The multivariate view sharpens how the two levers differ. They are not
orthogonal: both act mainly on the same "strength" component of the mechanical profile,
with suspension concentration much the stronger of the two. What distinguishes them is
that ethanol additionally raises recovery and residual stress, and it does so while
*lowering* stiffness — so the two properties trade off along the ethanol axis. Suspension
concentration changes stiffness with no *detectable* effect on recovery (*P* = 0.93), which
is what makes the two separately addressable here. That last statement deserves its own
caveat, of the kind this document has applied throughout: with 40 sponges the data bound
any suspension effect on recovery to at most about 7 percentage points — an upper limit
almost as large as ethanol's own effect. The 95% interval for the largest suspension
contrast on recovery runs from −4.8 to +6.9 percentage points, against an estimated
ethanol effect of +7.4. "No effect detected" is not "no effect", and a larger study could well
find one.
For the mechanistic interpretation of these patterns (cell-wall thickening,
ice-templating control, nanofilm delamination and slippage), see the two papers cited at
the top; this document is meant to let readers retrace, probe and extend the statistical
side of that story.
