---
title: "NIPS-JS Sponge — Data & Statistical Analysis"
author: "De Nguyen"
output:
  html_document:
    toc: yes
    number_sections: yes
    df_print: default
    fig_width: 10
    fig_height: 6
  word_document:
    toc: true
    fig_width: 10
    fig_height: 6
    keep_md: true
  pdf_document:
    toc: yes
editor_options:
  markdown:
    wrap: 120
---

```{r setup, include=FALSE}
# Global chunk options: code is hidden in the rendered report (echo = FALSE), so the
# knitted document shows only figures and tables. Read this .Rmd file for the code.
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.retina = 5)

library(psych)       # pairs.panels(): scatter-plot matrices with correlations
library(factoextra)  # fviz_*(): publication-ready PCA graphics
library(FactoMineR)  # PCA(): principal component analysis
library(ggplot2)     # grammar-of-graphics plotting
library(dplyr)       # %>% pipe used throughout
library(cowplot)     # plot_grid(): combining plots into labelled panels
library(broom)       # tidy(): turn model summaries into data frames
library(knitr)       # kable(): formatted tables in the rendered document
library(MASS)        # boxcox(): Box–Cox transformation profiles
library(rsm)         # response-surface models: FO()/SO(), canonical analysis, contour.lm()
# emmeans is called via emmeans::emmeans() without attaching the package.
```

# About this document

This notebook contains the complete dataset documentation and statistical analysis behind
the NIPS-JS sponge studies (it merges the repository's former `NIPS.Rmd` and `RSM.Rmd`
into one systematic workflow):

1. *Nonsolvent-Induced Phase Separation–Jet Spinning: An Innovative Technique for Producing
   Cellulosic Nanofilms, Suspensions, and Nanofilm-Based Sponges* (2025), ACS Omega,
   <https://doi.org/10.1021/acsomega.5c02353>
2. *Mechanically tunable nanofilm-based cellulose acetate sponges via crosslinker-free
   cryo-templating* (2025), Polymer Journal,
   <https://doi.org/10.1038/s41428-025-01083-z>

Forty sponges were fabricated in a 3 × 3 full factorial design over two experimental
factors — the polymer suspension concentration (`SUS`: 0.24, 0.3, 0.4 wt%) and the ethanol
cosolvent concentration (`EtOH`: 0.2, 0.5, 1 wt%), with 4–7 replicates per combination —
and characterised by a set of physical and mechanical response variables.

The document is organised as a pipeline from raw data to models:

1. **Study data** — import and overview of all responses.
2. **Distributional checks** — Shapiro–Wilk tests, normal Q–Q plots and Box–Cox profiles,
   run before any modelling.
3. **Factor effects I: distributions by factor level** — one-factor boxplots for each
   response and two-factor boxplots for selected responses.
4. **Factor effects II: regression & ANOVA** — a linear model per response with `SUS` and
   `EtOH` as categorical predictors, plus interaction models for selected responses.
5. **Correlations between responses** — Pearson and Spearman analyses.
6. **Multivariate analysis** — PCA with biplots, multi-response regression, regression on
   component scores, and MANOVA.
7. **Response-surface analysis** — `SUS` and `EtOH` as continuous predictors: first- and
   second-order surfaces for Shrinkage, canonical analysis, steepest descent, and
   cross-checks with ordinary regression.
8. **Structure–property scaling** — Gibson–Ashby power-law fit of relative modulus vs
   relative density.
9. **Appendix** — an unevaluated ASCA sketch kept for reference.

Two representations of the dataset are used: the factor version `nips` (the design levels
as discrete groups; used everywhere except the response-surface part) and the numeric
version `dat` (the response-surface models need continuous predictors) — see the import
section.

**Data.** The full dataset ships with this repository as `NIPS.csv` (40 sponges), licensed
CC BY 4.0 — see `DATA_LICENSE.md`. The geometry variables (Weight … Shrinkage) correspond
to Table S1 of the ACS Omega Supporting Information and the mechanical variables
(Recovery … RS) to Table S1 of the Polymer Journal Supplementary Information; both SI
tables list the same 40 sponges in the same row order. The derived geometry responses
(Volume, Density, Porosity, Shrinkage) are stored at full precision, recomputed from
Weight/Diameter/Height with the papers' equations — the SI tables print them rounded, and
the published statistics were computed on the unrounded values. Three additional sponges
used to validate the RSM prediction (EtOH = 3 wt%) are provided in `NIPS_validation.csv`.

**Column dictionary** (measurement protocols are described in the papers):

| Column | Role | Description |
|---|---|---------|
| `ID` | identifier | Sample number; becomes the row name on import. |
| `SUS` | design factor | Polymer suspension concentration (wt%; levels 0.24 / 0.3 / 0.4). |
| `EtOH` | design factor | Ethanol cosolvent concentration (wt%; levels 0.2 / 0.5 / 1). |
| `Weight` | response | Sponge weight (g); primary measurement (electronic balance). |
| `Diameter`, `Height` | response | Sponge dimensions (mm); primary measurements (caliper). |
| `Volume` | response | Bulk volume (cm³), derived as a cylinder: V = π/4 · h · d². |
| `Density` | response | Bulk density (kg·m⁻³), derived: ρ = m/V. |
| `Porosity` | response | Porosity (%), derived: 100 · (1 − ρ/ρ~CA~) with ρ~CA~ = 1311 kg·m⁻³. |
| `Shrinkage` | response | Volumetric shrinkage (%), derived: 100 · (1 − V/V~suspension~) with 15 mL of suspension. |
| `Recovery` | response | Percentage of the initial height retained after the stress-relaxation test (%). |
| `Modulus` | response | Apparent elastic modulus E\* from compression (kPa). |
| `A0`; `A1`,`t1`; `A2`,`t2`; `A3`,`t3` | response | Three-branch generalized Maxwell fit of the stress-relaxation curves: equilibrium stress σ~e~ (`A0`, kPa), branch stresses σ~1–3~ (`A1`–`A3`, kPa) and relaxation times τ~1–3~ (`t1`–`t3`, s). |
| `T` | response | Characteristic relaxation time τ (s): the sum of the constituent relaxation times, t1 + t2 + t3. Caution: `T` is also R's shorthand for `TRUE`; the column is therefore always accessed through the data frame (e.g. `nips$T`), never as a bare `T`. |
| `RS` | response | Residual stress σ~r~: normalized stress at the conclusion of the relaxation test (dimensionless). |

**Correspondence to the published results.** The additive models and ANOVA for Height,
Diameter, Shrinkage and Density reproduce ACS Omega Table 1 and Tables S2–S5; the one-factor
boxplots of those variables correspond to ACS Omega Figure 5. The interaction models for
Modulus, RS, T and Recovery reproduce Polymer Journal Table S2, the Spearman analysis
corresponds to Polymer Journal Figure S6 (which uses the subset Modulus, RS, T, Recovery),
and the Gibson–Ashby fit reproduces the scaling constants of Polymer Journal Figure 5b
(C = 2.04, n = 2.48). The first-order response-surface model reproduces ACS Omega Table 2,
Table S6 and Figure 6j. One caveat: the statistics involving `T` in Polymer Journal
Table S2 and Figure S6 were derived from an earlier version of the Maxwell-fit parameters,
so re-running the analysis on this final dataset yields slightly different values for that
one variable (the qualitative conclusions are unchanged). Analyses beyond these — the
Pearson panel, PCA, MANOVA, multi-response models, Box–Cox profiles and the second-order
response surface — are exploratory work that supported, but did not appear in, the
publications.

# Study data

## Import

```{r import-data}
# Read the data; the first column (ID) becomes row names, so samples keep their labels
# in plots such as the labelled PCA biplots.
nips <- read.csv("NIPS.csv", header = TRUE, row.names = 1)

# Two representations of the same dataset are used in this document:
#   dat  — SUS and EtOH stay numeric: the response-surface models need continuous
#          predictors;
#   nips — SUS and EtOH become factors: everywhere else the design levels are treated
#          as discrete groups.
dat <- nips

nips$SUS  <- as.factor(nips$SUS)
nips$EtOH <- as.factor(nips$EtOH)

# Make the baseline explicit: the first (lowest) level of each factor is the reference
# group that regression coefficients are contrasted against. (Level 1 is already the
# default reference, so this is documentation more than transformation.)
nips <- within(nips, SUS  <- relevel(SUS,  ref = 1))
nips <- within(nips, EtOH <- relevel(EtOH, ref = 1))

# head(nips)  # uncomment to preview the imported data
```

## Overview boxplots

```{r overview-boxplots}
# Quick overview of every response variable (SUS and EtOH excluded).
# Left: raw values — scales differ wildly between variables.
# Right: the same data standardised (z-scores), which makes spreads comparable
# and helps to spot outliers on a common scale.
resp_vars <- setdiff(names(nips), c("SUS", "EtOH"))
boxplot(nips[, resp_vars], main = "Box plot — raw values")
boxplot(scale(nips[, resp_vars], center = TRUE), main = "Box plot — standardised (z-scores)")
```

# Distributional checks

The linear models below assume approximately normal residuals, so each response is first
screened for normality.

## Shapiro–Wilk test

H₀: the sample comes from a normal distribution. A p-value **above** 0.05 means the data
show no significant departure from normality (consistent with the normal assumption); a
p-value below 0.05 flags a departure.

```{r shapiro-tests}
shapiro.test(nips$Weight)
shapiro.test(nips$Diameter)
shapiro.test(nips$Height)
shapiro.test(nips$Volume)
shapiro.test(nips$Density)
shapiro.test(nips$Porosity)
shapiro.test(nips$Shrinkage)
shapiro.test(nips$Recovery)
shapiro.test(nips$Modulus)
shapiro.test(nips$A0)
shapiro.test(nips$A1)
shapiro.test(nips$t1)
shapiro.test(nips$A2)
shapiro.test(nips$t2)
shapiro.test(nips$A3)
shapiro.test(nips$t3)
shapiro.test(nips$T)
```

## Normal Q–Q plots

Visual counterpart of the Shapiro–Wilk test: points falling along the reference line are
consistent with normality; systematic curvature indicates skew, and points peeling off at
the ends indicate heavy tails or outliers.

```{r qq-plots}
par(mfrow = c(3, 6))   # 17 plots on a 3 × 6 grid
qqnorm(nips$Weight, main = "Weight"); qqline(nips$Weight)
qqnorm(nips$Diameter, main = "Diameter"); qqline(nips$Diameter)
qqnorm(nips$Height, main = "Height"); qqline(nips$Height)
qqnorm(nips$Volume, main = "Volume"); qqline(nips$Volume)
qqnorm(nips$Density, main = "Density"); qqline(nips$Density)
qqnorm(nips$Porosity, main = "Porosity"); qqline(nips$Porosity)
qqnorm(nips$Shrinkage, main = "Shrinkage"); qqline(nips$Shrinkage)
qqnorm(nips$Recovery, main = "Recovery"); qqline(nips$Recovery)
qqnorm(nips$Modulus, main = "Modulus"); qqline(nips$Modulus)
qqnorm(nips$A0, main = "A0"); qqline(nips$A0)
qqnorm(nips$A1, main = "A1"); qqline(nips$A1)
qqnorm(nips$t1, main = "t1"); qqline(nips$t1)
qqnorm(nips$A2, main = "A2"); qqline(nips$A2)
qqnorm(nips$t2, main = "t2"); qqline(nips$t2)
qqnorm(nips$A3, main = "A3"); qqline(nips$A3)
qqnorm(nips$t3, main = "t3"); qqline(nips$t3)
qqnorm(nips$T, main = "T"); qqline(nips$T)
```

## Box–Cox transformation

A complementary check on the distribution of each response. `boxcox()` profiles the
log-likelihood of the power-transformation parameter λ for the model `y^(λ) ~ 1`
(λ = 1: no transformation needed; λ = 0: log transform; λ = −1: reciprocal). If the
95% confidence interval around the peak includes 1, the untransformed variable is
adequate. Note that Box–Cox requires strictly positive data.

### Log-likelihood profiles

```{r boxcox-profiles}
par(mfrow = c(3, 3))   # 16 profiles: rendered as two 3 × 3 pages
boxcox(nips$Diameter ~ 1); mtext("Diameter", side = 3, line = 1, at = 0, col = "blue", las = 1)
boxcox(nips$Height ~ 1); mtext("Height", side = 3, line = 1, at = 0, col = "blue", las = 1)
boxcox(nips$Volume ~ 1); mtext("Volume", side = 3, line = 1, at = 0, col = "blue", las = 1)
boxcox(nips$Density ~ 1); mtext("Density", side = 3, line = 1, at = 0, col = "blue", las = 1)
boxcox(nips$Porosity ~ 1); mtext("Porosity", side = 3, line = 1, at = 0, col = "blue", las = 1)
boxcox(nips$Shrinkage ~ 1); mtext("Shrinkage", side = 3, line = 1, at = 0, col = "blue", las = 1)
boxcox(nips$Recovery ~ 1); mtext("Recovery", side = 3, line = 1, at = 0, col = "blue", las = 1)
boxcox(nips$Modulus ~ 1); mtext("Modulus", side = 3, line = 1, at = 0, col = "blue", las = 1)
boxcox(nips$A0 ~ 1); mtext("A0", side = 3, line = 1, at = 0, col = "blue", las = 1)
boxcox(nips$A1 ~ 1); mtext("A1", side = 3, line = 1, at = 0, col = "blue", las = 1)
boxcox(nips$A2 ~ 1); mtext("A2", side = 3, line = 1, at = 0, col = "blue", las = 1)
boxcox(nips$A3 ~ 1); mtext("A3", side = 3, line = 1, at = 0, col = "blue", las = 1)
boxcox(nips$t1 ~ 1); mtext("t1", side = 3, line = 1, at = 0, col = "blue", las = 1)
boxcox(nips$t2 ~ 1); mtext("t2", side = 3, line = 1, at = 0, col = "blue", las = 1)
boxcox(nips$t3 ~ 1); mtext("t3", side = 3, line = 1, at = 0, col = "blue", las = 1)
boxcox(nips$T ~ 1); mtext("T", side = 3, line = 1, at = 0, col = "blue", las = 1)
```

### Optimal λ per response

The λ value that maximises each profile above, collected into a single table.

```{r boxcox-lambda}
boxcox_vars <- c("Diameter", "Height", "Volume", "Density", "Porosity", "Shrinkage",
                 "Recovery", "Modulus", "A0", "A1", "A2", "A3", "t1", "t2", "t3", "T")

lambda_opt <- sapply(boxcox_vars, function(v) {
  # boxcox() refits the model internally, so the response must be reachable from the
  # formula's own environment — extract it into a local variable and use the formula
  # interface (boxcox(y ~ 1)); boxcox(lm(nips[[v]] ~ 1)) breaks under knitting because
  # the refit cannot see `v`.
  y <- nips[[v]]
  prof <- boxcox(y ~ 1, plotit = FALSE)
  prof$x[which.max(prof$y)]     # λ at the log-likelihood maximum
})

kable(data.frame(Variable = names(lambda_opt), `Optimal lambda` = round(lambda_opt, 2),
                 check.names = FALSE),
      row.names = FALSE)
```

# Factor effects I: distributions by factor level

Distribution of the responses by factor level. Boxes show the median and interquartile
range; overlaid dots are the individual samples, so group sizes and spread are visible at
a glance.

```{r boxplot-theme}
# Shared theme for all boxplot panels in this section (identical styling throughout).
theme_box <- theme(
  legend.title = element_text(size = 14, face = "bold.italic"),
  legend.text = element_text(size = 13, face = "bold"),
  axis.title = element_text(size = 15, face = "bold"),
  axis.text = element_text(size = 14, face = "bold"),
  plot.title = element_text(size = 15, face = "bold"),
  panel.border = element_rect(fill = FALSE),
  aspect.ratio = 1,
  legend.position = "bottom"
)
```

## One-factor boxplots

Each response against one factor at a time: panel A groups by `SUS`, panel B by `EtOH`.

### Diameter

```{r box1f-diameter}
p1.1 <- ggplot(nips, aes(x = SUS, y = Diameter)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p2.1 <- ggplot(nips, aes(x = EtOH, y = Diameter)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p1.1, p2.1, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Height

```{r box1f-height}
p1.2 <- ggplot(nips, aes(x = SUS, y = Height)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p2.2 <- ggplot(nips, aes(x = EtOH, y = Height)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p1.2, p2.2, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Volume

```{r box1f-volume}
p1.3 <- ggplot(nips, aes(x = SUS, y = Volume)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p2.3 <- ggplot(nips, aes(x = EtOH, y = Volume)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p1.3, p2.3, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Density

```{r box1f-density}
p1.4 <- ggplot(nips, aes(x = SUS, y = Density)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p2.4 <- ggplot(nips, aes(x = EtOH, y = Density)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p1.4, p2.4, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Porosity

```{r box1f-porosity}
p1.5 <- ggplot(nips, aes(x = SUS, y = Porosity)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p2.5 <- ggplot(nips, aes(x = EtOH, y = Porosity)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p1.5, p2.5, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Shrinkage

```{r box1f-shrinkage}
p1.6 <- ggplot(nips, aes(x = SUS, y = Shrinkage)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p2.6 <- ggplot(nips, aes(x = EtOH, y = Shrinkage)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p1.6, p2.6, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Recovery

```{r box1f-recovery}
p1.7 <- ggplot(nips, aes(x = SUS, y = Recovery)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p2.7 <- ggplot(nips, aes(x = EtOH, y = Recovery)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p1.7, p2.7, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Modulus

```{r box1f-modulus}
p1.8 <- ggplot(nips, aes(x = SUS, y = Modulus)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p2.8 <- ggplot(nips, aes(x = EtOH, y = Modulus)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p1.8, p2.8, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### A0

```{r box1f-a0}
p1.9 <- ggplot(nips, aes(x = SUS, y = A0)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p2.9 <- ggplot(nips, aes(x = EtOH, y = A0)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p1.9, p2.9, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### A1

```{r box1f-a1}
p1.10 <- ggplot(nips, aes(x = SUS, y = A1)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p2.10 <- ggplot(nips, aes(x = EtOH, y = A1)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p1.10, p2.10, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### t1

```{r box1f-t1}
p1.11 <- ggplot(nips, aes(x = SUS, y = t1)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p2.11 <- ggplot(nips, aes(x = EtOH, y = t1)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p1.11, p2.11, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### A2

```{r box1f-a2}
p1.12 <- ggplot(nips, aes(x = SUS, y = A2)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p2.12 <- ggplot(nips, aes(x = EtOH, y = A2)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p1.12, p2.12, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### t2

```{r box1f-t2}
p1.13 <- ggplot(nips, aes(x = SUS, y = t2)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p2.13 <- ggplot(nips, aes(x = EtOH, y = t2)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p1.13, p2.13, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### A3

```{r box1f-a3}
p1.14 <- ggplot(nips, aes(x = SUS, y = A3)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p2.14 <- ggplot(nips, aes(x = EtOH, y = A3)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p1.14, p2.14, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### t3

```{r box1f-t3}
p1.15 <- ggplot(nips, aes(x = SUS, y = t3)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p2.15 <- ggplot(nips, aes(x = EtOH, y = t3)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p1.15, p2.15, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### T

```{r box1f-bigT}
# Note: inside aes(), `T` refers to the data column T (data masking), not to TRUE.
p1.16 <- ggplot(nips, aes(x = SUS, y = T)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p2.16 <- ggplot(nips, aes(x = EtOH, y = T)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p1.16, p2.16, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

## Two-factor boxplots

Both factors shown at once: one factor on the x-axis, the other as the fill colour (and
vice versa in the second panel). Parallel patterns across groups suggest additive effects;
crossing patterns hint at an interaction.

### Diameter

```{r box2f-diameter}
p3.1 <- ggplot(nips, aes(x = EtOH, y = Diameter)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p4.1 <- ggplot(nips, aes(x = SUS, y = Diameter)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p3.1, p4.1, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Height

```{r box2f-height}
p3.2 <- ggplot(nips, aes(x = EtOH, y = Height)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p4.2 <- ggplot(nips, aes(x = SUS, y = Height)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p3.2, p4.2, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Volume

```{r box2f-volume}
p3.3 <- ggplot(nips, aes(x = EtOH, y = Volume)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p4.3 <- ggplot(nips, aes(x = SUS, y = Volume)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p3.3, p4.3, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Density

```{r box2f-density}
p3.4 <- ggplot(nips, aes(x = EtOH, y = Density)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p4.4 <- ggplot(nips, aes(x = SUS, y = Density)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p3.4, p4.4, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Porosity

```{r box2f-porosity}
p3.5 <- ggplot(nips, aes(x = EtOH, y = Porosity)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p4.5 <- ggplot(nips, aes(x = SUS, y = Porosity)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p3.5, p4.5, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Shrinkage

```{r box2f-shrinkage}
p3.6 <- ggplot(nips, aes(x = EtOH, y = Shrinkage)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p4.6 <- ggplot(nips, aes(x = SUS, y = Shrinkage)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p3.6, p4.6, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Recovery

```{r box2f-recovery}
p3.7 <- ggplot(nips, aes(x = EtOH, y = Recovery)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p4.7 <- ggplot(nips, aes(x = SUS, y = Recovery)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p3.7, p4.7, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

### Modulus

```{r box2f-modulus}
p3.8 <- ggplot(nips, aes(x = EtOH, y = Modulus)) + geom_boxplot(aes(fill = SUS), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = SUS), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

p4.8 <- ggplot(nips, aes(x = SUS, y = Modulus)) + geom_boxplot(aes(fill = EtOH), position = position_dodge(.8)) +
  geom_dotplot(aes(fill = EtOH), binaxis = 'y', stackdir = 'center', position = position_dodge(.8)) +
  theme_box

plot_grid(p3.8, p4.8, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

# Factor effects II: regression & ANOVA

The graphical impressions from the boxplot gallery are quantified here with linear
models on the factor version of the data.

## Per-response linear models & ANOVA

Each response is modelled as `response ~ SUS + EtOH`, i.e. an additive two-factor model
with both factors categorical. For every response, four outputs are shown:

- `summary()` — coefficient estimates (each level contrasted against the reference level),
  their significance, and the model's R².
- `anova()` — the ANOVA table testing the overall effect of each factor.
- The same two tables re-rendered with `tidy()`/`kable()` as compact, publication-ready
  tables.
- Standard diagnostic plots (residuals vs fitted, normal Q–Q, scale–location, residuals vs
  leverage) to check the model assumptions.

For selected responses, a second model adds the `SUS:EtOH` interaction to test whether the
effect of one factor depends on the level of the other. (In R's formula notation,
`SUS + EtOH + SUS*EtOH` expands to main effects plus the interaction term — it is
equivalent to `SUS*EtOH`.) The four responses analysed with interaction in the Polymer
Journal paper (Table S2) are Modulus, RS, T and Recovery; the additive models for Height,
Diameter, Shrinkage and Density are the ones published in ACS Omega (Table 1, Tables S2–S5).

### Diameter

```{r lm-diameter}
mod1 <- lm(nips$Diameter ~ SUS + EtOH, data = nips)
mod1 %>% summary()
mod1 %>% anova()
mod1 %>% tidy() %>% kable(digits = 3)
mod1 %>% anova %>% kable(digits = 3)
```

```{r lm-diameter-diag}
par(mfrow = c(2, 2))
mod1 %>% plot()
```

### Height

```{r lm-height}
mod2 <- lm(nips$Height ~ SUS + EtOH, data = nips)
mod2 %>% summary()
mod2 %>% anova()
mod2 %>% tidy() %>% kable(digits = 3)
mod2 %>% anova %>% kable(digits = 3)
```

```{r lm-height-diag}
par(mfrow = c(2, 2))
mod2 %>% plot()
```

### Volume

```{r lm-volume}
mod3 <- lm(nips$Volume ~ SUS + EtOH, data = nips)
mod3 %>% summary()
mod3 %>% anova()
mod3 %>% tidy() %>% kable(digits = 3)
mod3 %>% anova %>% kable(digits = 3)
```

```{r lm-volume-diag}
par(mfrow = c(2, 2))
mod3 %>% plot()
```

### Density

```{r lm-density}
mod4 <- lm(nips$Density ~ SUS + EtOH, data = nips)
mod4 %>% summary()
mod4 %>% anova()
mod4 %>% tidy() %>% kable(digits = 3)
mod4 %>% anova %>% kable(digits = 3)
```

```{r lm-density-diag}
par(mfrow = c(2, 2))
mod4 %>% plot()
```

### Porosity

```{r lm-porosity}
mod5 <- lm(nips$Porosity ~ SUS + EtOH, data = nips)
mod5 %>% summary()
mod5 %>% anova()
mod5 %>% tidy() %>% kable(digits = 3)
mod5 %>% anova %>% kable(digits = 3)
```

```{r lm-porosity-diag}
par(mfrow = c(2, 2))
mod5 %>% plot()
```

### Shrinkage

```{r lm-shrinkage}
mod6 <- lm(nips$Shrinkage ~ SUS + EtOH, data = nips)
mod6 %>% summary()
mod6 %>% anova()
mod6 %>% tidy() %>% kable(digits = 3)
mod6 %>% anova %>% kable(digits = 3)
```

```{r lm-shrinkage-diag}
par(mfrow = c(2, 2))
mod6 %>% plot()
```

### Recovery

```{r lm-recovery}
mod7 <- lm(nips$Recovery ~ SUS + EtOH, data = nips)
mod7 %>% summary()
mod7 %>% anova()
mod7 %>% tidy() %>% kable(digits = 3)
mod7 %>% anova %>% kable(digits = 3)
```

```{r lm-recovery-diag}
par(mfrow = c(2, 2))
mod7 %>% plot()
```

#### Recovery with interaction

```{r lm-recovery-int}
mod7a <- lm(nips$Recovery ~ SUS + EtOH + SUS*EtOH, data = nips)
mod7a %>% summary()
mod7a %>% anova()
mod7a %>% tidy() %>% kable(digits = 3)
mod7a %>% anova %>% kable(digits = 3)
```

```{r lm-recovery-int-diag}
par(mfrow = c(2, 2))
mod7a %>% plot()
```

### Modulus

```{r lm-modulus}
mod8 <- lm(nips$Modulus ~ SUS + EtOH, data = nips)
mod8 %>% summary()
mod8 %>% anova()
mod8 %>% tidy() %>% kable(digits = 3)
mod8 %>% anova %>% kable(digits = 3)
```

```{r lm-modulus-diag}
par(mfrow = c(2, 2))
mod8 %>% plot()
```

#### Modulus with interaction

```{r lm-modulus-int}
mod8a <- lm(nips$Modulus ~ SUS + EtOH + SUS*EtOH, data = nips)
mod8a %>% summary()
mod8a %>% anova()
mod8a %>% tidy() %>% kable(digits = 3)
mod8a %>% anova %>% kable(digits = 3)
```

```{r lm-modulus-int-diag}
par(mfrow = c(2, 2))
mod8a %>% plot()
```

### A0

```{r lm-a0}
mod9 <- lm(nips$A0 ~ SUS + EtOH, data = nips)
mod9 %>% summary()
mod9 %>% anova()
mod9 %>% tidy() %>% kable(digits = 3)
mod9 %>% anova %>% kable(digits = 3)
```

```{r lm-a0-diag}
par(mfrow = c(2, 2))
mod9 %>% plot()
```

#### A0 with interaction

```{r lm-a0-int}
mod9a <- lm(nips$A0 ~ SUS + EtOH + SUS*EtOH, data = nips)
mod9a %>% summary()
mod9a %>% anova()
mod9a %>% tidy() %>% kable(digits = 3)
mod9a %>% anova %>% kable(digits = 3)
```

```{r lm-a0-int-diag}
par(mfrow = c(2, 2))
mod9a %>% plot()
```

### A1

```{r lm-a1}
mod10 <- lm(nips$A1 ~ SUS + EtOH, data = nips)
mod10 %>% summary()
mod10 %>% anova()
mod10 %>% tidy() %>% kable(digits = 3)
mod10 %>% anova %>% kable(digits = 3)
```

```{r lm-a1-diag}
par(mfrow = c(2, 2))
mod10 %>% plot()
```

### t1

```{r lm-t1}
mod11 <- lm(nips$t1 ~ SUS + EtOH, data = nips)
mod11 %>% summary()
mod11 %>% anova()
mod11 %>% tidy() %>% kable(digits = 3)
mod11 %>% anova %>% kable(digits = 3)
```

```{r lm-t1-diag}
par(mfrow = c(2, 2))
mod11 %>% plot()
```

#### t1 with interaction

```{r lm-t1-int}
mod11a <- lm(nips$t1 ~ SUS + EtOH + SUS*EtOH, data = nips)
mod11a %>% summary()
mod11a %>% anova()
mod11a %>% tidy() %>% kable(digits = 3)
mod11a %>% anova %>% kable(digits = 3)
```

```{r lm-t1-int-diag}
par(mfrow = c(2, 2))
mod11a %>% plot()
```

### A2

```{r lm-a2}
mod12 <- lm(nips$A2 ~ SUS + EtOH, data = nips)
mod12 %>% summary()
mod12 %>% anova()
mod12 %>% tidy() %>% kable(digits = 3)
mod12 %>% anova %>% kable(digits = 3)
```

```{r lm-a2-diag}
par(mfrow = c(2, 2))
mod12 %>% plot()
```

### t2

```{r lm-t2}
mod13 <- lm(nips$t2 ~ SUS + EtOH, data = nips)
mod13 %>% summary()
mod13 %>% anova()
mod13 %>% tidy() %>% kable(digits = 3)
mod13 %>% anova %>% kable(digits = 3)
```

```{r lm-t2-diag}
par(mfrow = c(2, 2))
mod13 %>% plot()
```

#### t2 with interaction

```{r lm-t2-int}
mod13a <- lm(nips$t2 ~ SUS + EtOH + SUS*EtOH, data = nips)
mod13a %>% summary()
mod13a %>% anova()
mod13a %>% tidy() %>% kable(digits = 3)
mod13a %>% anova %>% kable(digits = 3)
```

```{r lm-t2-int-diag}
par(mfrow = c(2, 2))
mod13a %>% plot()
```

### A3

```{r lm-a3}
mod14 <- lm(nips$A3 ~ SUS + EtOH, data = nips)
mod14 %>% summary()
mod14 %>% anova()
mod14 %>% tidy() %>% kable(digits = 3)
mod14 %>% anova %>% kable(digits = 3)
```

```{r lm-a3-diag}
par(mfrow = c(2, 2))
mod14 %>% plot()
```

### t3

```{r lm-t3}
mod15 <- lm(nips$t3 ~ SUS + EtOH, data = nips)
mod15 %>% summary()
mod15 %>% anova()
mod15 %>% tidy() %>% kable(digits = 3)
mod15 %>% anova %>% kable(digits = 3)
```

```{r lm-t3-diag}
par(mfrow = c(2, 2))
mod15 %>% plot()
```

#### t3 with interaction

```{r lm-t3-int}
mod15a <- lm(nips$t3 ~ SUS + EtOH + SUS*EtOH, data = nips)
mod15a %>% summary()
mod15a %>% anova()
mod15a %>% tidy() %>% kable(digits = 3)
mod15a %>% anova %>% kable(digits = 3)
```

```{r lm-t3-int-diag}
par(mfrow = c(2, 2))
mod15a %>% plot()
```

### T

```{r lm-bigT}
mod16 <- lm(nips$T ~ SUS + EtOH, data = nips)
mod16 %>% summary()
mod16 %>% anova()
mod16 %>% tidy() %>% kable(digits = 3)
mod16 %>% anova %>% kable(digits = 3)
```

```{r lm-bigT-diag}
par(mfrow = c(2, 2))
mod16 %>% plot()
```

#### T with interaction

```{r lm-bigT-int}
mod16a <- lm(nips$T ~ SUS + EtOH + SUS*EtOH, data = nips)
mod16a %>% summary()
mod16a %>% anova()
mod16a %>% tidy() %>% kable(digits = 3)
mod16a %>% anova %>% kable(digits = 3)
```

```{r lm-bigT-int-diag}
par(mfrow = c(2, 2))
mod16a %>% plot()
```

### Residual stress (RS)

```{r lm-rs}
mod17 <- lm(nips$RS ~ SUS + EtOH, data = nips)
mod17 %>% summary()
mod17 %>% anova()
mod17 %>% tidy() %>% kable(digits = 3)
mod17 %>% anova %>% kable(digits = 3)
```

```{r lm-rs-diag}
par(mfrow = c(2, 2))
mod17 %>% plot()
```

#### Residual stress (RS) with interaction

```{r lm-rs-int}
mod17a <- lm(nips$RS ~ SUS + EtOH + SUS*EtOH, data = nips)
mod17a %>% summary()
mod17a %>% anova()
mod17a %>% tidy() %>% kable(digits = 3)
mod17a %>% anova %>% kable(digits = 3)
```

```{r lm-rs-int-diag}
par(mfrow = c(2, 2))
mod17a %>% plot()
```

# Correlations between responses

Pairwise associations between selected responses. Each panel figure shows scatter plots
(below the diagonal), variable histograms with density curves (diagonal), and correlation
coefficients with significance stars (above the diagonal): \*\*\* p < 0.001, \*\* p < 0.01,
\* p < 0.05.

## Pearson

Pearson's *r* measures **linear** association and is used here for the wider variable set.

```{r pearson-panel}
# Variables entering the Pearson panel (selection used in the papers).
vars_pearson <- c("Density", "Shrinkage", "Recovery", "Modulus", "A0", "t1", "t2", "t3", "T")

pairs.panels(
  nips[, vars_pearson],
  pch = 21,
  cex.cor = 1.2,     # size of the correlation labels
  gap = 0.5,
  scale = FALSE,     # do not scale label size by correlation strength
  density = TRUE,    # add density curves to the diagonal histograms
  ellipses = FALSE,
  lm = FALSE,
  smooth = FALSE,
  main = "Pearson_Correlation_Analysis",
  cex.main = 1.5,
  font.main = 2,
  hist.col = 4,
  stars = TRUE       # significance stars next to the coefficients
)
```

## Spearman

Spearman's ρ is rank-based, so it captures **monotonic** (not necessarily linear)
relationships and is robust to outliers. A reduced variable set is examined here. (Note:
the published Spearman figure — Polymer Journal Figure S6 — uses the subset `Modulus`,
`RS`, `T`, `Recovery`; the panel below is the repository's variant of that analysis.)

```{r spearman-panel}
# Variables entering the Spearman panel and the exported matrices below.
# The published figure (Polymer Journal Fig. S6) used c("Modulus", "RS", "T", "Recovery").
vars_spearman <- c("Density", "Recovery", "Modulus", "A0", "T")

pairs.panels(
  nips[, vars_spearman],
  method = "spearman",
  pch = 21,
  cex.cor = 1,
  gap = 0.5,
  scale = FALSE,
  density = TRUE,
  ellipses = FALSE,
  lm = FALSE,
  smooth = FALSE,
  main = "Spearman_Correlation",
  cex.main = 1.5,
  font.main = 2,
  hist.col = 8,
  stars = TRUE
)
```

```{r spearman-export}
# Spearman correlation matrix for the same variable set...
correlation_matrix <- cor(nips[, vars_spearman], method = "spearman")

# ...and the matching matrix of p-values. cor() alone does not return p-values,
# so this helper runs cor.test() on every variable pair and collects the results
# into a symmetric matrix (diagonal = 0 by convention).
cor.mtest <- function(mat, ...) {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], method = "spearman", ...)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  p.mat
}

p_matrix <- cor.mtest(nips[, vars_spearman])

decimal_places <- 3
rounded_correlation_matrix <- round(correlation_matrix, decimal_places)
rounded_p_matrix <- round(p_matrix, decimal_places)
rounded_correlation_matrix
rounded_p_matrix

# Side effect: export both matrices as CSV files next to this document
# (used for the papers' supplementary tables).
write.csv(rounded_correlation_matrix, "spearman_correlation_matrix.csv")
write.csv(rounded_p_matrix, "spearman_p_matrix.csv")
```

# Multivariate analysis

The responses are interrelated, so this part analyses them jointly: principal component
analysis condenses the eight mechanical variables into a few uncorrelated components,
multi-response regression and MANOVA test the factor effects on several responses at
once, and the component scores themselves are regressed on the factors.

## Principal component analysis (PCA)

Principal component analysis of the **mechanical/stress-relaxation variables** (modulus plus
the exponential-fit parameters). PCA re-expresses the correlated variables as a small number
of uncorrelated components, revealing which properties vary together and how the samples
separate by processing condition. Variables are standardised internally by `PCA()`.

```{r pca-fit}
# Variables entering the PCA (8 variables → at most 8 components).
vars_pca <- c("Modulus", "A0", "A1", "t1", "A2", "t2", "A3", "t3")
res.pca <- PCA(nips[, vars_pca], graph = FALSE, ncp = 8)

# Eigenvalues: variance explained by each component.
eig <- as.data.frame(res.pca[["eig"]])
eig

# Variable coordinates (loadings) on the components.
coord <- as.data.frame(res.pca[["var"]][["coord"]])
coord

# Contribution (%) of each variable to each component.
contrib <- as.data.frame(res.pca[["var"]][["contrib"]])
contrib

# Squared cosines: quality of representation of each variable on each component.
cos2 <- as.data.frame(res.pca[["var"]][["cos2"]])
cos2

# Individual (sample) coordinates, i.e. the PC scores — reused later for
# regression on principal components.
coord_ind <- as.data.frame(res.pca[["ind"]][["coord"]])
coord_ind
```

### Scree plot and variable map

The scree plot shows how much variance each component explains — components before the
"elbow" summarise most of the structure. The variable correlation plot (loading map) places
each variable by its correlation with PC1 and PC2; variables pointing the same way are
positively correlated, and colour encodes representation quality (cos²). The contribution
bar charts identify which variables define each of the first three components (the dashed
line marks the contribution expected if all variables contributed equally).

```{r pca-plot-objects}
# Build the PCA summary plots; they are arranged into panels in the next two chunks.
a1 <- fviz_eig(res.pca, addlabels = TRUE, ylim = c(1, 90),
               title = "Scree plot", ylab = "Explained variances (%)") +
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 11, face = "bold"),
    plot.title = element_text(size = 16, face = "bold"),
    panel.border = element_rect(fill = FALSE),
    legend.text = element_text(size = 15, face = "bold"),
    aspect.ratio = 1.7,
  )

b1 <- fviz_pca_var(res.pca,
                   title = "Variables \ncorrelation plot",
                   col.var = "cos2",
                   gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
                   repel = TRUE,
                   arr.width = 25,) +
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 11, face = "bold"),
    plot.title = element_text(size = 16, face = "bold"),
    legend.title = element_text(size = 15, face = "bold"),
    legend.text = element_text(size = 15, face = "bold"),
    aspect.ratio = 1,
  )

# Shared theme for the three per-dimension contribution charts.
theme_contrib <- theme(
  axis.title = element_text(size = 14, face = "bold"),
  axis.text = element_text(size = 12, face = "bold"),
  plot.title = element_text(size = 16, face = "bold"),
  panel.border = element_rect(fill = FALSE),
  legend.text = element_text(size = 15, face = "bold"),
  aspect.ratio = 1.8,
)
c1 <- fviz_contrib(res.pca, choice = "var", axes = 1, title = "Contribution \nDim1") + theme_contrib
d1 <- fviz_contrib(res.pca, choice = "var", axes = 2, title = "Contribution \nDim2") + theme_contrib
e1 <- fviz_contrib(res.pca, choice = "var", axes = 3, title = "Contribution \nDim3") + theme_contrib
```

```{r pca-scree-loadings}
plot_grid(a1, b1, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15,
          rel_widths = c(1, 1), rel_heights = c(1, 1))
```

```{r pca-contributions}
plot_grid(c1, d1, e1, ncol = 3, nrow = 1, labels = c("A", "B", "C"), label_size = 15,
          rel_widths = c(1, 1), rel_heights = c(1, 1))
```

### Biplots

Each biplot overlays the samples (points, scored on the components) with the variable
loadings (arrows). Points close together are samples with similar mechanical profiles;
a sample lying in the direction of an arrow has a high value of that variable. The same
plot is drawn twice — coloured by `SUS` (panel A) and by `EtOH` (panel B) — with 95%
concentration ellipses per group, to show which factor drives the separation.

```{r pca-theme-defs}
# Shared themes for the biplot figures below (values identical across panels).
theme_biplot <- theme(          # side-by-side biplot pairs
  legend.title = element_text(size = 14, face = "bold.italic"),
  legend.text = element_text(size = 13, face = "bold"),
  axis.title = element_text(size = 15, face = "bold"),
  axis.text = element_text(size = 14, face = "bold"),
  plot.title = element_text(size = 16, face = "bold"),
  panel.border = element_rect(fill = FALSE),
  aspect.ratio = .6,
  legend.position = "bottom"
)
theme_biplot_lab <- theme(      # single, sample-labelled biplots
  legend.title = element_text(size = 14, face = "bold.italic"),
  legend.text = element_text(size = 13, face = "bold"),
  axis.title = element_text(size = 15, face = "bold"),
  axis.text = element_text(size = 13, face = "bold"),
  plot.title = element_text(size = 15, face = "bold"),
  panel.border = element_rect(fill = FALSE),
  aspect.ratio = .65
)
```

#### Biplot: PC1 vs PC2

```{r pca-biplot-12}
a2 <- fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(1, 2),
  geom.ind = "point",
  pointshape = 21,
  pointsize = 4,
  fill.ind = nips$SUS,          # colour samples by suspension concentration
  col.ind = "black",
  addEllipses = TRUE,          # 95% concentration ellipse per group
  col.var = "black",
  arr.width = 25,
  legend.title = list(fill = "SUS\n (%)"),
  mean.point = FALSE,
  repel = TRUE,
  ) +
  guides(fill = guide_legend(order = 1), color = "none") +
  theme_biplot

b2 <- fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(1, 2),
  geom.ind = "point",
  pointshape = 21,
  pointsize = 4,
  fill.ind = nips$EtOH,         # same plot, coloured by ethanol concentration
  col.ind = "black",
  addEllipses = TRUE,
  col.var = "black",
  arr.width = 25,
  legend.title = list(fill = "EtOH"),
  mean.point = FALSE,
  repel = TRUE,
  ) +
  guides(fill = guide_legend(order = 1), color = "none") +
  theme_biplot

plot_grid(a2, b2, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

#### Biplot: PC1 vs PC3

```{r pca-biplot-13}
a3 <- fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(1, 3),
  geom.ind = "point",
  pointshape = 21,
  pointsize = 4,
  fill.ind = nips$SUS,
  col.ind = "black",
  addEllipses = TRUE,
  col.var = "black",
  arr.width = 25,
  legend.title = list(fill = "SUS\n (%)"),
  mean.point = FALSE,
  repel = TRUE,
  ) +
  guides(fill = guide_legend(order = 1), color = "none") +
  theme_biplot

b3 <- fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(1, 3),
  geom.ind = "point",
  pointshape = 21,
  pointsize = 4,
  fill.ind = nips$EtOH,
  col.ind = "black",
  addEllipses = TRUE,
  col.var = "black",
  arr.width = 25,
  legend.title = list(fill = "EtOH\n (%)"),
  mean.point = FALSE,
  repel = TRUE,
  ) +
  guides(fill = guide_legend(order = 1), color = "none") +
  theme_biplot

plot_grid(a3, b3, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

#### Biplot: PC2 vs PC3

```{r pca-biplot-23}
a4 <- fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(2, 3),
  geom.ind = "point",
  pointshape = 21,
  pointsize = 4,
  fill.ind = nips$SUS,
  col.ind = "black",
  addEllipses = TRUE,
  col.var = "black",
  arr.width = 25,
  legend.title = list(fill = "SUS\n (%)"),
  mean.point = FALSE,
  repel = TRUE,
  ) +
  guides(fill = guide_legend(order = 1), color = "none") +
  theme_biplot

b4 <- fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(2, 3),
  geom.ind = "point",
  pointshape = 21,
  pointsize = 4,
  fill.ind = nips$EtOH,
  col.ind = "black",
  addEllipses = TRUE,
  col.var = "black",
  arr.width = 25,
  legend.title = list(fill = "EtOH\n (%)"),
  mean.point = FALSE,
  repel = TRUE,
  ) +
  guides(fill = guide_legend(order = 1), color = "none") +
  theme_biplot

plot_grid(a4, b4, ncol = 2, nrow = 1, labels = c("A", "B"), label_size = 15)
```

#### Biplot: PC1 vs PC2 with sample labels

Same projection as above, but each point is labelled with its sample ID (the `ID` column,
kept as row names on import) — useful for identifying individual samples and outliers.

```{r pca-biplot-12-labels-sus}
fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(1, 2),
  geom.ind = c("point", "text"),   # points plus sample-ID labels
  pointshape = 21,
  pointsize = 4,
  fill.ind = nips$SUS,
  col.ind = "black",
  addEllipses = FALSE,
  col.var = "black",
  arr.width = 25,
  legend.title = list(fill = "SUS (%)"),
  mean.point = FALSE,
  repel = TRUE,
  ) +
  guides(fill = guide_legend(order = 1), color = "none") +
  theme_biplot_lab
```

```{r pca-biplot-12-labels-etoh}
fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(1, 2),
  geom.ind = c("point", "text"),
  pointshape = 21,
  pointsize = 4,
  fill.ind = nips$EtOH,
  col.ind = "black",
  addEllipses = FALSE,
  col.var = "black",
  arr.width = 25,
  legend.title = list(fill = "EtOH (%)"),
  mean.point = FALSE,
  repel = TRUE,
  ) +
  guides(fill = guide_legend(order = 1), color = "none") +
  theme_biplot_lab
```

#### Biplot: PC7 vs PC8

Diagnostic view of the two **last** components. These carry almost no variance; a
near-zero component corresponds to an almost exact linear dependency among the variables,
so this plot mainly serves as a check on redundancy/collinearity in the variable set.

```{r pca-biplot-78}
fviz_pca_biplot(
  title = "PCA biplot",
  res.pca,
  axes = c(7, 8),
  geom.ind = c("point", "text"),
  pointshape = 21,
  pointsize = 4,
  fill.ind = nips$SUS,
  col.ind = "black",
  addEllipses = FALSE,
  col.var = "black",
  arr.width = 25,
  legend.title = list(fill = "SUS (%)"),
  mean.point = FALSE,
  repel = TRUE,
  ) +
  guides(fill = guide_legend(order = 1), color = "none") +
  theme_biplot_lab
```

## Multi-response (multivariate) regression & ANOVA

Here several responses are fitted **jointly** against the same predictors with
`lm(cbind(...) ~ ...)`. The coefficients equal those of the separate per-response models,
but `anova()` on a multi-response fit performs multivariate (MANOVA-type) tests, taking the
correlations between the responses into account.

### lm(cbind(Modulus, A0, T, Recovery) ~ SUS + EtOH)

```{r mlm-additive}
mod2.1 <- lm(cbind(nips$Modulus, nips$A0, nips$T, nips$Recovery) ~ SUS + EtOH, data = nips)
mod2.1 %>% summary()
mod2.1 %>% anova()
mod2.1 %>% tidy() %>% kable(digits = 3)
mod2.1 %>% anova %>% kable(digits = 3)
```

### lm(cbind(Modulus, A0, T, Recovery) ~ SUS + EtOH + SUS\*EtOH)

```{r mlm-interaction}
mod2.1a <- lm(cbind(nips$Modulus, nips$A0, nips$T, nips$Recovery) ~ SUS + EtOH + SUS*EtOH, data = nips)
mod2.1a %>% summary()
mod2.1a %>% anova()
mod2.1a %>% tidy() %>% kable(digits = 3)
mod2.1a %>% anova %>% kable(digits = 3)
```

## Regression & ANOVA on principal-component scores

Instead of modelling each variable separately, the first PCA scores are used as *composite*
responses: Dim.1 and Dim.2 summarise the joint variation of all eight mechanical variables,
so factor effects on the scores indicate effects on the overall mechanical profile.

### lm(cbind(Dim.1, Dim.2) ~ SUS + EtOH + SUS\*EtOH)

```{r pca-scores-mlm}
# Sample scores on the principal components, from the PCA fitted above.
score <- data.frame(res.pca$ind$coord)

mod3.1 <- lm(cbind(score$Dim.1, score$Dim.2) ~ SUS + EtOH + SUS*EtOH, data = nips)
mod3.1 %>% summary()
mod3.1 %>% anova()
mod3.1 %>% tidy() %>% kable(digits = 3)
mod3.1 %>% anova %>% kable(digits = 3)
```

### lm(Dim.1 ~ SUS + EtOH)

```{r pca-scores-dim1}
mod3.2 <- lm(score$Dim.1 ~ SUS + EtOH, data = nips)
mod3.2 %>% summary()
mod3.2 %>% anova()
mod3.2 %>% tidy() %>% kable(digits = 3)
mod3.2 %>% anova %>% kable(digits = 3)
```

```{r pca-scores-dim1-diag}
par(mfrow = c(2, 2))
mod3.2 %>% plot()
```

### lm(Dim.1 ~ SUS + EtOH + SUS\*EtOH)

```{r pca-scores-dim1-int}
mod3.3 <- lm(score$Dim.1 ~ SUS + EtOH + SUS*EtOH, data = nips)
mod3.3 %>% summary()
mod3.3 %>% anova()
mod3.3 %>% tidy() %>% kable(digits = 3)
mod3.3 %>% anova %>% kable(digits = 3)
```

```{r pca-scores-dim1-int-diag}
par(mfrow = c(2, 2))
mod3.3 %>% plot()
```

## MANOVA

Multivariate analysis of variance tests whether the mean vector of the eight mechanical
variables (`Modulus`, `A0`, `A1`, `t1`, `A2`, `t2`, `A3`, `t3`) differs between factor
levels — one joint test instead of eight separate ANOVAs. Four standard test statistics
are reported: **Pillai's trace** (most robust to assumption violations), **Wilks' lambda**
(the classical likelihood-ratio criterion), **Hotelling–Lawley trace**, and **Roy's largest
root** (most powerful when group differences lie along a single dimension, but most
liberal). Agreement across all four strengthens the conclusion. The `summary.aov()` output
then breaks the joint result down into the univariate follow-up ANOVA for each response.

### SUS

`manova(cbind(Modulus, A0, A1, t1, A2, t2, A3, t3) ~ SUS)` — joint test of the SUS effect:

```{r manova-sus}
mod4.1 <- manova(cbind(nips$Modulus, nips$A0, nips$A1, nips$t1, nips$A2, nips$t2, nips$A3, nips$t3) ~ SUS, data = nips)
mod4.1 %>% summary.manova(intercept = TRUE, "Pillai")
mod4.1 %>% summary.manova(intercept = TRUE, "Wilks")
mod4.1 %>% summary.manova(intercept = TRUE, "Hotelling-Lawley")
mod4.1 %>% summary.manova(intercept = TRUE, "Roy")
```

Univariate follow-up ANOVAs (one per response):

```{r manova-sus-aov}
mod4.1 %>% summary.aov()
mod4.1 %>% tidy() %>% kable(digits = 3)
```

### EtOH

`manova(cbind(Modulus, A0, A1, t1, A2, t2, A3, t3) ~ EtOH)` — joint test of the EtOH effect:

```{r manova-etoh}
mod5.1 <- manova(cbind(nips$Modulus, nips$A0, nips$A1, nips$t1, nips$A2, nips$t2, nips$A3, nips$t3) ~ EtOH, data = nips)
mod5.1 %>% summary.manova(intercept = TRUE, "Pillai")
mod5.1 %>% summary.manova(intercept = TRUE, "Wilks")
mod5.1 %>% summary.manova(intercept = TRUE, "Hotelling-Lawley")
mod5.1 %>% summary.manova(intercept = TRUE, "Roy")
```

Univariate follow-up ANOVAs (one per response):

```{r manova-etoh-aov}
mod5.1 %>% summary.aov()
mod5.1 %>% tidy() %>% kable(digits = 3)
```

# Response-surface analysis (continuous factors)

Everywhere above, `SUS` and `EtOH` are treated as categorical design levels. This part
instead treats them as **continuous** variables and uses response surface methodology
(RSM) to model `Shrinkage` (and, at the end, `Density`) as a smooth function of both
factors, working on the numeric data copy `dat`. This view supports interpolation between
the tested levels and indicates in which direction the response decreases fastest.

The first-order Shrinkage model below is the analysis published in the ACS Omega paper:
its coefficients appear there as Table 2, its ANOVA (including the lack-of-fit test) as
Table S6, and its contour plot as Figure 6j. The RSM prediction was validated with three
additional sponges fabricated at EtOH = 3 wt% (one per SUS level; ACS Omega SI Table S1,
rows 43–45; provided here as `NIPS_validation.csv`):
the SUS 0.4 / EtOH 3 sponge showed 49.4% shrinkage, consistent with the predicted trend.
The second-order model and canonical analysis are exploratory extensions beyond the
published analysis.

## First-order (FO) model

`FO(SUS, EtOH)` fits a first-order response surface — a plane over the two factors. The
`summary()` includes a lack-of-fit test, which indicates whether this simple plane is
adequate or whether curvature (a second-order model) is needed. The predictors enter in
their original units (uncoded), so coefficients are interpretable directly on the
concentration scales. The coefficient table equals ACS Omega Table 2, and the ANOVA block
of the summary equals Table S6 (model P = 0.011; lack of fit insignificant, P = 0.260).

```{r rsm-fo-fit}
rsm1 <- rsm(Shrinkage ~ FO(SUS, EtOH), data = dat)
summary(rsm1)
rsm1 %>% tidy() %>% kable(digits = 3)
```

### Contour plot of the fitted plane

Filled contour map of predicted Shrinkage. The red rectangle marks the experimental
(design) region actually covered by the study — predictions outside it are extrapolation
and should be read with caution.

```{r rsm-fo-contour}
contour(rsm1, ~ SUS + EtOH, image = TRUE, bounds = list(EtOH = c(0.1, 4), SUS = c(0, 0.5)),
        zlim = c(40, 60), decode = FALSE)
lines(c(0.4, 0.24, 0.24, 0.4, 0.4), c(0.2, 0.2, 1, 1, 0.2), col = "red")  # design region
```

## Second-order (SO) model & canonical analysis

`SO(SUS, EtOH)` adds quadratic and interaction terms, allowing a curved surface. The
canonical analysis then locates the stationary point of that surface and reports the
eigenvalues of the quadratic form: all negative → maximum, all positive → minimum, mixed
signs → saddle. `threshold = 0` keeps the eigenvalues exactly as estimated (no
thresholding of near-zero values).

```{r rsm-so-fit}
rsm2 <- rsm(Shrinkage ~ SO(SUS, EtOH), data = dat)
canonical(rsm2, threshold = 0)
```

### Contour plot with the steepest-descent path

The blue circles trace the **path of steepest descent** — the direction along which
predicted Shrinkage decreases fastest from the stationary point (marked with a distinct
symbol at distance 0). The filled blue square highlights the validation condition
SUS = 0.4 wt%, EtOH = 3 wt%, where a fabricated sponge showed 49.4% shrinkage, consistent
with the predicted trend. The red rectangle is again the design region.

```{r rsm-so-contour-path}
contour(rsm2, EtOH ~ SUS, image = TRUE, bounds = list(EtOH = c(0.1, 4), SUS = c(0, 0.5)),
        zlim = c(30, 80), decode = FALSE)
lines(c(0.4, 0.24, 0.24, 0.4, 0.4), c(0.2, 0.2, 1, 1, 0.2), col = "red")  # design region
points(EtOH ~ SUS, data = canonical.path(rsm2, descent = TRUE, threshold = 0),
       col = "blue", pch = 1 + 6*(dist == 0))   # pch 7 marks the stationary point (dist == 0)
points(0.4, 3, col = "black", pch = 22, cex = 2, bg = "blue")  # validation condition (SUS 0.4, EtOH 3 wt%; measured shrinkage 49.4%)
```

### Contour plot centred at the stationary point

The same second-order surface, re-drawn with the slice taken through the stationary point
(`at = summary(rsm2)$canonical$xs`) and a wider EtOH range.

```{r rsm-so-contour-stationary}
contour(rsm2, EtOH ~ SUS, image = TRUE, at = summary(rsm2)$canonical$xs,
        bounds = list(EtOH = c(0.1, 5), SUS = c(0, 0.5)), zlim = c(40, 100), decode = FALSE)
```

## Predicted means over the design grid

Estimated marginal means of Shrinkage from the first-order model, evaluated at the nine
design combinations of the study (SUS 0.24/0.3/0.4 wt% × EtOH 0.2/0.5/1 wt%;
`mode = "decoded"` reports them on the original concentration scales).

```{r rsm-emmeans-grid}
# Note: emmeans reconstructs the data from the model call, so `dat` must exist in the
# workspace — which is the case when this document is knitted normally.
emmeans::emmeans(rsm1, ~ SUS * EtOH, mode = "decoded",
                 at = list(SUS = c(0.24, 0.3, 0.4), EtOH = c(0.2, 0.5, 1)))
```

## Cross-checks with ordinary lm()

The same first-order surface can be fitted with plain `lm()`; these fits reproduce the
RSM results and add the standard regression diagnostics.

### First-degree polynomial fit

`poly(SUS, EtOH, degree = 1)` spans the same model space as `FO(SUS, EtOH)`, so this
contour should match the FO contour above. (The point SUS = 0.4, EtOH = 3 is the same
validation condition as before.)

```{r rsm-fo-poly-check}
fo_poly <- lm(Shrinkage ~ poly(SUS, EtOH, degree = 1), data = dat)
contour(fo_poly, ~ SUS + EtOH, image = TRUE, bounds = list(EtOH = c(0.1, 4), SUS = c(0, 0.5)),
        zlim = c(40, 60), decode = FALSE)
lines(c(0.4, 0.24, 0.24, 0.4, 0.4), c(0.2, 0.2, 1, 1, 0.2), col = "red")  # design region
points(0.4, 3, col = "black", pch = 22, cex = 2, bg = "blue")  # validation condition (SUS 0.4, EtOH 3 wt%; measured shrinkage 49.4%)
```

### Additive linear model for Shrinkage, with diagnostics

Plain additive regression `Shrinkage ~ SUS + EtOH` on the numeric factors, its ANOVA
table, a contour zoomed to the design region, and the standard diagnostic plots
(residuals vs fitted, normal Q–Q, scale–location, residuals vs leverage).

```{r rsm-lm-shrinkage}
lm2 <- lm(Shrinkage ~ SUS + EtOH, data = dat)
summary(lm2)
anova(lm2)
contour(lm2, ~ SUS + EtOH, image = TRUE, bounds = list(EtOH = c(0, 1), SUS = c(0, 0.5)),
        zlim = c(50, 65), decode = FALSE)
lines(c(0.4, 0.24, 0.24, 0.4, 0.4), c(0.2, 0.2, 1, 1, 0.2), col = "red")  # design region
points(0.4, 3, col = "black", pch = 22, cex = 2, bg = "blue")  # validation condition SUS 0.4 / EtOH 3 (outside these axis limits, so not visible here)
par(mfrow = c(2, 2))
lm2 %>% plot()
```

### Additive linear model for Density

The same approach applied to Density as the response.

```{r rsm-lm-density}
lm3 <- lm(Density ~ SUS + EtOH, data = dat)
summary(lm3)
contour(lm3, ~ SUS + EtOH, image = TRUE, bounds = list(EtOH = c(0, 1), SUS = c(0, 0.5)), decode = FALSE)
lines(c(0.4, 0.24, 0.24, 0.4, 0.4), c(0.2, 0.2, 1, 1, 0.2), col = "red")  # design region
points(0.4, 3, col = "black", pch = 22, cex = 2, bg = "blue")  # validation condition SUS 0.4 / EtOH 3 (outside these axis limits, so not visible here)
```

# Structure–property scaling (Gibson–Ashby)

The apparent elastic modulus scales with density following the Gibson–Ashby model for
cellular solids, E\*/E~s~ = C · (ρ/ρ~s~)^n^, with the solid cellulose acetate reference
values E~s~ = 1178 MPa and ρ~s~ = 1311 kg·m⁻³. The model is fitted by nonlinear least
squares on the original scale, which reproduces the constants reported in Polymer Journal
Figure 5b (C = 2.04, n = 2.48): an exponent near 2 indicates bending-dominated
deformation of the cell structure, and C > 1 a comparatively rigid cell arrangement.

```{r gibson-ashby}
Erel   <- (nips$Modulus / 1000) / 1178   # kPa -> MPa, relative to solid CA (Es = 1178 MPa)
rhorel <- nips$Density / 1311            # relative to solid CA (rho_s = 1311 kg m^-3)

ga_fit <- nls(Erel ~ C * rhorel^n, start = list(C = 1, n = 2))
summary(ga_fit)

plot(rhorel, Erel, log = "xy",
     pch = 21, cex = 1.4, bg = as.integer(nips$SUS) + 1,
     xlab = "Relative density", ylab = "Relative modulus",
     main = "Gibson-Ashby scaling")
curve(coef(ga_fit)[["C"]] * x^coef(ga_fit)[["n"]], add = TRUE, col = "red", lwd = 2)
legend("topleft", legend = levels(nips$SUS), pt.bg = seq_along(levels(nips$SUS)) + 1,
       pch = 21, title = "SUS (%)", bty = "n")
```

# Appendix

Kept for reference; not part of the analysis pipeline.

## ASCA (not evaluated)

ANOVA–simultaneous component analysis was explored as an alternative multivariate approach
(factor 1 = SUS, factor 2 = EtOH). The code is kept for reference but not evaluated: it
requires an ASCA implementation such as the archived **MetStaT** package, which is not
loaded in this document.

```{r asca, eval=FALSE, echo=TRUE}
set.seed(1234)

# Design matrix: recode the factor levels as 1/2/3 integers, as ASCA.Calculate expects.
# (The original draft of this block mapped SUS levels "0.27"/"0.33"/"0.44", which do not
# match the study's actual levels of 0.24/0.3/0.4 wt% — corrected here.)
nips.F <- nips[, c("SUS", "EtOH")]
map.1 <- c("0.24" = 1, "0.3" = 2, "0.4" = 3)
nips.F$SUS <- map.1[nips.F$SUS]
map.2 <- c("0.2" = 1, "0.5" = 2, "1" = 3)
nips.F$EtOH <- map.2[nips.F$EtOH]
names(nips.F) <- c("1", "2")
nips.F <- as.matrix(nips.F)

# Response matrix (columns 8–16 of the data).
nips.X <- as.matrix(nips[, 8:16])

# Decompose into factor-1, factor-2 and interaction effects; permutation test.
ASCA.nips <- ASCA.Calculate(nips.X, nips.F, equation.elements = "1,2,12", scaling = TRUE)
ASCA.DoPermutationTest(ASCA.nips, perm = 1000)
```

