---
title: "NIPS-JS Sponge — Response Surface Methodology"
author: "De Nguyen"
output:
  html_document:
    toc: yes
    number_sections: yes
    df_print: default
    fig_width: 10
    fig_height: 6
  word_document:
    toc: true
    fig_width: 10
    fig_height: 6
    toc_depth: 1
    keep_md: true
  pdf_document:
    toc: yes
editor_options:
  markdown:
    wrap: 120
---

```{r setup, include=FALSE}
# Code is hidden in the rendered report (echo = FALSE); read this .Rmd file for the code.
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.retina = 5)
library(rsm)    # response-surface models: FO()/SO(), canonical analysis, contour.lm()
library(dplyr)  # %>% pipe
library(knitr)  # kable(): formatted tables
library(broom)  # tidy(): model output as data frames
# emmeans is called below via emmeans::emmeans() without attaching the package.
```

# About this document

Companion analysis to `NIPS.Rmd`. There, the design factors `SUS` (polymer suspension
concentration, wt%) and `EtOH` (ethanol concentration, wt%) are treated as **categorical**
levels; here they are treated as **continuous** variables, and response surface
methodology (RSM) is used to model `Shrinkage` (and, at the end, `Density`) as a smooth
function of both factors. This view supports interpolation between the tested levels and
indicates in which direction the response decreases fastest.

The first-order Shrinkage model below is the analysis published in the ACS Omega paper:
its coefficients appear there as Table 2, its ANOVA (including the lack-of-fit test) as
Table S6, and its contour plot as Figure 6j. The RSM prediction was validated with three
additional sponges fabricated at EtOH = 3 wt% (one per SUS level; SI Table S1, rows
43–45): the SUS 0.4 / EtOH 3 sponge showed 49.4% shrinkage, consistent with the predicted
trend. The second-order model and canonical analysis further below are exploratory
extensions beyond the published analysis.

The data file `NIPS.csv` contains the full study dataset (licensed CC BY 4.0 — see
`DATA_LICENSE.md`); the column dictionary is in `NIPS.Rmd` / `README.md`. The three
RSM-validation sponges are provided separately in `NIPS_validation.csv` and are not part
of the models below.

# Import data

```{r import-data}
# SUS and EtOH stay numeric here: response-surface models need continuous predictors.
# (NIPS.Rmd, in contrast, converts them to factors.)
dat <- read.csv("NIPS.csv", header = TRUE, row.names = 1)
# head(dat)  # uncomment to preview the imported data
```

# First-order (FO) model

`FO(SUS, EtOH)` fits a first-order response surface — a plane over the two factors. The
`summary()` includes a lack-of-fit test, which indicates whether this simple plane is
adequate or whether curvature (a second-order model) is needed. The predictors enter in
their original units (uncoded), so coefficients are interpretable directly on the
concentration scales. The coefficient table equals ACS Omega Table 2, and the ANOVA block
of the summary equals Table S6 (model P = 0.011; lack of fit insignificant, P = 0.260).

```{r fo-fit}
rsm1 <- rsm(Shrinkage ~ FO(SUS, EtOH), data = dat)
summary(rsm1)
rsm1 %>% tidy() %>% kable(digits = 3)
```

## Contour plot of the fitted plane

Filled contour map of predicted Shrinkage. The red rectangle marks the experimental
(design) region actually covered by the study — predictions outside it are extrapolation
and should be read with caution.

```{r fo-contour}
contour(rsm1, ~ SUS + EtOH, image = TRUE, bounds = list(EtOH = c(0.1, 4), SUS = c(0, 0.5)),
        zlim = c(40, 60), decode = FALSE)
lines(c(0.4, 0.24, 0.24, 0.4, 0.4), c(0.2, 0.2, 1, 1, 0.2), col = "red")  # design region
```

# Second-order (SO) model & canonical analysis

`SO(SUS, EtOH)` adds quadratic and interaction terms, allowing a curved surface. The
canonical analysis then locates the stationary point of that surface and reports the
eigenvalues of the quadratic form: all negative → maximum, all positive → minimum, mixed
signs → saddle. `threshold = 0` keeps the eigenvalues exactly as estimated (no
thresholding of near-zero values).

```{r so-fit}
rsm2 <- rsm(Shrinkage ~ SO(SUS, EtOH), data = dat)
canonical(rsm2, threshold = 0)
```

## Contour plot with the steepest-descent path

The blue circles trace the **path of steepest descent** — the direction along which
predicted Shrinkage decreases fastest from the stationary point (marked with a distinct
symbol at distance 0). The filled blue square highlights the validation condition
SUS = 0.4 wt%, EtOH = 3 wt%, where a fabricated sponge showed 49.4% shrinkage, consistent
with the predicted trend. The red rectangle is again the design region.

```{r so-contour-path}
contour(rsm2, EtOH ~ SUS, image = TRUE, bounds = list(EtOH = c(0.1, 4), SUS = c(0, 0.5)),
        zlim = c(30, 80), decode = FALSE)
lines(c(0.4, 0.24, 0.24, 0.4, 0.4), c(0.2, 0.2, 1, 1, 0.2), col = "red")  # design region
points(EtOH ~ SUS, data = canonical.path(rsm2, descent = TRUE, threshold = 0),
       col = "blue", pch = 1 + 6*(dist == 0))   # pch 7 marks the stationary point (dist == 0)
points(0.4, 3, col = "black", pch = 22, cex = 2, bg = "blue")  # validation condition (SUS 0.4, EtOH 3 wt%; measured shrinkage 49.4%)
```

## Contour plot centred at the stationary point

The same second-order surface, re-drawn with the slice taken through the stationary point
(`at = summary(rsm2)$canonical$xs`) and a wider EtOH range.

```{r so-contour-stationary}
contour(rsm2, EtOH ~ SUS, image = TRUE, at = summary(rsm2)$canonical$xs,
        bounds = list(EtOH = c(0.1, 5), SUS = c(0, 0.5)), zlim = c(40, 100), decode = FALSE)
```

# Predicted means over the design grid

Estimated marginal means of Shrinkage from the first-order model, evaluated at the nine
design combinations of the study (SUS 0.24/0.3/0.4 wt% × EtOH 0.2/0.5/1 wt%;
`mode = "decoded"` reports them on the original concentration scales).

```{r emmeans-grid}
# Note: emmeans reconstructs the data from the model call, so `dat` must exist in the
# workspace — which is the case when this document is knitted normally.
emmeans::emmeans(rsm1, ~ SUS * EtOH, mode = "decoded",
                 at = list(SUS = c(0.24, 0.3, 0.4), EtOH = c(0.2, 0.5, 1)))
```

# Cross-checks with ordinary lm()

The same first-order surface can be fitted with plain `lm()`; these fits reproduce the
RSM results and add the standard regression diagnostics.

## First-degree polynomial fit

`poly(SUS, EtOH, degree = 1)` spans the same model space as `FO(SUS, EtOH)`, so this
contour should match the FO contour above. (The point SUS = 0.4, EtOH = 3 is the same
validation condition as before.)

```{r fo-poly-check}
fo_poly <- lm(Shrinkage ~ poly(SUS, EtOH, degree = 1), data = dat)
contour(fo_poly, ~ SUS + EtOH, image = TRUE, bounds = list(EtOH = c(0.1, 4), SUS = c(0, 0.5)),
        zlim = c(40, 60), decode = FALSE)
lines(c(0.4, 0.24, 0.24, 0.4, 0.4), c(0.2, 0.2, 1, 1, 0.2), col = "red")  # design region
points(0.4, 3, col = "black", pch = 22, cex = 2, bg = "blue")  # validation condition (SUS 0.4, EtOH 3 wt%; measured shrinkage 49.4%)
```

## Additive linear model for Shrinkage, with diagnostics

Plain additive regression `Shrinkage ~ SUS + EtOH` on the numeric factors, its ANOVA
table, a contour zoomed to the design region, and the standard diagnostic plots
(residuals vs fitted, normal Q–Q, scale–location, residuals vs leverage).

```{r lm-shrinkage}
lm2 <- lm(Shrinkage ~ SUS + EtOH, data = dat)
summary(lm2)
anova(lm2)
contour(lm2, ~ SUS + EtOH, image = TRUE, bounds = list(EtOH = c(0, 1), SUS = c(0, 0.5)),
        zlim = c(50, 65), decode = FALSE)
lines(c(0.4, 0.24, 0.24, 0.4, 0.4), c(0.2, 0.2, 1, 1, 0.2), col = "red")  # design region
points(0.4, 3, col = "black", pch = 22, cex = 2, bg = "blue")  # validation condition SUS 0.4 / EtOH 3 (outside these axis limits, so not visible here)
par(mfrow = c(2, 2))
lm2 %>% plot()
```

## Additive linear model for Density

The same approach applied to Density as the response.

```{r lm-density}
lm3 <- lm(Density ~ SUS + EtOH, data = dat)
summary(lm3)
contour(lm3, ~ SUS + EtOH, image = TRUE, bounds = list(EtOH = c(0, 1), SUS = c(0, 0.5)), decode = FALSE)
lines(c(0.4, 0.24, 0.24, 0.4, 0.4), c(0.2, 0.2, 1, 1, 0.2), col = "red")  # design region
points(0.4, 3, col = "black", pch = 22, cex = 2, bg = "blue")  # validation condition SUS 0.4 / EtOH 3 (outside these axis limits, so not visible here)
```
nips <- within(nips, SUS <- relevel(SUS, ref = 1))
nips <- within(nips, EtOH <- relevel(EtOH, ref = 1))
# nips
```


```{r}
rsm1 <- rsm(Shrinkage ~ FO(SUS, EtOH), data = dat)
summary(rsm1)
rsm1 %>% tidy() %>% kable(digits = 3)
```

```{r}
contour(rsm1, ~ SUS + EtOH, image = T,bounds = list(EtOH = c(0.1, 4), SUS = c(0, 0.5)), 
        zlim = c(40, 60), decode = FALSE)
lines(c(0.4,0.24,0.24,0.4,0.4), c(0.2,0.2,1,1,0.2), col = "red") # design region
```

```{r}
rsm2 <- rsm(Shrinkage ~ SO(SUS, EtOH), data = dat)
canonical(rsm2, threshold = 0)
```

```{r}
contour(rsm2, EtOH ~ SUS, image = T, bounds = list(EtOH = c(0.1, 4), SUS = c(0, 0.5)), 
        zlim = c(30, 80), decode = FALSE)
lines(c(0.4,0.24,0.24,0.4,0.4), c(0.2,0.2,1,1,0.2), col = "red") # design region
points(EtOH ~ SUS, data = canonical.path(rsm2, descent = T, threshold = 0), 
        col = "blue", pch = 1 + 6*(dist == 0))
points(0.4, 3, col="black", pch = 22, cex = 2, bg = "blue")
```

```{r}
contour(rsm2, EtOH ~ SUS, image = T, at = summary(rsm2)$canonical$xs, bounds = list(EtOH = c(0.1, 5), SUS = c(0, 0.5)), zlim = c(40, 100), decode = F)
```


```{r}
emmeans::emmeans(rsm1, ~ SUS * EtOH, mode = "decoded", 
        at = list(SUS = c(0.24, 0.3, 0.4), EtOH = c(0.2, 0.5, 1)))
```

```{r}
lm <- lm(Shrinkage ~ poly(SUS, EtOH, degree = 1), data = dat)
contour(lm, ~ SUS + EtOH, image = T,bounds = list(EtOH = c(0.1, 4), SUS = c(0, 0.5)), 
        zlim = c(40, 60), decode = FALSE)
lines(c(0.4,0.24,0.24,0.4,0.4), c(0.2,0.2,1,1,0.2), col = "red") # design region
points(0.4, 3, col="black", pch = 22, cex = 2, bg = "blue")
```


```{r}
lm2 <- lm(Shrinkage ~ SUS + EtOH, data = dat)
summary(lm2)
anova(lm2)
contour(lm2, ~ SUS + EtOH, image = T,bounds = list(EtOH = c(0, 1), SUS = c(0, 0.5)), 
        zlim = c(50, 65), decode = FALSE)
lines(c(0.4,0.24,0.24,0.4,0.4), c(0.2,0.2,1,1,0.2), col = "red") # design region
points(0.4, 3, col="black", pch = 22, cex = 2, bg = "blue")
par(mfrow = c(2,2))
lm2 %>% plot()
```

```{r}
lm3 <- lm(Density ~ SUS + EtOH, data = dat)
summary(lm3)
contour(lm3, ~ SUS + EtOH, image = T,bounds = list(EtOH = c(0, 1), SUS = c(0, 0.5)), decode = FALSE)
lines(c(0.4,0.24,0.24,0.4,0.4), c(0.2,0.2,1,1,0.2), col = "red") # design region
points(0.4, 3, col="black", pch = 22, cex = 2, bg = "blue")
```

